### 2019-07-19 vesicle_code/docs/adhesion/revisions/summary_of_changes.tex

\documentclass[11pt]{article}

\usepackage{fullpage}
\usepackage{todonotes}
\usepackage{amsmath,amsfonts,amssymb,stmaryrd}
\usepackage{color}
\newcommand{\comment}[1]{{\color{blue} #1}}
\newcommand{\xxi}{\boldsymbol{\xi}}

\begin{document}
\noindent
We thank the referees for their comments.   We have made revisions based on these comments and criticisms.
Below is a list of changes in the revision.
\\ \\ \noindent 
Sincerely, \\ \noindent
Yuan-Nan Young, Shravan Veerapaneni, and Bryan Quaife

\vspace{20pt}

\begin{enumerate}

\item We modified the abstract.

\item We modified the introduction with better clarity and more complete citation of references.

\item We modified all the figure captions to make sure that all the figures are self-contained.

\item In Section IV: Adhesion of two vesicles in a planar extensional
  flow, we conducted new simulations to elucidate the interaction
    between two vesicles in a planar extensional flow. These results are
    summarized in Figure 6.

\item In Section IV: Adhesion of two vesicles in a planar extensional flow, we added a set of simulations with reduced area $\triangle A=0.65$ and these results are summarized in the new Figure 10(b) and 10(c).

\item In Section V: Adhesion of two vesicles in a shear flow, we conducted new  simulations with higher Hamaker constant. These results are summarized in Figure 12.

\item In Section V: Adhesion of two vesicles in a shear flow, we conducted more simulations to examine the oscillation of the vesicle doublet and how the oscillation frequency depends on 
the shear rate, separation distance, and Hamaker constant. These results are summarized in Figure 13(b).

\item In Section V: We did some analysis to compare the double
  oscillation with the Jeffery orbit of a rigid ellipse in shear flow.
    These results are summarized in Table I.

\item In Section V: Adhesion of two vesicles in a shear flow, we
  conducted more simulations to examine the effect of the separation
    distance $\delta$ based on reviewers' suggestions.

\item In Section V: Adhesion of two vesicles in a shear flow, we
  corrected the calculation of the intrinsic viscosity and the new
    results are summarized in Figure 15.  We also discussed the effect
    of the separation distance on the intrinsic viscosity.

\item In Section VI: We have modified the conclusions to include a
  discussion of our new results.

\item In Acknowledgements,  a sentence is added to describe the contribution from each author.

\end{enumerate}
\end{document}