### 2019-07-21 vesicle_code/docs/adhesion/revisions/reviewer1.tex
finished rebuttal for reviewer 1

\documentclass[11pt]{article}

\usepackage{fullpage}
\usepackage{todonotes}
\usepackage{amsmath,amsfonts,amssymb,stmaryrd}
\usepackage{color}
\newcommand{\comment}[1]{{\color{blue} #1}}

\begin{document}
\noindent
We thank the referee for the comments.  Below are our point-to-point
responses to all the reviewer 1's comments.  We have made revisions
based on these comments.  We also note that a few typos from the
original manuscript have been corrected.  A separate list of changes can
be found along with this list of responses.

\medskip
\noindent 
Sincerely, \\ \noindent
Yuan-Nan Young, Shravan Veerapaneni, and Bryan Quaife

\vspace{20pt}
\noindent

\begin{enumerate}
\item\comment{While the numerical simulations performed are based on standard
approaches, and the results are supposedly correct, I find the results
not to have a significant impact. Most results presented, for example,
those on vesicle adhesion in quiescent flow, adhered vesicle dynamics in
shear flow, and adhesion of two approaching vesicles in a shear flow,
are not surprising, and similar results were obtained before for red
blood cells, capsules, drops, and adhered particles. Most (maybe all) of
the principals discussed are more-or-less understood, and would
therefore not of themselves constitute a novel contribution.}
\begin{itemize}
\item We thank the referee for this comment. The referee is absolutely
  correct by labeling boundary integral simulation as a standard
    approach. However, the boundary integral algorithm as applied to the
    adhesion problem here is quite novel for two reasons: (1) First, we
    applied a new adaptive time marching developed recently in Quaife
    and Biros {\em Journal of Computational Physics}, {\bf
    306}:478--499, 2016, allowing us to efficiently run long
    time-horizon simulations. (2) Second, the close vesicle-vesicle
    interactions in the doublet case is quite challenging to capture
    accurately; here, we employed a novel near-singular integral
    evaluation scheme which is an extension of the method developed in
    Quaife and Biros {\em Journal of Computational Physics}, {\bf
    274}:245--267, 2014. Without adaptive time marching, the boundary
    integral simulations of vesicles often give unreliable results as
    shown in Kabacao\u{g}lu {\em et al.}, {\em Journal of Computational
    Physics}, {\bf 357}:43--77, 2018. In addition it is very time
    consuming to simulate two vesicles to equilibrate in close vicinity
    of each other in a quiescent flow.  Without a good near-singular
    evaluation in the boundary integral calculation, the simulations
    will be contaminated by the errors incurred when two moving
    deformable boundaries are close to each other.

 We thank the referee for pointing out that similar physical setup has
    been studied for red blood cells, capsules, drops, and adhering
    particles. However, it is not clear to us if our results for
    vesicles are so similar to those for red blood cells, drops, and
    particles from the literature. For example, we are able to verify
    the scaling of the draining time with respect to adhesion strength
    and equilibrium separation distance by analyzing the numerical
    simulation results.  A similar scaling is found in Ramachandran {\em
    et al.}, Phys. Fluids (2010) where they assumed the vesicles are
    spherical cap with a flat contact region.  It is not clear how their
    scaling for the draining time can differentiate the contribution
    from either adhesion strength or equilibrium separation distance,
    while our scaling clearly illustrates how each component contributes
    differently to the draining time. It is also not clear how such
    scaling can be verified by comparing against experimental data.

For the vesicles in a planar extensional flow, we cannot find any work
    in the literature that shows how two red blood cells or capsules move
    toward each other around the stagnation point. Furthermore we cannot
    find any work in the literature that uses extensional flow to
    measure the adhesion strength between two
    interfaces/particles/membranes. All the work we can find on a drop
    in an extensional flow/four-roll mill is for a single drop (clean
    drop or a surfactant-laden drop) or a vesicle.

For the vesicles in a planar shear flow, there are numerical simulations
    and experiments on how two drops/vesicles interact with each other.
    However, to our best attempt, we did not find any work on the
    effects of adhesion on drop-drop/vesicle-vesicle interactions in a
    shear flow (possibly due to the numerical issues that we mentioned
    earlier). Our simulation results show that (1) adhesion
    significantly affects the vesicle-vesicle interaction in the shear
    flow, and in the revision we provided means to quantify such
    effects; and (2) adhesion also induces additional dissipation in a
    dilute suspension of vesicle doublets. To the best of our knowledge,
    these results are not available in the literature. Even though our
    simulations are two-dimensional, previous works have demonstrated
    that two-dimensional vesicle hydrodynamics bear some general
    similarity to their three-dimensional counterparts.

{\bf Changes:} We made revisions with this comment in mind, to make sure
that our results are highlighted and the readers can see clearly how
these results are unique and maybe useful.

\end{itemize}

\noindent
\item \comment{Second, the simulations are 2D, which makes it difficult
  to accept whether the predicted results would be observed in
  experiments. There was also no direct comparison with any experimental
  studies.}
\begin{itemize}
  \item We thank the referee for this comment. The referee is absolutely
    correct: two-dimensional and three-dimensional hydrodynamics are in
    general very different. There are cases where the two-dimensional
    and three-dimensional vesicle hydrodynamics are in good agreement.
    For example, the two-dimensional simulations of packing of RBCs in
    Flormann {\em et al.}, Sci.~Rep.~(2017) are shown to be in good
    agreement with both experiments and three-dimensional simulations.
    In addition, the two-dimensional theoretical prediction of the
    buckling instability of the contact region between RBCs is validated
    by comparing with experiments in that paper.  Another example where
    two-dimensional vesicle hydrodynamics in shear flow is in general
    agreement with the three-dimensional vesicle hydrodynamics is
    illustrated in Ghigliotti {\em et al.}, J.~Fluid Mech.~(2010).

  The current work is the first step we took to investigate the effects
    of adhesion on vesicle hydrodynamics. We are working on the
    three-dimensional version and with the two-dimensional results as a
    guidance we will be able to elucidate more physical insight when we
    analyze the results from three-dimensional simulations in the near
    future.

  \noindent
  {\bf Changes:} We have added some comparison with three-dimensional
    (both numerical and experimental) results of two RBCs in shear flow
    in the revision.
\end{itemize}

\noindent
\item\comment{My third concern is the use of L-J potential. In practice, the
vesicle surface would be coated with some adhesive molecules (e.g.
receptor/ligand). Is the L-J type adhesive field appropriate for such
interactions? Also, in reality, these bonds are stochastic in nature.}
\begin{itemize}
  \item We thank the referee for this point.  In this work we focus on
    the adhesion effects on the hydrodynamics of two vesicles drawn
    together by an external flow.  Thus the L.-J.~type potential is to
    keep the equilibrium distance, following the modeling works on red
    blood cells (RBCs) by Flormann {\em et al.}, Sci.~Rep.~(2017) and
    Hoore {\em et al.} Soft Matt.!(2018).  In both papers the
    L.-J.~potential is used to model the interaction between RBCs. The
    focus of this work is to show the hydrodynamic implication of such
    interaction between two vesicles without any coating of adhesive
    protein macromolecules. Results from these modeling works using a
    L.-J.~potential are found to be in good agreement with experiments
    and previous modeling results using different adhesion potential
    (Ziherl and Svetina, PNAS (2007)).

    The referee is correct that the biological adhesion requires
    physical contact and is stochastic in nature. We are modeling such
    stochastic binding for adhesion between two membranes in a viscous
    solvent.

  \noindent
  {\bf Changes:} We have added more references and a better description
    on the usage of L.-J.~potential for vesicle adhesion to explain its
    validity and applicability.

\end{itemize}

\noindent
\item\comment{The authors talks about the ''fluid trap". This is indeed
  a tricky issue in any experiment dealing with extensional flow as how
  to maintain the drop/vesicle at the stagnation point. But in a
  numerical simulation this should not be any issue at all., because
  even of the vesicle moves away from the stagnation point its dynamics
  is still governed by the velocity gradient which is same everywhere in
  the flow field.}
\begin{itemize}
  \item We thank the referee for this comment. The ``fluid trap", in its
    earliest form, was first achieved in an experiment by Bentley and
    Leal in 1986: A computer-controlled four-roll mill for
    investigations of particle and drop dynamics in two-dimensional
    linear shear flows, {\em Journal of Fluid Mechanics} {\bf
    167}:219-240, 1986.  Bentley and Leal showed that their
    computer-aided feedback control in the four-roll mill can keep the
    particle/drop at the center of the four-roll mill for a very long
    time.  Over the years the techniques of using the four-roll mill as
    a fluid trap have been improved and extended to microfluidics (see
    Hudson {\em et al.}, Applied Phys.~Lett.~(2004) and Lee {\em et
    al.}, Applied Phys.~Lett.~(2007), for example.) In the original
    manuscript we simulate the fluid trap based on the setups by both
    the Schroeder's group at UIUC and S. Muller's group that can
    efficiently control the location of the stagnation point.

    Our motivation is to present these simulation results so the
    experimentalist may be interested to use the fluid trap as a
    non-invasive tool to measure the adhesion strength between vesicles.
    We hope that there will be experimental data for us to compare our
    simulations results in the near future.

   \noindent
  {\bf Changes:} We have added these references on fluid trap in the
    revised manuscript. In addition we have rewritten a couple of
    relevant paragraphs to convey our motivation.
 \end{itemize}

\noindent
\item\comment{There was not enough physical insights on the vesicle
  alignment in extensional flow as presented in section IV. If the
  vesicles are placed centrally, what causes the pair to become
  asymmetric in the first place, since all components in the model are
  deterministic? The authors used vorticity contours, but the external
  flow is irrotational; so what is the origin of vorticity, and is it
  appropriate to use vorticity to explain the behavior?}
\begin{itemize}
  \item We have reorganized section IV based on the referee's feedback.
    To investigate the vesicle hydrodynamics in a planar extensional
    flow, we conducted new simulations where a left vesicle is placed at
    the stagnation point and the right vesicle is moving towards it.
    Depending on the initial elevation of the right vesicle off the
    $x$-axis, the vesicle hydrodynamics is different and we find
    configurations that are similar to those that we observe in the
    fluid trap. The simulation data shows that the membrane tension is
    highly non-uniform, and we speculate that the Marangoni stress
    contributes to asymmetric configuration of vesicles around the
    stagnation point, as for a single surfactant-laden drop in an
    extensional flow.

  \noindent {\bf Changes:} We have removed the part on visualizing flow
    around vesicles in the fluid trap. Instead we added simulations at
    the beginning of new Section IV to examine the vesicle hydrodynamics
    in a planar extensional flow.
\end{itemize}

\noindent
\item\comment{I did not find figure 14 relevant for the present article.
  This result is not relevant, and also presented just to show what the
  authors intend to do in future.}
\begin{itemize}
  \item We agree with the referee and  have removed this example from
    the manuscript.

    \noindent
    {\bf Changes:} We removed this example from the conclusion and
    instead we expand future work to include a suspension of many
    vesicles.
\end{itemize}

\end{enumerate}

\end{document}