### 2019-07-21 vesicle_code/docs/adhesion/revisions/reviewer2.tex
finished reviewer 2 reubuttal

\documentclass[11pt]{article}

\usepackage{fullpage}
\usepackage{todonotes}
\usepackage{amsmath,amsfonts,amssymb,stmaryrd}
\usepackage{color}
\newcommand{\comment}[1]{{\color{blue} #1}}
\newcommand{\xxi}{\boldsymbol{\xi}}

\begin{document}
\noindent
We thank the referee for the comments.  Below are our point-to-point
responses to all the comments from the reviewer.  We have made revisions
based on these comments.  We also note that a few typos from the
original manuscript have been corrected.  A separate list of changes can
be found along with this list of responses.

\medskip
\noindent 
Sincerely, \\ \noindent
Yuan-Nan Young, Shravan Veerapaneni, and Bryan Quaife

\vspace{20pt}

\begin{enumerate}
\item\comment{General comment: I found that the figure captions are often
incomplete and that details of the legends are missing and are only
given in the text. I think figures should be self explanatory (e.g.
signification of symbols of fig. 9 right (angles corresponding to symbol
sizes); in fig. 10: scale/unit of vorticity, Fig 11: indicate in the
caption that the coloraturas corresponds to tension; Fig 12: color code
in figure caption...)}
\begin{itemize}
  \item The captions now do a more complete job of explaining the
    figure.  We paid particular attention to Figure 10 (old Figure 9)
    since it contains a lot of information.
\end{itemize}

\item\comment{Abstract: it should be mentioned in the abstract that this is
2D modelling. 2D simulations of vesicles are often perfectly fine to
investigate the physics of phenomena but in some cases 3D effects may
have their importance, at least for quantitative comparisons with
experiments. The reader should identify either from the title or
abstract that this study is 2D.}
\begin{itemize}
  \item We now mention in the abstract that we consider the dynamics of
    and adhesive two-dimensional vesicle doublet.
\end{itemize}

\item\comment{Introduction, 2nd paragraph: the transition in the effective
viscosity of vesicle or RBC suspensions was also validated
experimentally in ref [18] (Vitkova et al.), not only in [20-22].}
\begin{itemize}
  \item We have included reference [18] in the list of experimental
    works that validated the transition in the effective viscosity
\end{itemize}

\item\comment{Intro, 3rd paragraph: the buckling instability of aggregating RBC
membranes was not only predicted theroretically and numerically in
Flormann et al.~[32] but also experimentally. In addition, a more
recent paper by Hoore et al.~(Soft Matter 14, 6278-6289 (2018)) showed
the influence of the membrane shear elasticity on doublet shapes by 3D
modelling. This reference may be added.}
\begin{itemize}
  \item The paper by Hoore et al.~has been added to the introduction.
\end{itemize}

\item\comment{Concerning the stability and dynamics of clusters of
  vesicles with adhesive interaction, there have been at least two
    papers on clusters in micro-channel flow (computational and
    experimental) that should be cited: Brust et al., Scientific Reports
    4, 4348 (2014) and Claveria et al., Soft Matter. 2016 vol
    12(39):8235-8245. About rheology, at least the historical paper by
    Chien on the shear-thinning of blood related to RBC aggregation
    should be cited : Chien, S., Usami, S., Dellenback, R.~J.,
    Gregersen, M.~I., Nanninga, L.~B.~\& Guest, M.~M. Blood viscosity:
    Influence of erythrocyte aggregation. Science 157, 829–831 (1967).}
\begin{itemize}
  \item The reference of Brust et al.~has been added.
  \item The reference of Claveria et al~has been added.
  \item The reference of Chien et al.~has been added.
\end{itemize}

\item\comment{In addition to ref [54], Gires et al. also published experimental
measurements of hydrodynamic interactions between vesicles in Gires et
al., Phys.~Fluids 26, 013304 (2014)}
\begin{itemize}
  \item The reference of Gires et al.~has been added
\end{itemize}

\item\comment{In II. Governing equations and fig. 1: the right panel of the
figure and comments at the end of the section say that the part of
vesicle membranes that is out of the contact zone, including the parts
that are facing the outer fluid, experience an attractive interaction.
This is an artifact due to the long-range potential chosen by the
authors to model the interaction. In most experimental systems, the
adhesive forces leading to cluster formation are either due to direct
bridging of membranes or depletion phenomena that lead to osmotic
pressure gradients. This is the case for RBC aggregation for instance.
In both mechanisms, there is indeed no attractive force between membrane
parts that are out of the contact zone. It is perfectly fine to model
the interaction by an adhesion potential such as represented in fig 1
left, but its use introduces this artifact that should be acknowledged
by the authors. It is also due to the fact that the value of parameter
$\delta$ (the equilibrium distance) is not small compared to vesicle
dimensions. This is a numerical necessity to avoid the need to have
extremely refined meshes but again in experiments, the equilibrium
distance is usually much smaller that vesicle sizes and is more
comparable to molecular sizes (size of bridging molecules or of
depletion-inducing macromolecules). This deserves to be acknowledged
also, even though it may not impact too much the results.}
\begin{itemize}
  \item The referee is correct that the long-range attraction in the
    L.-J.~is an artifact.  In this work we focus on the adhesion effects
    on the hydrodynamics of two vesicles drawn together by an external
    flow.  Thus the L.-J.~type potential is to keep the equilibrium
    distance, following the works on red blood cells by Flormann {\em et
    al.}~Scientific Reports (2017) and Hoore {\em et al.}~Soft Matter
    (2018).  In both these papers, the L.-J.~potential is used to model
    the adhesive interaction between red blood cells.  Results from
    these modeling works are found to be in good agreement with
    experiments and other models for adhesion (Ziherl and Svetina, PNAS
    (2017)).

  \item We have added more references and a better description on the
    usage of the Lennard-Jones potential for vesicle adhesion to explain
    its validity and applicability.

  \item We have added additional simulations that use a smaller value
    for the separation distance $\delta$.  First, this demonstrates that
    we are able to stably simulate adhering vesicles where the
    lubrication layer is thin.  Second, the decreased value of $\delta$
    gives similar results.  For example, in Figure 14, we show
    numerically that, for two different separation distances, there is a
    linear relationship between the shear rate and Hamaker constant that
    determines if two vesicles in a shear flow will form a doublet or
    not.
\end{itemize}

\item\comment{Along the same line, the repulsing spots in fig 1 right or Fig 3
right as well as the amplitude of the deformation in the contact zone
are again a consequence of the choice of parameter values for the
interaction potential and would probably not be so strong in real
cases where the equilibrium distance between membranes is much shorter
and the range of the potential is shorter.}
\begin{itemize}
  \item As mentioned in the Introduction and Section II, we are focusing
    on intermediate adhesive strengths.  In the Introduction, we give
    several references where strong adhesion is studied, but the effect
    of the intermediate adhesion strengths on the vesicle dynamics has
    not received much attention.
\end{itemize}

\item\comment{I cannot figure out why the external part of vesicles on
  profile 1 in fig 3-left is yellow. Shouldn't the force be weakly
  attractive in that region, as in fig 1-right?}
\begin{itemize}
  \item In Figure 3 left, the color is not the adhesive force---it is the
    tension.  This is mentioned in the caption.  However, in the plot on
    the right that shows a zoomed in view of the contact zone, the
    adhesive force is plotted using the same method as in Figure 1.  In
    this case, the color is yellow only if the vesicle is sufficiently
    close to the other.
\end{itemize}

\item\comment{Figs 4-5 and related text: it would have been interesting
  to compare results with the bending energy of vesicles without
  adhesion potential in their equilibrium shape, to show the energy gain
  due to adhesion (bending and total energy). This could be a way to
  compare the total energy of the doublet to the energy of dissociated
  vesicles and infer about the stability of doublets.}
\begin{itemize}
  \item We thank the referee for this suggestion. We did follow the
    referee's suggestion to check out this comparison. As pointed out by
    the referee in the seventh comment, the long-range attraction of the
    L.-J.~potential that is used for membrane-membrane adhesion is not
    realistic and consequently the doublet always has lower total energy
    than dissociated vesicles. We are working on a more realistic
    adhesion potential for RBCs and we believe that this comparison will
    reveal more information.
\end{itemize}

\item\comment{Fig 9(right): I regret that the authors did not extend the
  study to lower values of the reduced area, to reach values comparable
  to red blood cells which are more deflated that $\Delta A=0.7$. Also,
  the range of shear rates could have been slightly increased to reach
  the stability limit of the most deflated vesicles. Here, we do not
  know what happens below reduced areas of 0.8.}
\begin{itemize}
  \item We have extended the results to include vesicles of reduced area
    0.65, a typical value for two-dimensional red blood cells.
  \item We have also reported results for larger extensional rates so
    that the transition from a fluid trap to two separating vesicles can
    be observed at all the reduced areas.
\end{itemize}

\item\comment{Figs 9(right) and 12(right): the stability diagrams are
  presented in different parameter spaces for the extensional and shear
  flow cases. It would have been interesting to see e.g. the diagram in
  the interaction energy vs.~flow strength $\chi$ for both cases, for
  comparison.}
\begin{itemize}
  \item In the revision for the extensional example we first focus on
    how an initial vertical displacement may affect the vesicle dynamics
    in a planar extensional flow (beginning of the new Section III).
    There we find that a vesicle doublet is not likely to form unless
    the two vesicles are aligned on the converging axis of the
    extensional flow. In the fluid trap simulations we find that the
    reduced area has a large effect on determining if the doublet is
    broken or not, and on the inclination angle if the doublet is not
    broken.  Therefore, we have elected to investigate the effect of the
    reduced area for this example.

  \item For the shear flow simulations, we place the two vesicles on the
    same horizontal axis at $t=0$ (initial displacement is zero).  Such
    head-on collisions have been well-studied for drops (clean or
    surfactant-laden), capsules, vesicles, and red blood cells (without
    the L.-J.~potential between membranes), and will serve as a good
    reference point to draw comparison with our results. For head-on
    collisions between two vesicles, we find that  a smaller reduced
    area will increase the size of the contact area, and this will
    affect the critical Hamaker constant required to form a doublet.
    Thus we expect that the critical Hamaker constant for vesicle
    doublet formation reported will depend on the reduced area, vertical
    displacement and the shear rate.  To avoid diverting the focus of
    the paper, we have elected to divert an investigation of this
    behavior to future work.
\end{itemize}

\item\comment{Fig.~10: the black arrows, as I understand, represent the
  total attractive force due to the interaction potential. Would the
  authors be able to represent also the total hydrodynamic forces on
  vesicles which presumably balance the torque of interaction forces? Is
  there a tank-treading motion of membranes?}
\begin{itemize}
  \item We have removed Figure 10 from the previous version.

  \item There is a transient tank-treading motion that eventually
    (asymptotically) stops towards equilibrium.  Since we removed Figure
    10 from the original version, we have chosen to not focus on the
    transient tank-treading motion and the vorticity and torque of the
    vesicles.
\end{itemize}

\item\comment{Fig 12: the authors did not comment the evolution of the
  frequency of distance oscillations in doublets when the Hamaker
  constant is varied. Can it be compared to the natural rotation
  frequency of the shear flow $\chi/2$? Does a strongly adhesive doublet
  tend to behave as a tumbling rigid object following something similar
  to Jeffery orbits?}
\begin{itemize}
  \item We now report the period of the doublets at three different
    shear rates and at a variety of Hamaker constants and separation
    distances.

  \item For each of the shear rates and two of the Hamaker constants
    that we consider, we compute an effective aspect ratio which is the
    aspect ratio of a rigid body ellipse that would undergo the same
    period in the same shear flow.  The results indicate that the
    effective aspect ratio depends on the shear rate, and this suggests
    that the dynamics of the doublet in a shear flow is more complicated
    than Jeffery orbits.
\end{itemize}

\item\comment{Fig 12-right: is the stability limit between the red and
  blue zones linear? This would mean that a stability criterion could
  take the form of a dimensionless number involving the ratio of Hamaker
  constant/shear rate, a potentially interesting result.}
\begin{itemize}
  \item We thank the referee for this observation.  We increased the
    parameter space of the phase diagram, and it does appear that there
    is a linear relationship.

  \item We have superimposed a straight line to show the transition
    between the two doublet and unbounded configurations.

  \item We also have repeated the phase diagram, but with a smaller
    separation distance.  With this new separation distance, the
    transition between the two regions appears to also be linear.
\end{itemize}

\item\comment{Fig 13 and comments about rheology: I am not sure to
  understand the meaning of the left figure where it is shown that the
  intrinsic viscosity of the doublet in suspension is about twice the
  viscosity of a single vesicle. By definition, the intrinsic viscosity
  does not depend on volume fraction, so I do not expect such a big
  difference between bound vesicles and isolated vesicles. With the
  values chosen for the Hamaker constant and the shear rate, it seems
  (from fig 12-left) that the interaction is quite strong and that the
  distance between the two vesicles in the doublet is always very short
  (close to the equilibrium distance $\delta$). So the doublet must be
  roughly tumbling as a single object and unless a co-rotating
  tank-treading motion of the two membranes produces a strong
  dissipation by shear flow in the gap between the two vesicles, the
  difference should not be so large. It is also quite curious that the
  intrinsic viscosity of the doublet is so close to exactly twice the
  viscosity of a single vesicle. Could the authors check and elaborate
  on this? Was the volume fraction properly scaled out here? Also in the
  right panel, the bending/tension part is compared to twice the
  intrinsic viscosity of a suspension of isolated vesicles. Why twice?
  Finally, it is not clear how technically the bending/tension and
  adhesion parts are separated.  Overall I think the study on rheology
  needs much more development than what is presented here with only one
  set of parameter values and very little quantitative comments.}
\begin{itemize}
  \item The referee is correct that we miscalculated the intrinsic
    viscosity.  We had initially divided by only the area inside a
    single vesicle rather than both vesicles in the doublet.  This has
    been corrected, and now there is only a small increase in the
    intrinsic viscosity when compared to a single tank treading vesicle.

  \item With the numbers corrected, the intrinsic viscosity is clearly
    not equal to the intrinsic viscosity of a tank treading vesicle.

  \item We have double-checked our calculations and the volume fraction
    has been properly accounted for.

  \item Using twice the intrinsic viscosity of a tank treading vesicle
    was incorrect.  We now compare the intrinsic viscosity of a single
    tank treading vesicle with the doublet.

  \item We also report numbers for the intrinsic viscosity for a larger
    Hamaker constant that has a similar period to $\mathcal{H} = 0.7$.
    The result is an increase in the intrinsic viscosity due to
    adhesion, a decrease in the intrinsic viscosity due to bending and
    tension, and a net increase in the intrinsic viscosity.

  \item The total force $\xxi$ is the sum of a bending, tension, and
    adhesion force.  Therefore, the contribution to the intrinsic
    viscosity due to each of these forces can be individually computed.
\end{itemize}

\item\comment{In the conclusions, an example of aggregation of vesicles
  enclosed in a bigger one is presented. I do not think this brings much
  to this paper. Why not keep this for a future paper dedicated to this
  specific problem and instead show examples of aggregates without the
  enclosing large vesicle? In what conditions (reduced volume, Hamaker
  constant) can linear structures (rouleaux) be formed for instance (see
  the paper by Hoore et al whose reference is given above for
  instance).}
\begin{itemize}
  \item We have removed this example from the manuscript.  We thank the
    referee for the great suggestion to examine the conditions for
    formation of linear structures. We have added a few new results in
    the revision and consequently we mention that future work includes a
    suspension of many vesicles.
\end{itemize}

\end{enumerate}

\end{document}