### 2020-12-29 vesicle_code/docs/permeableVes/main.tex
first draft of supplementary material

\documentclass[9pt,twocolumn,twoside,lineno]{pnas-new}
% Use the lineno option to display guide line numbers if required.

\templatetype{pnasresearcharticle} % Choose template 
% {pnasresearcharticle} = Template for a two-column research article
% {pnasmathematics} %= Template for a one-column mathematics article
% {pnasinvited} %= Template for a PNAS invited submission

%\title{Semi-Permeable Vesicles: Mathematical Formulation, Numerics, and
%Physics}
\title{Effects of Water Permeability on the Hydrodynamics of a Vesicle Under Flow and Confinement}

% Use letters for affiliations, numbers to show equal authorship (if applicable) and to indicate the corresponding author
\author[a]{Bryan Quaife}
\author[a]{Ashley Gannon}
\author[b,1]{Y.-N. Young}

\affil[a]{Department of Scientific Computing, Florida State University}
\affil[b]{Department of Mathematical Sciences, New Jersey Institute of
Technology}

% Please give the surname of the lead author for the running footer
\leadauthor{Quaife}

% Please add a significance statement to explain the relevance of your work
\significancestatement{Authors must submit a 120-word maximum statement
about the significance of their research paper written at a level
understandable to an undergraduate educated scientist outside their
field of speciality. The primary goal of the significance statement is
to explain the relevance of the work in broad context to a broad
readership. The significance statement appears in the paper itself and
is required for all research papers.}

% Please include corresponding author, author contribution and author declaration information
\authorcontributions{All authors contributed equally.}
%\authordeclaration{Please declare any competing interests here.}
%\equalauthors{\textsuperscript{1}A.O.(Author One) contributed equally to this work with A.T. (Author Two) (remove if not applicable).}
%\correspondingauthor{\textsuperscript{1} E-mail: bquaife\@fsu.edu}
\correspondingauthor{\textsuperscript{1} E-mail: yyoung@njit.edu}

% At least three keywords are required at submission. Please provide three to five keywords, separated by the pipe symbol.
\keywords{Keyword 1 $|$ Keyword 2 $|$ Keyword 3 $|$ ...}

\begin{abstract}
We describe a mathematical model for a semi-permeable vesicle and derive
  a boundary integral equation formulation for simulating its
  hydrodynamics.

Please provide an abstract of no more than 250 words in a single paragraph. Abstracts should explain to the general reader the major contributions of the article. References in the abstract must be cited in full within the abstract itself and cited in the text.
\end{abstract}

\dates{This manuscript was compiled on \today}
\doi{\url{www.pnas.org/cgi/doi/10.1073/pnas.XXXXXXXXXX}}

\input my_macros2.tex

\newif\ifTikz
%\Tikztrue
\Tikzfalse

\usepgfplotslibrary{external}
\tikzexternalize

\begin{document}

\maketitle
\thispagestyle{firststyle}
\ifthenelse{\boolean{shortarticle}}{\ifthenelse{\boolean{singlecolumn}}{\abscontentformatted}{\abscontent}}{}

\dropcap{W}ater exchange is essential for a living cell to adapt to its environment over a wide range of timescales \cite{CadartVenkovaRechoEtAl2019_NaturePhys,AlbertsMolecularBiology,YangMaVerkman2001_JBC,SugieIntaglietta2018_AmJPhysiolHCP,SaadounPapadopoulosWatanabeEtAl2005_JCS,Verkman2008_JMM,BerthaudEtAl2016_SM,Keren2011_EurBJ,TaloniKardashSalmanEtAl2015_PRL}.  
%
%Water permeation across the cell membrane due to an osmotic shock has been well studied \cite{AlbertsMolecularBiology}. 
Recent findings confim that both osmotic and mechanical stresses contribute to significant water permeation for cells to migrate under strong confinement \cite{JiangSun2013_BJ,StrokaJiangChenEtAl2014_Cell,LiMoriSun2015_PRL,yao-mor2017}, leading to a dynamic surface-to-volume ratio of a migrating cell. 
While stress-induced release of messengers in cells has been well studied \cite{Wan2008_PNAS, ForsythWan2011_PNAS,Russell-PuleriPazAdams2016_AJHCP,ZhangShenHoganBarakatMisbah2018_BJ,GordonShimmelFrye2020_FP},
%osmotic effects on cellular dynamics have been well studied \cite{AlbertsMolecularBiology}, 
mechanically induced water permeation is often associated with membrane poration or rupture under extreme stresses \cite{HarmanBertrandJoos2017_CJP,RazizadehNikfarPaulLiu2020_BJ}. In this work we use modeling and direct numerical simulations to show that, 
over the appropriate time scales or under strong confinement,
the intrinsic membrane permeability to water can give rise to significantly distinct hydrodynamics of a semipermeable vesicle without poration or rupture in the lipid bilayer membrane. 
%
%
%One of the most direct ways to induce water permeation across cell membranes is to impose an osmotic shock }.
%
%A gradient in the osmotic stress across the cellular membrane is the most direct way to induce water permeation, yet 
%%
%%In response to external stimuli, cells move, deform, and release signaling molecules, such as ATP in red blood cells \cite{ZhangShenHoganBarakatMisbah2018_BJ} and cyclooxygenase-2 in endothelial cells \cite{Russell-PuleriPazAdams2016_AJHCP}. 
%%
%%Water permeation is often neglected in such release processes, yet 
%%
%recent findings suggest that both osmotic and mechanical stresses can induce water permeation that is significant enough for tumor cells to migrate under strong confinement \cite{StrokaJiangChenEtAl2014_Cell,LiMoriSun2015_PRL,yao-mor2017}, leading to a dynamic surface-to-volume ratio over time. 
%
%
%To elucidate the effects of mechanically induced water permeation (which is a great challenge to quantify in experiments), in this work we use accurate numerical simulations to study the hydrodynamics of a semi-permeable vesicle in various configurations.

%Cells also send out chemicals in response to 
%
% Erythrocytes are known to  participate in efficient body water homeostasis by regulating its exchange of water  \cite{YangMaVerkman2001_JBC,SugieIntaglietta2018_AmJPhysiolHCP}.  Cells achieve motility by simultaneously regulating their cell volume and cell membrane tension through trafficking of aquaporin proteins \cite{SaadounPapadopoulosWatanabeEtAl2005_JCS,Verkman2008_JMM,BerthaudEtAl2016_SM} and flattening of caveolae \cite{Keren2011_EurBJ}. 
%Cells also control water transport  as they activate shape fluctuations in the membrane for the bleb formation \cite{TaloniKardashSalmanEtAl2015_PRL}.
%
%
%%\dropcap{W}ater balance is an essential process in the adaptation of a living cell to its environment \cite{AlbertsMolecularBiology}: Erythrocytes are known to  participate in efficient body water homeostasis by regulating its exchange of water  \cite{YangMaVerkman2001_JBC,SugieIntaglietta2018_AmJPhysiolHCP}.  Cells achieve motility by simultaneously regulating their cell volume and cell membrane tension through trafficking of aquaporin proteins \cite{SaadounPapadopoulosWatanabeEtAl2005_JCS,Verkman2008_JMM,BerthaudEtAl2016_SM} and flattening of caveolae \cite{Keren2011_EurBJ}. 
%%Cells also control water transport  as they activate shape fluctuations in the membrane for the bleb formation \cite{TaloniKardashSalmanEtAl2015_PRL}.

The lipid bilayer membrane is permeable to water and water-soluble molecules \cite{Dick1964_JTB,FettiplaceHaydon1980_PhysRev,DeamerBramhall1986_ChemPhysLipids,Grafmueller2019_ABLS}. 
%
%In the presence of an osmotic gradient across the membrane (such as a contrast in the concentration of an impermeable solute), water flows against the osmotic gradient  to neutralize the osmolarity. 
%
Without aquaporins or other membrane channels that promotes water permeability, the intrinsic water permeability is known to depend on the temperature \cite{OlbrichRawiczNeedhamEtAl2000_BJ}, lipid composition \cite{OlbrichRawiczNeedhamEtAl2000_BJ}, and the hydration of the membrane \cite{MarrinkBerendsen1994_JPhysChem}. In the red blood cell membrane, aquaporin-1 channels give rise to a much higher water permeability of $\sim 1.8\times 10^{-2}$ cm/s in erythrocytes \cite{YangMaVerkman2001_JBC} while the intrinsic water permeability of a lipid bilayer membrane is in the range of $10^{-4}-10^{-3}$ cm/s \cite{ThompsonHuang1966_ANYAS,FettiplaceHaydon1980_PhysRev,Grafmueller2019_ABLS,Dimova2020_GVB,BhatiaRobinsonDimova2020_SoftMatt}. 
%
%The swelling/shrinking of a cell is inevitably coupled to changes in membrane tension and deformation. Under a strong confinement (such as going through a microcapillary or submicron slit),  the transmembrane water flow may have comparable contributions from both osmotic and mechanical stresses \cite{StrokaJiangChenEtAl2014_Cell,LiMoriSun2015_PRL,yao-mor2017}.  
Molecular dynamic (MD) 
%and coarse-grained (CG) MD 
simulations show that a stretched RBC membrane porates when the membrane tension reaches the order of $2-4$ mN/m (\cite{RazizadehNikfarPaulLiu2020_BJ} and references therein), giving rise to enhanced membrane permeability to both water and macromolecules. Consequently, as long as the lipid bilayer membrane remains intact (no poration or rupture), RBCs and vesicles are assumed to be of the same volume throughout experiments of durations no longer than tens of minutes. 
%
%membrane poration is inevitable when membrane tension is of the order of $2-4$ mN/m (\cite{RazizadehNikfarPaulLiu2020_BJ} and references therein). 
%%The CGMD results further show that a stretched RBC membrane may have pores that might contribute to transmembrane transport of both water and ATP molecules \cite{RazizadehNikfarPaulLiu2020_BJ}. 
%Thus an important outstanding question in stress-induced ATP release of highly deformed RBCs is if there is any transmembrane flow when RBCs release ATP as they experience large
%mechanical load due to confinement or large shear stress?  To address such question, the first step we take is to focus on the permeable water flux induced by mechanical stress.
%%
%%RBCs are known to release ATP under large deformation and microfluidic experiments have been used to study the correlation between RBC deformation and ATP release \cite{Wan2008_PNAS,ForsythWan2011_PNAS,ZhangShenHoganBarakatMisbah2018_BJ}.
%%
%%Why is it interesting to study the dynamics of a semi-permeable membrane induced by mechanical stress/confinement?
%%
%%The osmolarity-induced water flux has been well studied, while mechanical pressure may also contribute to the permeable water flow \cite{yao-mor2017}.
%%The response of a cell to an osmotic gradient is inevitably both mechanical (tension redistribution and shape change) and chemical (water flux across the membrane).  
Under what condition is this a good assumption? Do vesicles conserve volume over hours under flow? What is the dynamic consequence when water influx/efflux affects the membrane hydrodynamics?
To answer these questions, we investigate the balance between hydrodynamics, tension and elasticity of a semipermeable membrane under various flowing conditions and confinement. 
%Specifically our numerical simulations yield the quantification of the time scale and length scale over which water permeation becomes important.

% In this work we use numerical simulations to examine the hydrodynamics of a semipermeable vesicle in two dimensions. 
In microfluidic experiments, it is challenging to keep a freely suspended RBC or vesicle in a steady flow for more than a few hours. It is also difficult to quantify the vesicle area/volume ratio accurately. Using numerical simulations we can quantify  the effects of water permeability on the hydrodynamics of a vesicle under mechanical stresses. 
Vesicles (self-enclosed lipid bilayer membranes) have been used as a model system to study reshaping, remodeling and scission of cell membranes due to osmotic stress \cite{OgleckaEtAl2014_eLife,CamposSaric2020_bioRxiv,Dimova2020_GVB,BhatiaChrist2020_SoftMatt}. Under an osmotic shock, a vesicle may undergo oscillatory variation in its enclosed volume \cite{ChabanonHoLiedberg2017_BJ}. When trapped in a microfluidic channel, an osmotic shock induces 
 vesicle shape changes, tank-treading and tumbling motions \cite{BhatiaRobinsonDimova2020_SoftMatt,BhatiaChrist2020_SoftMatt}, an illustration that the permeable water flow is inevitably coupled with the mechanical pressure jump across the membrane\cite{yao-mor2017}, which could become comparable to osmotic stress when the membrane deformation is large \cite{LiMoriSun2015_PRL,yao-mor2017}.  In the absence of an osmotic gradient, we first show that a freely suspended semipermeable vesicle behaves differently from an impermeable vesicle over a long time. Next we show that, at normal physiological conditions, the permeable water flow can be amplified by extreme confinement and results in observable change in both vesicle hydrodynamics and reduced area.  These results will shed light on the dynamic aging process of living cells that constantly go through flows and constrictions over a period of days.

%
%\paragraph{Literature to maybe cite}
%\cite{ali-ovr-ran2020} show how beading can occur on nanotubes. This
%happens in regions where proteins congregate and this can lead to
%semi-permeability?
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[htp]
  \centering
  \includegraphics[width=0.45\textwidth]{figures/schematic.pdf}
  \caption{\label{fig:sketch} Schematics showing the effects of
  semipermeability on vesicle hydrodynamics in various configurations.
  The slipper in (a) is the steady-state shape of a semi-permeable
  vesicle in a Poiseuille flow.}
\end{figure}

\section*{Formulation}
We consider a two-dimensional semi-permeable vesicle (permeable only to water) suspended in a viscous fluid, see figure~\ref{fig:sketch}.
To focus on the effect of water flow on vesicle hydrodynamics in various conditions, we assume the fluid is the same inside and outside the vesicle membrane.
%($\gamma$) with an outward normal $\nn$.
On the membrane the permeable water flux gives rise to a difference between the fluid velocity ${\bf u}({\bf x}\in \gamma)$
and the membrane velocity $\dot{\bf x}$ \cite{yao-mor2017}:
\begin{align}
  \label{eqn:vesicleFlux}
  \uu - \dot{\xx} = - k_w (R T \triangle c + \ff_\mathrm{mem} \cdot \nn) \nn, \qquad
  \xx \in \gamma,
\end{align}
where $\gamma$ denotes the membrane boundary with an outward normal $\nn$, $k_w$ is the water permeability, $R$ is the ideal gas constant, $T$ is the temperature ($^{\circ}$K), $\triangle c$ is contrast in a solute concentration (mol/volume),
and $\ff_\mathrm{mem}$ is the membrane stress that consists of both bending $ \ff_\mathrm{ben}= -k_b \xx_{ssss}$ and tension $\ff_\mathrm{ten}=(\sigma \xx_s)_s$ ($s$
is the arclength along the membrane, see figure~\ref{fig:sketch}. Using $k_w\sim 10^{-12}$ m$^2$s/kg (typical of mammalian cell membranes \cite{LiMoriSun2015_PRL}) and an osmolarity $\triangle c = 0.1$ mol/L (consistent with the osmotic forces in vesicle aspiration experiments \cite{OlbrichRawiczNeedhamEtAl2000_BJ}), the corresponding water flow is of the order $v_\mathrm{o}=0.1 \mu$m/s. Assuming that the membrane bending stiffness $k_b = 10^{-19}$J, a membrane tension $10^{-3}$N/m (one tenth of the lysis tension) and a characteristic length scale of $10^{-6}$m, the tension contribution ($v_\mathrm{t}$) and elastic stress contribution 
($v_\mathrm{e}$) to the permeable water flow can be scaled to $v_\mathrm{o}$ as $v_\mathrm{o}:v_\mathrm{t}:v_\mathrm{e}=1:10^{-2}:10^{-6}$.  Furthermore, we note that in the absence of an osmotic stress, the water permeable flow can be significant when the membrane tension is large due to the external stress such as the hydrodynamics stress due to an external flow and/or due to confinement.
In the following we will explore both.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section*{Effects of membrane water permeability}
We consider a single semi-permeable vesicle in familiar configurations to highlight the effects of membrane water permeability.
In free space we put a semi-permeable vesicle in a quiescent flow, a planar shear flow and a Poiseuille flow.
We also consider a semi-permeable vesicle in confinement: a long channel and a contracting channel.
For each example, we fix the vesicle length and examine the enclosed area as a dynamical consequence of water flow, vesicle shape deformation and membrane tension distribution.

In experiments vesicle under shear flows are measured for a minute (chapter 19 in \cite{Dimova2020_GVB}), a duration over which vesicle can be assumed impermeable \cite{AbkarianViallat2005_BJ}. Our simulations of a semi-permeable vesicle
in these flows show that water permeability may lead to different vesicle hydrodynamics over a time scale of 10 to 100 minutes. 
%{\bf YNY: is this correct? The shear flow results seem to suggest that the permeability may lead to change in area in less than a second???}
For vesicles in a narrow micro-fluidic channel, the typical time scale is of $ms$ \cite{abk-fai-sto2006}. Our simulations show that strong confinement can lead to amplification of water permeability at such short time scale.
%
%all vesicles have a length
%$L$, but the area is not conserved because of semi-permeability. We vary
%the initial vesicle reduced area, the permeability rate, and the flow
%rate. For all simulations, the bending modulus is $10^{-19}J$.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection*{Quiescent Flow} 
\begin{figure*}[htp]
  \centering
  \ifTikz
  \input{figures/relaxationComposite.tikz}
  \else
  \includegraphics{figures/relaxationComposite.pdf}
  \fi
  \caption{\label{fig:relaxationComposite} (a): Snapshots of four
  semi-permeable vesicles with $\beta = 10^{-3}$ submerged in a
  quiescent flow. The fluid flux along the red regions is into the
  vesicle, and the fluid flux along the blue regions is out of the
  vesicle. In all cases, the vesicle reaches a steady state circular
  shape. (b): The reduced area of each of the simulations. The
  particular snapshots in (a) occur at the marks along the curve.}
\end{figure*}
In a quiescent flow, a semipermeable vesicle in free space can be quantified in terms of its area dynamics and its transition to the equilibrium shape:
Our analysis in Supplementary Material shows that a semi-permeable vesicle
relaxes to a circle at a rate that is proportional to $\beta$, and the
steady state tension is $\Lambda = 2\pi^2/L^2$. Our analysis further shows that the time for a semipermeable
vesicle to deviate from its initial reduced area and to reach its steady
state is proportional to $\beta^{-1}$. For this reason, our numerical
results below are reported in terms of the scaled time $\beta t$.

%This is verified for a
%vesicle with initial reduced area $\nu = 0.65$ and with different
%semi-permeability rates. In Table~\ref{tbl:ellipseRelaxTimes} we report
%the required time, $t_\mathrm{steady}$, for the vesicle's reduced area
%to be within 1.3\% of its steady state reduced area, and the required
%time, $t_\mathrm{inflate}$, for the vesicle's reduced area to exceed
%1.3\% of its initial reduced area. Both times are proportional to
%$\beta^{-1}$, and for this reason, our numerical results are reported in
%terms of the scaled time $\beta t$.
%\begin{table}[htp]
%  \centering
%  \caption{\label{tbl:ellipseRelaxTimes} The time for a semi-permeable
%  vesicle to its steady state ($t_\mathrm{steady}$) and to begin
%  inflating ($t_\mathrm{inflate}$).}
%  \begin{tabular}{C{0.8cm}C{2cm}C{0.8cm}C{2cm}C{0.8cm}}
%    $\beta$ & $t_\mathrm{steady}$  & ratio 
%            & $t_\mathrm{inflate}$ & ratio \\
%    \midrule
%    $10^{0}$  & $6.45 \times 10^{0}$ & ---  
%              & $2.17 \times 10^{-2}$ & --- \\ 
%    $10^{-1}$ & $2.00 \times 10^{1}$ & 3.10 & 
%                $1.25 \times 10^{-1}$ & 5.76 \\
%    $10^{-2}$ & $9.61 \times 10^{1}$ & 4.81 & 
%                $1.59 \times 10^{0}$ & 12.72 \\
%    $10^{-3}$ & $8.36 \times 10^{2}$ & 8.71 & 
%                $2.19 \times 10^{1}$ & 13.77 \\
%    $10^{-4}$ & $8.15 \times 10^{3}$ & 9.74 & 
%                $2.14 \times 10^{2}$ & 9.77 \\
%    $10^{-5}$ & $8.03 \times 10^{4}$ & 9.85 & 
%                $2.20 \times 10^{3}$ & 10.14 \\
%    \bottomrule
%  \end{tabular}
%\end{table}

In Figure~\ref{fig:relaxationComposite}, we examine the relaxation
dynamics of a semipermeable vesicle with four initial shapes and reduced areas as labeled.
%$\beta = 10^{-3}$ and a variety of initial configurations. 
The vesicles on the left are color-coded by the water permeation flux: red denotes an influx into the vesicle, and blue denotes an efflux out of the vesicle.
%
First we observe that the initial vesicle shape determines the direction of water flux.
For the vesicle with an initial reduced area ($\alpha(0)=0.12$ in the bottom row) the vesicle deflates (blue regions) first due to the high interior capillary pressure that corresponds to the high membrane curvatures.
For the other three cases efflux is found at some locations of high curvature, while mostly water permeates into the vesicle and significant change in the reduced area is found at $t\sim 10$ minutes.  Over time (about ten hours) all vesicles inflate and as expected, each vesicle eventually
reaches a circular shape with reduced area $\alpha=1$. 
%%These regions are denoted in blue and are only present at early times. 
%Finally, the vesicles's reduced area do not steady state tension is
%$\Lambda = 2\pi^2/L^2$. 
%Therefore, the times for the vesicle to deviate from its initial reduced area and to reach its
%steady state are proportional to $\beta^{-1}$. 
%
%Each of the four cases in Figure~\ref{fig:relaxationComposite} eventually
%reaches a circular shape with reduced area $\alpha=1$. Next, we observe
%that even though the vesicle is inflating to a circle, 
%%meaning that there is a net flux into the vesicle, 
%there are regions where the flux
%is out of the vesicle. These regions are denoted in blue and are only
%present at early times. 
%
%Finally, the vesicles's reduced area do not
%change by much at early times (up to ten minutes), and

Between $t=10$ seconds and $t=10$ minutes, the 
intermediate shapes of semi-permeable
vesicles resemble the shape of an impermeable vesicle. This can be
explained by considering the two terms that contribute to the vesicle
velocity: (i) the permeability velocity, $\beta(\ff \cdot \nn) \nn$, and
(ii) the force balance velocity, $\SS[\ff]$. Since $\beta = 10^{-6}$,
the permeability velocity is much smaller than the force balance
velocity at early times, and the semi-permeable vesicle dynamics
resembles those of an impermeable vesicle. As the force balance velocity
decreases, the permeability velocity dominates, and the vesicle begins
to inflate to a circle with no vanishing force and velocity. 
%{\bf YNY:We should put a tick on the time when this transition occurs for each of the vesicles}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection*{Planar Shear Flow}
\begin{figure}[htp]
  \centering
  \ifTikz
  \input{figures/shearComposite.tikz}
  \else
  \includegraphics{figures/shearComposite.pdf}
  \fi
  \caption{\label{fig:shearComposite} (a) The reduced area of a
  semi-permeable vesicle with $\beta = 10^{-3}$ initialized with four
  different reduced areas in a shear flow with two different flow rates.
  (b) The steady state shape of a semi-permeable vesicle with varying
  flow rates and semi-permeability rates. The circled vesicles
  correspond to the simulations in part (a). {\bf YNY: should this be two-column wide?}}
\end{figure}

We consider a semi-permeable vesicle in a planar shear flow
$\uu_{\infty}(\xx) = \dot{\gamma} (y,0)$ with different permeability  $\beta$ and shear rates $\dot{\gamma}$. 
When $\beta=0$ (impermeable case), a vesicle's reduced area
and viscosity contrast determine whether it tank treads or
tumbles~\cite{fin-lam-sei-gom2008, kra-win-sei-lip1996}. 
Focusing on semi-permeable vesicles with no viscosity contrast, 
we find that a semi-permeable vesicle tilts to an
inclination angle and undergoes tank treading dynamics, similar to the case of  an impermeable vesicle.

In Figure~\ref{fig:shearComposite}(a), using four initial reduced areas
and two flow rates, we plot the reduced area of a semi-permeable vesicle
with $\beta = 10^{-3}$. As observed in a quiescent flow, the final
vesicle shape is independent of the initial reduced area. However, it
does depend on the flow rate.  We also observe that the reduced area remains constant until $t\sim 10$ minutes, when
the vesicle starts to tank-tread and the normal component of the force-balance velocity becomes significantly smaller than the permeable velocity, leading to a changing reduced area.

For a variety of semi-permeability rates and flow rates, a semipermeable
vesicle reaches a steady equilibrium as in
Figure~\ref{fig:shearComposite}(a).  Since the final shape is
independent of the initial reduced area,  we initialize the vesicle with
reduced area $\alpha = 0.65$ for all simulations in
Figure~\ref{fig:shearComposite}(b), where we summarize the steady state
vesicle shape and its reduced area. We see that the flow rate
significantly affects the final vesicle shape, but the semi-permeability
rate has a minor effect. Instead, as we saw in the quiescent analysis,
$\beta$ sets a time scale since the time required for a vesicle to reach
its steady state configuration is proportional to $\beta^{-1}$.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection*{Poiseuille Flow}
\begin{figure*}[htp]
  \ifTikz
  \centering
  \input{figures/parabolicComposite.tikz}
  \else
  \includegraphics{figures/parabolicComposite.pdf}
  \fi
  \caption{\label{fig:parabolicComposite} (a) The steady state shape of
  a semi-permeable vesicle submerged in a Poiseuille flow with varying
  initial reduced areas and flow rates. The fluid flux along the red
  regions is into the vesicle, and the fluid flux along the blue regions
  is out of the vesicle.  (b) The steady state inclination angle for
  four different flow velocities. The steady state shapes are
  superimposed. (c) The reduced area of the slipper formed with the flow
  rate of 800 $\mu m/s$. (d) The vesicle shape shown with the background
  imposed flow. The channel width is 12.5 times larger than the vesicle
  radius. The corresponding reduced area of each vesicle is indicated by
  the marks in plot (c).}
\end{figure*}

We consider a semi-permeable vesicle in the Poiseuille flow $\uu =
U(1-(y/W)^2,0)$, where $U$ is the maximum velocity at $y=0$. The non-linear effects of the Poiseuille flow
on the migration of an impermeable vesicle have been well-studied \cite{kao-bir-mis2009}: At a given value of the maximum velocity $U$,
there exists a critical reduced volume above which the equilibrium vesicle shape is symmetric (parachute or bullet shapes) and the vesicle stays at $y=0$.
Below this critical reduced volume the equilibrium vesicle moves off the center and is asymmetric (tank-treading slipper shape).
We validated our numerical codes by comparing against results in Kaoui {\it et al.} \cite{kao-bir-mis2009}, see Supplementary Material.

%
%
%We examine the effect of the non-linear Poiseuille flow on the migration of
%semi-permeable vesicle. In this experiment, the vesicle is initialized
%above the center line $y=0$, and it migrates towards the center
%line~\cite{dan-vla-mis2009}. 
%
%For an impermeable vesicle, depending on
%the vesicles reduced area and the flow rate, the equilibrium vesicle shape
%is either an axisymmetric parachute or bullet, or an asymmetric
%tank-treading slipper~\cite{kao-bir-mis2009}. 
%We validated our numerical codes by comparing against results in Kaoui {\it et al.} \cite{kao-bir-mis2009}.
%One of the most salient features of an impermeable vesicle in the Poiseuille flow is that, 
%%
%%To validate our methods,
%%in Figure~\ref{fig:parabolicComposite}(a), we construct a phase diagram
%%for the steady state shape of an impermeable vesicle
%%vesicle~\cite{kao-bir-mis2009} (see Figure 2). While the vesicle shapes
%%qualitatively agree, Kaoui et al.~report that the vesicles with
%%$\nu=0.7$ and flow rates $U = 1200 \mu m/s$ and $U = 1600 \mu m/s$ are
%%axisymmetric. In our simulations these vesicles are slightly asymmetric,
%%are undergoing a tank treading motion, and we therefore report them as
%%asymmetric slippers.
%
%Having validated our method, we investigate the dynamics of a semi-permeable vesicle in a Poiseuille flow. 
%
%Using several initial
%$U(1-(y/W)^2,0)$, where $U$ is the maximum imposed velocity. We examine
%the effect of the non-linear Poiseuille flow on the migration of
%semi-permeable vesicle. 
%

At the beginning of the simulations a semipermeable vesicle is initialized
above the center line $y=0$, and it migrates towards the center
line as in the impermeable case~\cite{dan-vla-mis2009}. However, for an impermeable vesicle, depending on
the initial reduced area and the maximum flow velocity $U$, the equilibrium vesicle shape
is either an axisymmetric parachute or bullet, or an asymmetric
tank-treading slipper~\cite{kao-bir-mis2009}. 
%
%To validate our methods,
%in Figure~\ref{fig:parabolicComposite}(a), we construct a phase diagram
%for the steady state shape of an impermeable vesicle
%vesicle~\cite{kao-bir-mis2009} (see Figure 2). While the vesicle shapes
%qualitatively agree, Kaoui et al.~report that the vesicles with
%$\alpha=0.7$ and flow rates $U = 1200 \mu m/s$ and $U = 1600 \mu m/s$ are
%axisymmetric. In our simulations these vesicles are slightly asymmetric,
%are undergoing a tank treading motion, and we therefore report them as
%asymmetric slippers.
%
Using several initial reduced areas and flow rates, we recreate the impermeable vesicle phase
diagram for vesicles with a permeability rate of $\beta = 10^{-3}$. The equilibrium vesicle shape for different $U$ and initial reduced area is summarized in
Figure~\ref{fig:parabolicComposite}(a). An example of a semi-permeable vesicle with $\alpha(0) = 0.9$ and $U=800$ $\mu m/s$ is
shown in Figure~\ref{fig:parabolicComposite}(c) and (d). 
%

As observed for the planar shear flow,
the steady state reduced area is independent of the initial reduced
area, and larger flow rates result in smaller steady state reduced
areas. We report the reduced area of the equilibrium vesicle shape in
Figure~\ref{fig:parabolicComposite}(a). At the smallest flow rate of $U= 200 \mu m/s$, the steady state reduced area is large, and the result
is an axisymmetric bullet. At the larger flow rates, the steady state
reduced area is smaller, and the result is an asymmetric tank-treading
slipper. We observe that the membrane permeability to water drastically alters the vesicle shape and position relative to the flow at equilibrium: At $\beta=10^{-3}$ the equilibrium vesicle depends only on the maximum velocity $U$ of the nonlinear shear flow. The higher $U$ the smaller the equilibrium reduced area and the further away vesicle is relative to the flow center at $y=0$.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection*{Semi-Permeable Vesicle in a Narrow Channel (Stenosis)}
Cells experience stretching and release messengers when they pass through capillary vessels \cite{Wan2008_PNAS,ForsythWan2011_PNAS,Pak2015_PNAS}.
Many cells exhibit extraordinary flexibility as they go through narrow channels \cite{AuStoreyMoore2016_PNAS}, and adjustment of area to volume ratio may be a key factor for a successful passage.
Here we quantify the effect of water permeability on a vesicle going through a stenosis (a long microfluidic channel with a width smaller than the cell radius).
Based on the findings in 
%We consider the effect of stenosis on a semi-permeable vesicle. Inspired
%by an experiment that measures the effects of cells on the dynamical
%pressure-drop variations along a micrometer-sized
\cite{abk-fai-sto2006}, we first examine the effect of
semi-permeability on the pressure fluctuation across a stenosed
geometry. In this experiment, a pressure drop is used to squeeze a
red blood cell (RBC) through a constriction with diameter $5\mu$m and
length $45\mu$m. We set up a similar computational geometry with a
Poiseuille flow imposed at the inlet and outlet
(Figure~\ref{fig:stenosisComposite}(a)). The flow velocity $U_{max} = 1000 \mu$m/s is chosen so that
the residency time of the vesicle  in the stenosis channel $t_R \approx 0.5$ s agrees with the
experimental results \cite{abk-fai-sto2006}. When the RBC enters the channel, an increase in the pressure drop is required to maintain a constant flow
rate, and the difference in the pressure drop is called the excess
pressure.

We consider both an impermeable and semi-permeable vesicle with initial
reduced area $\alpha = 0.65$ passing through a long narrow channel and
compute the excess pressure as a function of vesicle position in the
channel, see Figure~\ref{fig:stenosisComposite}(c). A 
Poiseuille flow with a maximum value of $1000 \mu$m/s is maintained
at the inlet and outlet. The impermeable case agrees with experimental
results in Figure 3 of~\cite{abk-fai-sto2006}. Our simulations show that
less excess pressure is needed to drive a semi-permeable vesicle through the closely-fit channel.
Once in the channel, the vesicle takes a bullet shape with maximum tension in the front and minimum tension in the back (see Figure~\ref{fig:stenosisComposite}(a)),
consistent with earlier results \cite{Pak2015_PNAS,HarmanBertrandJoos2017_CJP}.

Figure~\ref{fig:stenosisComposite}(d)) shows
that the vesicle
deflates from the beginning due to the high flow velocity, consistent
with the results for a vesicle in a Poiseuille flow in open space. As
the vesicle enters the stenosis channel, it continues to
deflate in the front even though we observe influx (red segment) at
various rear locations in the membrane. Overall, the vesicle loses about 6.1\% of its initial total water at a constant rate as shown in
Figure~\ref{fig:stenosisComposite}(d).  Thus we expect that a larger amount of deflation is possible
if the vesicle is inside a longer stenosis: we consider a vesicle passing
through a channel of triple the length in
Figure~\ref{fig:stenosisComposite}(b) with $U_{max}=8.4 \mu$m/s, and we observe that vesicle stays inside the stenosis  for a residence time $t_R=16.2$ seconds and the reduced
area decreases by 13.2\% (Figure~\ref{fig:stenosisComposite}(e)).
%\begin{itemize}
%  \item In plot (a), flow rate is $1000 \mu m/s$ and residency time is
%    $t_R = 45.6 \times 10^{-2}$ s.
%  \item In plot (b), flow rate is $8.4 \mu m/s$ and residency time is
%    $t_R =  16.2$ s.
%\end{itemize}

\begin{figure*}[htp]
  \ifTikz
  \input{figures/stenosisComposite.tikz}
  \else
  \includegraphics{figures/stenosisComposite.pdf}
  \fi
  \caption{\label{fig:stenosisComposite} (a) A semi-permeable vesicle
  passing through a constricted geometry similar to the experimental
  device in~\cite{abk-fai-sto2006}. (b) A semi-permeable vesicle passing
  through a constricted geometry that is over five times longer. (c) The
  excess pressure of an impermeable and semi-permeable vesicle in the
  short geometry in (a). (d) The reduced area of the semi-permeable
  vesicles in the geometries in (a) and (b) as a function of the
  vesicles' center of mass. The $x$-axis is scaled by $1/L$ where $L$ is
  half the length of the constriction. (e) The reduced area of the
  semi-permeable vesicles in the geometries in (a) and (b) as a function
  of time. The time axis is scaled by $1/t_R$ where $t_R$ is the
  residency time. In parts (c) and (d), the dashed lines correspond to
  the locations of the inlet and outlet. In part (e), the dashed lines
  correspond to the times that the vesicle enters and exits the
  constriction.}
\end{figure*}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection*{Semi-Permeable Vesicle in a Contracting Geometry} 
Squeezing a cell or an elastic vesicle through a slit in microfluidics has gained popularity for its application in stress-induced release of macromolecules
\cite{ShareiEtAl2013_PNAS,Pak2015_PNAS,ZhangShenHoganBarakatMisbah2018_BJ,LuoBai2019_PoF}. RBCs go through submicron slits many times during their lifetime, and such process selects healthy, flexible RBCs to survive \cite{wu2015critical,LuPeng2019_PoF}.
Here we investigate the role of membrane permeability to water when a semi-permeable vesicle is squeezed through a contracting microfluidic channel similar to~\cite{wu2015critical},
where the width of the channel gradually decreases from $16 \mu$m to $1.6 \mu$m. The channel then immediately opens
again to $16 \mu$m (Figure~\ref{fig:contractingComposite}). A
Poiseuille flow is imposed at the inlet and outlet with a constant maximum flow
rate of $U_{\max}$. We initialize a semi-permeable vesicle with $\alpha
= 0.65$ that spans $10 \mu$m of the total channel width and investigate
the effects of the membrane permeability to water and the maximum flow velocity on whether a vesicle can successfully pass through the slit. 
We color the vesicle according to the sign of its flux---red for influx and blue for efflux.

In Figures~\ref{fig:contractingComposite}a
and~\ref{fig:contractingComposite}b, we plot snapshots of vesicles with
different permeability $\beta$. 
Figure~\ref{fig:contractingComposite}c
shows the corresponding reduced area as a function of the vesicles' location. 
The reduced area in Figure~\ref{fig:contractingComposite}c and f shows that the vesicle always inflates gradually over seconds
before it reaches the neck, where vesicle quickly deflates within a fraction of a second.
%{\bf YNY: inflates over seconds but deflates over ms?}
However, if the
semi-permeability rate is too low, the vesicle is too large when it
reaches the narrowest part of the channel, and it is unable to pass. In
Figures~\ref{fig:contractingComposite}d
and~\ref{fig:contractingComposite}e, we plot snapshots of vesicles with
different imposed maximum velocities, and
Figure~\ref{fig:contractingComposite}f shows that the reduced area as a
function of the vesicles' location. At low velocities, the vesicle
inflates similar to the unbounded parabolic flow example, and it is
unable to pass through the constrictions. However, higher flow rates
result in additional deflation and the vesicle passes through the
contraction.

To understand the effects of water permeability on cells that repeatedly go through micro-capillary vessels or sub-micron slits, 
we construct a contracting geometry (see Figures~\ref{fig:contractingComposite}g)
to simulate a vesicle passing through the contracting geometry six times.
In Figure~\ref{fig:contractingComposite}h, we show that the vesicle
deflates each time it passes through the constriction followed by a
gradual inflation (over seconds). The total area loss over the history of the vesicle is
almost 50\%. In Figure~\ref{fig:contractingComposite}i, we show the
vesicle velocity as a function of the vesicle's center of mass. As the
vesicle passes through the contraction, it first accelerates until it reaches the neck where it abruptly decelerates
to go through the contraction. Their corresponding variation with respect to time is shown in Figure~\ref{fig:contractingComposite2}.

%
%\begin{itemize}
%  \item Vesicle velocity in plot (a) varies from 5 to 24 microns/s.
%  \item Vesicle velocity in plot (b) varies from 5 to 28 microns/s. Time
%    that vesicle first completely passes through constriction is 4.75s
%  \item Vesicle velocity in plot (c) varies from 0.2 to 2 microns/s.
%  \item Vesicle velocity in plot (d) varies from 5 to 25 microns/s. Time
%    that vesicle first completely passes through constriction is 4.76s
%  \item Vesicle velocity in plot (g) varies from 14 to 95 microns/s (see
%    plot (i)). Time that vesicle first completely passes through
%    constriction is 1.43s
%\end{itemize}

\begin{figure*}[htp]
  \centering
  \ifTikz
  \input{figures/contractingComposite.tikz}
  \else
  \includegraphics{figures/contractingComposite.pdf}
  \fi
  \caption{\label{fig:contractingComposite} A contracting geometry
  similar to the experimental device in~\cite{wu2015critical}. The
  channel width varies from $18\mu$m to $1.8\mu$m, and it repeats every
  $45\mu$m. The fluid flux along the red regions is into the vesicle,
  and the fluid flux along the blue regions is out of the vesicle.
  (a--c) The ability for a semi-permeable vesicle to pass through the
  constriction depends on the semi-permeability rate. (d--f) The ability
  for a semi-permeable vesicle to pass through the constriction depends
  on the flow rate. (g) A semi-permeable vesicle passing through six
  periods of the constricting geometry. (h) The reduced area with the
  marks representing shapes in (g). (i) The vesicle velocity with the
  marks representing shapes in (g).}
\end{figure*}

\begin{figure}[htp]
  \ifTikz
  \input{figures/contractingComposite2.tikz}
  \else
  \includegraphics{figures/contractingComposite2.pdf}
  \fi
  \caption{\label{fig:contractingComposite2} Vesicle reduced area and
  velocity vs.~time.}
\end{figure}

\section*{Discussion}

% \bibliographystyle{plain}
 \bibliography{refs}

\section*{Supplementary Material: Formulation}
We consider a two-dimensional inextensible semi-permeable vesicle,
$\gamma$, suspended in unconfined and confined geometries $\Omega$. The
suspending fluid is assumed to be viscous so that the fluid velocity
$\uu$ and pressure $p$ are governed by the incompressible Stokes
equations
\begin{align}
  \label{eqn:governing}
  -\nabla p + \mu \triangle \uu = 0, \quad
  \nabla \cdot \uu = 0, \qquad \xx \in \Omega \setminus \gamma,
\end{align}
where $\mu$ is the fluid viscosity. In an unconfined geometry, the
condition $\uu(\xx) \rightarrow \uu_\infty(\xx)$ is enforced as $\xx
\rightarrow \infty$, and in a confined geometry, the boundary condition
$\uu(\xx) = \UU(\xx)$ is enforced on $\partial \Omega$, where $\UU(\xx)$
is prescribed. Along the vesicle $\gamma$, mass continuity, a force
balance, and local inextensibility are enforced by requiring
\begin{align}
  \label{eqn:jump}
  \jump{\uu} = 0, \quad
  \jump{T\cdot\nn} = \ff_\mathrm{mem}, \quad
  \xx_s \cdot \uu_s = 0,
\end{align}
where $\nn$ is the outward normal of $\gamma$, $\jump{\cdot}$ is the
jump across the interface, and $T$ is the hydrodynamic stress tensor.
The membrane force $\ff_\mathrm{mem}$ is the sum of a bending force
with bending modulus $k_b$ and a force due to the tension $\Lambda$.
Following~\cite{vee-gue-zor-bir2009}, we modify the bending and tension
forces as $\ff_\mathrm{ben} = -k_b \xx_{ssss}$, and $\ff_\mathrm{ten} =
(\sigma \xx_s)_s$, where $s$ is arclength. The physical tension can be
recovered using the identity $\Lambda = \sigma + 1.5 \kappa^2$, where
$\kappa$ is the membrane curvature.

Semi-permeability is introduced by modifying the vesicle velocity from
the standard no-slip condition $\cdot{\xx} = \uu(\xx)$, where $\xx(s,t)$
parameterizes the vesicle shape $\gamma$ at time $t$ with respect to
arclength. Instead, we consider the case where the flux through the
vesicle is proportional to the mechanical force~\cite{yao-mor2017}. That
is, 
\begin{align}
  \label{eqn:vesicleFlux}
  \uu - \dot{\xx} = - k_w (\ff_\mathrm{mem} \cdot \nn) \nn, \qquad
  \xx \in \gamma,
\end{align}
where $k_w$ is the water permeability.

The governing equations are non-dimensionalized by introducing a length
scale that we take to be $R_0 = 10^{-6}$m. Using a bending stiffness of
$k_b = 10^{-19}$J and fluid viscosity $5 \times 10^{-2}$kg/ms, we define
the time scale $\tau = \mu R_0^3/k_b = 0.5s$, the velocity scale $U =
R_0/\tau = 2\mu$m/s, the pressure scale $P = k_b/R_0^3 = 10^{-1}$Pa, the
tension scale $Q = k_b/R_0 = 10^{-13}$N, and the permeability scale $W =
R_0\mu = 2 \times 10^{-5}$m$^2$s/kg. Another standard dimensionless
variable in vesicle suspensions is the reduced area $\nu = 4\pi A/L^2
\in (0,1]$, where $L$ is the vesicle length and $A$ is its area. The
length is conserved by the local inextensibility condition, but the area
is not because of semi-permeability. Therefore, the reduced area is
dynamic.

Since the fluid satisfies the Stokes equations, its velocity can be
represented as a layer potential. Given the interfacial boundary
conditions~\eqref{eqn:jump}, the fluid velocity is
\begin{align}
  \uu(\xx) = \uu_\infty(\xx) + \SS[\ff_\mathrm{mem}](\xx), \quad
    \xx \in \Omega,
\end{align}
where $\SS$ is the single-layer potential
\begin{align}
  \SS[\ff](\xx) = \frac{1}{4\pi\mu} \int_{\gamma} \left(
    -\mathbf{I} \log\rho + \frac{\rr \otimes \rr}{\rho^2} \right)
    \ff(\yy) ds_{\yy},
\end{align}
with $\rr = \xx - \yy$ and $\rho = \norm{\rr}$. Imposing the flux
condition~\eqref{eqn:vesicleFlux} and the inextensibility condition, the
vesicle velocity satisfies the boundary integral equation
\begin{align}
  \label{eqn:vesVelocity}
  \dot{\xx} &= \uu_\infty(\xx) + \beta (\ff_\mathrm{mem}\cdot\nn)\nn
  + \SS [\ff_\mathrm{mem}](\xx),  \quad
  \xx_s \cdot \dot{\xx}_s = 0,
\end{align}
where $\beta = k_w/W$ is the dimensionless water permeability. In
confined geometries, the far field term $\uu_\infty$ is replaced by a
double-layer potential with a density defined on $\partial\Omega$. The
unknown density function allows us to satisfy the Dirichlet boundary
condition $\uu(\xx) = \UU(\xx)$ for $\xx \in \partial\Omega$.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section*{Supplementary Material: Vesicle Area}
In the absence of an imposed flow, the steady state vesicle shape and a
differential equation for the area of the vesicle can be derived. We
start by following the analysis of Veerapaneni et
al.~\cite{vee-raj-bir-pur2009} by computing the steady state shape of a
semi-permeable vesicle in a quiescent flow. The steady state shape is a
stationary point of the Lagrangian
\begin{align}
  \mathcal{L} = \frac{1}{2}\int_{\gamma} \kappa^2 \, ds +
    \Lambda \left(\int_{\gamma} ds  - L \right),
\end{align}
where the Lagrange multiplier $\Lambda$ is the tension. Note the absence
of the pressure Lagrange multiplier since the vesicle's area fluctuates.
Taking the variation of $\mathcal{L}$ with respect to $\gamma$ and
setting it to zero, the steady state shape satisfies $\kappa_{ss} +
\frac{1}{2}\kappa^3 - \Lambda \kappa = 0$, with initial conditions
$\kappa(0) = \kappa_0$ and $\kappa_s(0) = 0$.  Integrating once, we have
\begin{align}
  \frac{\kappa_s^2}{2} + \frac{\kappa^4}{8} - 
    \frac{\Lambda}{2}\kappa^2 = \frac{\kappa_0^4}{8} - 
    \frac{\Lambda}{2}\kappa_0^2.
  \label{eqn:curveODE}
\end{align}
Assuming $\kappa_s \neq 0$,~\eqref{eqn:curveODE} is separable, and can
be solved analytically in terms of the {\em EllipticF} function using
Mathematica. Unlike in the impermeable case, the solution
of~\eqref{eqn:curveODE} is not periodic unless $\kappa(s) = \kappa_0$.
Therefore, the steady state shape of a semi-permeable vesicle in a
quiescent flow is a circle with radius $L/2\pi$. Therefore, the steady
state curvature and tension are $\Lambda = 2\pi^2/L^2.$ Note that in
contrast to a impermeable vesicle, the tension is positive rather than
negative.

We next compute the transient area of the vesicle. Since the area of the
vesicle is
\begin{align}
  A(t) = \frac{1}{2}\int_{\gamma} (\xx \cdot \nn)\, ds,
  \label{eqn:area}
\end{align}
its derivative includes terms due to the time derivative of $\xx \cdot
\nn$ and a term due to interface stretching~\cite{lai-tse-hua2008}, but
this latter term term is zero because of the local inextensibility
condition. Therefore,
\begin{align}
  \dot{A}(t) =
  \frac{1}{2} \int_{\gamma} (\dot{\xx} \cdot \nn)\, ds  + 
  \frac{1}{2} \int_{\gamma} (\xx \cdot \dot{\nn})\, ds.
\end{align}
Applying integration by parts, the two integrals agree, and we have
%\begin{align}
%  \int_\gamma (\xx \cdot \dot{\nn})\, ds 
%    = \int_\gamma \left(\xx \cdot \dot{\xx}_s^\perp \right)
%    = -\int_\gamma \left(\xx^\perp \cdot \dot{\xx}_s\right)\, ds 
%    = \int_\gamma \left(\xx_s^\perp \cdot \dot{\xx}\right)\, ds 
%    = \int_\gamma (\dot{\xx} \cdot \nn)\, ds.
%\end{align}
\begin{align}
  \dot{A}(t) = \int_\gamma \left(\beta (\ff \cdot \nn)\nn 
    + \SS[\ff]\right) \cdot \nn\, ds 
  = \beta \int_\gamma (\ff \cdot \nn)\, ds,
\end{align}
where we have used the incompressibility of the single-layer potential.
Repeatedly applying integration by parts, we have
\begin{align}
  \dot{A}(t) &= \beta \int_\gamma \left( -\xx_{ssss} + 
    (\sigma \xx_s)_s \right) \cdot \nn \, ds \\
  &= -\beta \int_\gamma \left(-\xx_{sss} + \sigma \xx_s 
    \right) \cdot \kappa \xx_s \, ds \\
  &= -\beta \int_\gamma \left(\xx_{ss} \cdot 
    (\kappa \xx_s)_s + \kappa \sigma \right) \, ds \\
  &= -\beta \int_\gamma \left(\kappa (\xx_{ss} \cdot \xx_{ss}) + 
    \kappa_s (\xx_{ss} \cdot \xx_s) + \kappa \sigma \right) 
    \, ds \\
  &= -\beta \int_\gamma \left(\kappa^3 + \kappa \sigma \right) 
    \, ds = \beta \int_\gamma \left(
    \frac{\kappa^3}{2} - \kappa \Lambda \right) \, ds.
  \label{eqn:areaROC}
\end{align}
Note that the integrand in~\eqref{eqn:areaROC} is zero when the vesicle
reaches its steady state which guarantees that $\dot{A}(t) = 0$.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section*{Supplementary Material: Numerical Methods}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection*{Discretization in Space}
By using a boundary integral equation formulation, the unknowns are the
vesicle's position and tension that we track at $N$ collocation points.
Derivatives are computed with Fourier differentiation, and the
eighth-order quadrature~\cite{alp1999} is applied to the single-layer
potential. For confined geometries, the trapezoid rule is used to
approximate the double-layer potential since its kernel is smooth and
periodic. Nearly-singular integrals are computed with an
interpolation-based quadrature rule~\cite{qua-bir2014}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection*{Discretization in Time}
To eliminate stringent time step restrictions, we modify a time stepping
method that discretizes high-order derivatives
semi-implicitly~\cite{vee-gue-zor-bir2009}. We introduce the notation
\begin{alignat}{3}
  &B[\xx]\ff = -\frac{d^4}{ds^4} \ff,  \qquad
  &&T[\xx]\sigma = (\sigma \xx_s)_s, \\
  &D[\xx]\ff = \xx_s \cdot \ff_s, 
  &&P[\xx]\ff = (\ff \cdot \nn) \nn,
\end{alignat}
for the differential and projection operators. Then, the
governing equations~\eqref{eqn:vesVelocity} are
\begin{align}
  \dot{\xx} &= \uu_\infty(\xx) + \beta P(B\xx + T\sigma)
  + \SS (B\xx + T\sigma), \quad
  D \dot{\xx} = 0, 
\end{align}
We note that the integral and differential operators, which all depend
on the vesicle shape, are non-linear with respect to the vesicle
shape. We construct a time stepping scheme by letting $B^N$ be the
bending operator due to the vesicle configuration at time $t^N$, and use
similar notation for the other operators. Then, a first-order
semi-implicit time stepping method of~\eqref{eqn:vesVelocity} is
\begin{alignat}{3}  
  \frac{\xx^{N+1} - \xx^N}{\Delta t} &= \uu_\infty(\xx^N) 
  &&+ \beta P^N\left(B^N\xx^{N+1} + T^N\sigma^{N+1}\right) \nonumber \\
  & &&+ \SS^N\left(B^N\xx^{N+1} + T^N\sigma^{N+1}\right),  \\
  D^N\xx^{N+1} &= 1.
\end{alignat}
This is a linear system that we solve with matrix-free GMRES.

To achieve second-order accuracy over long time horizons, we modify an
adaptive spectral deferred correction time stepping
method~\cite{qua-bir2016}. Since the area of the vesicle is not
conserved, we only use the length of the vesicle to estimate the error
which determines if a time step is accepted or not and determines the
subsequent time step size.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection*{Numerical Validation}
To validate our methods, we construct a phase diagram of the steady
state shape of an impermeable vesicle in a Poiseuille flow. Depending on
the flow rate and the reduced area, the steady state vesicle becomes
either a symmetric bullet or parachute, or an asymmetric
slipper~\cite{kao-bir-mis2009}. Comparing our results
(Figure~\ref{fig:PoiseuillePhase}) with those in~\cite{kao-bir-mis2009}
(see Figure 2), we see that the results qualitatively agree. We note
that Kaoui et al.~report that the vesicles with $\alpha=0.7$ and flow
rates $U = 1200 \mu m/s$ and $U = 1600 \mu m/s$ are axisymmetric. In our
simulations these vesicles are slightly asymmetric, are undergoing a
tank treading motion, and we therefore report them as asymmetric
slippers.
\begin{figure}[htp]
  \centering
  \ifTikz
  \input{figures/parabolicImpermeable.tikz}
  \else
  \includegraphics{figures/parabolicImpermeable.pdf}
  \fi
  \caption{\label{fig:PoiseuillePhase} Phase diagram of an impermeable
  vesicle with various flow rates and reduced areas. The shapes agree
  qualitatively with Kaoui et al.~\cite{kao-bir-mis2009} (see Figure
  2).}
\end{figure}

\end{document}

% OLD TEXT AND PLOTS FROM THE NUMERICAL EXAMPLES

%We plot the area of a vesicle with six different permeability rates in
%Figure~\ref{fig:ellipseArea}. The area is computed directly from the
%vesicle shape and the time axis is scaled by $\beta$. The plot verifies
%that the time for the vesicle to reach a circular shape is proportional
%to $\beta^{-1}$. We also solve~\eqref{eqn:areaROC} for the $\beta=10$
%simulation and superimpose its solution with dots on
%Figure~\ref{fig:ellipseArea}.

%\begin{figure}[htp]
%  \centering
%%  \ifTikz
%%  \input{figures/ellipseArea.tikz}
%%  \fi
%  \includegraphics[width=0.9\linewidth]{figures/ellipseArea}
%  \caption{\label{fig:ellipseArea} The area of an elliptical
%  semi-permeable vesicle, with permeability rate $\beta$ in a quiescent
%  flow. The time axis is scaled by the permeability rate $\beta$. Since
%  the curves collapse to a single curve, the permeability rate $\beta$
%  sets a relaxation time scale.}
%\end{figure}

%We further analyze the transient dynamics by considering the vesicle
%velocity~\eqref{eqn:vesVelocity} for small $\beta$. Two terms contribute
%to the vesicle velocity: (i) the permeability velocity, $\beta(\ff \cdot
%\nn)\nn$, and (ii) the force balance velocity, $\SS[\ff]$. Since $\beta
%\ll 1$, the permeability velocity is much smaller than the force balance
%velocity at early times, and the semi-permeable vesicle dynamics
%resembles those of an impermeable vesicle. As the force balance velocity
%decreases, the permeability velocity becomes significant, and the
%vesicle begins to inflate to a circle. This results in the membrane
%force and velocity both converging to zero. Letting $\beta = 10^{-3}$,
%the normal component of the force balance velocity (blue) and the
%permeability velocity (red) are plotted at three times in
%Figure~\ref{fig:vesVelocity}. We also plot the velocity (green) of an
%identically initialized impermeable vesicle. At the early time $t=0.5$,
%the permeability velocity is negligible relative to the force balance
%velocity, and the force balance velocity matches with the impermeable
%vesicle velocity. At intermediate time $t=5$, all three velocities are
%comparable in size, and the semi-permeable vesicle begins to inflate.
%Finally, at time $t=50$, the impermeable vesicle is near steady state,
%while the semi-permeable vesicle is still inflating. The right-most
%plot of Figure~\ref{fig:vesVelocity} is the maximum norm of the
%different velocity contributions as a function of time, and the black
%vertical lines correspond to the times reported in the other three
%subplots.
%
%\begin{figure*}[htp]
%  \centering
%%  \scalebox{0.6}{\input{figures/ellipseVelocity1.tikz}}
%%  \scalebox{0.6}{\input{figures/ellipseVelocity2.tikz}}
%%  \scalebox{0.6}{\input{figures/ellipseVelocity3.tikz}}
%%  \scalebox{0.6}{\input{figures/ellipseVelocity4.tikz}}
%%  \input{figures/ellipseVelocityNorms.tikz}
%  \includegraphics[width=0.24\linewidth]{figures/ellipseVelocity1.pdf}
%  \includegraphics[width=0.24\linewidth]{figures/ellipseVelocity2.pdf}
%  \includegraphics[width=0.24\linewidth]{figures/ellipseVelocity3.pdf}
%  \includegraphics[width=0.24\linewidth]{figures/ellipseVelocity4.pdf}
%  \caption{\label{fig:vesVelocity} The normal velocity contributions due
%  to force balance (blue) and permeability (red). The green line is the
%  velocity of an impermeable vesicle with the same initial shape as the
%  semi-permeable vesicle.}
%\end{figure*}

%We next investigate the interplay between the tension, transmembrane
%flux, and curvature of a semi-permeable vesicle. We let $\beta = 1$ and
%initialize the vesicle with regions of both positive and negative
%curvature. The vesicle tension, $\Lambda$, and the flux, $\ff \cdot
%\nn$, are plotted in Figure~\ref{fig:starTensionFlux}. At early times,
%regions of large negative curvature have negative tension and regions
%with large positive curvature have positive tension, and the same result
%holds for an impermeable vesicle. In addition, regions with large
%curvature have a large inflow of fluid (positive flux), while the
%flatter regions have an outflow of fluid. At these early times, the net
%effect is vesicle deflation. As the vesicle shape relaxes, the same
%general tension and flux trends persist, but the net flux is positive
%and the vesicle inflates. At later times, the tension becomes positive
%and the vesicle continues to inflate towards a circle. Finally, the
%tension transitions to a positive constant function, the membrane force
%and flux approach zero, and the vesicle shape reaches its steady
%circular shape.

%\begin{figure*}[htp]
%  \centering
%  \includegraphics[width=0.48\linewidth,trim =2cm 5cm 0cm 5cm, clip=true]{figures/StarTensionTime1.pdf}
%  \includegraphics[width=0.48\linewidth,trim =2cm 5cm 0cm 5cm, clip=true]{figures/StarFluxTime1.pdf}
%
%  \includegraphics[width=0.48\linewidth,trim =2cm 5cm 0cm 5cm, clip=true]{figures/StarTensionTime2.pdf}
%  \includegraphics[width=0.48\linewidth,trim =2cm 5cm 0cm 5cm, clip=true]{figures/StarFluxTime2.pdf}
%
%  \includegraphics[width=0.48\linewidth,trim =2cm 5cm 0cm 5cm, clip=true]{figures/StarTensionTime3.pdf}
%  \includegraphics[width=0.48\linewidth,trim =2cm 5cm 0cm 5cm, clip=true]{figures/StarFluxTime3.pdf}
%
%  \includegraphics[width=0.48\linewidth,trim =2cm 5cm 0cm 5cm, clip=true]{figures/StarTensionTime4.pdf}
%  \includegraphics[width=0.48\linewidth,trim =2cm 5cm 0cm 5cm, clip=true]{figures/StarFluxTime4.pdf}
%
%  \includegraphics[width=0.48\linewidth,trim =2cm 5cm 0cm 5cm, clip=true]{figures/StarTensionTime5.pdf}
%  \includegraphics[width=0.48\linewidth,trim =2cm 5cm 0cm 5cm, clip=true]{figures/StarFluxTime5.pdf}
%
%  \includegraphics[width=0.48\linewidth,trim =2cm 5cm 0cm 5cm, clip=true]{figures/StarTensionTime6.pdf}
%  \includegraphics[width=0.48\linewidth,trim =2cm 5cm 0cm 5cm, clip=true]{figures/StarFluxTime6.pdf}
%  \caption{\label{fig:starTensionFlux} The vesicle tension (left) and
%  flux (right) of the semi-permeable vesicle in a quiescent flow with
%  $\beta=1$.}
%\end{figure*}

%\begin{figure*}[htp]
%  \centering
%  \includegraphics[width=0.16\linewidth,trim =2cm 5cm 0cm 5cm, clip=true]{figures/StarTensionTime1.pdf}
%  \includegraphics[width=0.16\linewidth,trim =2cm 5cm 0cm 5cm, clip=true]{figures/StarTensionTime2.pdf}
%  \includegraphics[width=0.16\linewidth,trim =2cm 5cm 0cm 5cm, clip=true]{figures/StarTensionTime3.pdf}
%  \includegraphics[width=0.16\linewidth,trim =2cm 5cm 0cm 5cm, clip=true]{figures/StarTensionTime4.pdf}
%  \includegraphics[width=0.16\linewidth,trim =2cm 5cm 0cm 5cm,clip=true]{figures/StarTensionTime5.pdf}
%  \includegraphics[width=0.16\linewidth,trim =2cm 5cm 0cm 5cm,clip=true]{figures/StarTensionTime6.pdf}
%
%  \includegraphics[width=0.16\linewidth,trim =2cm 5cm 0cm 5cm, clip=true]{figures/StarFluxTime1.pdf}
%  \includegraphics[width=0.16\linewidth,trim =2cm 5cm 0cm 5cm, clip=true]{figures/StarFluxTime2.pdf}
%  \includegraphics[width=0.16\linewidth,trim =2cm 5cm 0cm 5cm, clip=true]{figures/StarFluxTime3.pdf}
%  \includegraphics[width=0.16\linewidth,trim =2cm 5cm 0cm 5cm, clip=true]{figures/StarFluxTime4.pdf}
%  \includegraphics[width=0.16\linewidth,trim =2cm 5cm 0cm 5cm, clip=true]{figures/StarFluxTime5.pdf}
%  \includegraphics[width=0.16\linewidth,trim =2cm 5cm 0cm 5cm, clip=true]{figures/StarFluxTime6.pdf}
%  \caption{\label{fig:starTensionFlux} The vesicle tension (top) and
%  flux (bottom) of the semi-permeable vesicle in a quiescent
%  flow with $\beta=1$.}
%\end{figure*}

%\begin{figure}[htp]
%  \centering
%  \ifTikz
%  \input{figures/shearSS.tikz}
%  \fi
%%  \includegraphics[width=0.9\linewidth]{figures/shearSS.pdf}
%  \caption{\label{fig:shearSS} The reduced area of a semi-permeable
%  vesicle in a shear flow. The final reduced area is nearly independent
%  of the initial reduced area.}
%\end{figure}

%\begin{figure}[htp]
%  \centering
%  %\ifTikz
%  \input{figures/shearPhaseDiagramRA2.tikz}
%  %\fi
%%  \includegraphics[width=\linewidth]{figures/shearPhaseDiagramRA2.pdf}
%  \caption{\label{fig:shearPhaseDiagram} The steady state shape of a
%  semi-permeable vesicle with several different permeability rates
%  $\beta$ and shear rates $\dot{\gamma}$. The final reduced areas are
%  reported for the largest and smallest parameter values. The final
%  reduced area decreases with the flow rate, while the permeability rate
%  has a very small effect on the steady state shape.}
%\end{figure}

%\begin{figure}[htp]
%%  \ifTikz
%%  \input{figures/parabolicOffCenterPhaseImpermeable.tikz}
%%  \fi
%  \includegraphics[width=0.9\linewidth]{figures/parabolicOffCenterPhaseImpermeable.pdf}
%  \caption{\label{fig:parabolicOffCenterImpermeablePhaseDiagram} Phase
%  diagram of an impermeable vesicle with several different flow rates
%  $U$ and reduced areas $\nu$. As shown by Kaoui et
%  al.~\cite{kao-bir-mis2009}, the final vesicle shape is either an
%  axisymmetric parachute or bullet (pink region) or an asymmetric
%  slipper (yellow region).}
%\end{figure}

%\begin{figure}[htp]
%%  \ifTikz
%%  \input{figures/parabolicOffCenterPhaseSemipermeable.tikz}
%%  \fi
%  \includegraphics[width=0.9\linewidth]{figures/parabolicOffCenterPhaseSemipermeable.pdf}
%  \caption{\label{fig:parabolicOffCenterSemipermeablePhaseDiagram} Phase
%  diagram of a semi-permeable vesicle, with $\beta = 10^{-3}$, with
%  several different flow rates $U$ and reduced areas $\nu$. The
%  steady state reduced area for each flow rate, which is reported along
%  the top of the vesicles, is nearly independent of the initial reduced
%  area. Low flow rates result in axisymmetric vesicles with large
%  reduced area, and high flow rates result in asymmetric vesicles with
%  low reduced area. Results at higher flow rates are not eporeted since
%  the vesicle self-intersects.}
%\end{figure}

%\begin{figure}[htp]
%%  \ifTikz
%%  \scalebox{0.5}{\input{figures/stenosisShapeTension01.tikz}}
%%  \scalebox{0.5}{\input{figures/stenosisShapeFlux02.tikz}}
%%  \scalebox{0.5}{\input{figures/stenosisShapeTension03.tikz}}
%%  \scalebox{0.5}{\input{figures/stenosisShapeFlux04.tikz}}
%%  \scalebox{0.5}{\input{figures/stenosisShapeTension05.tikz}}
%%  \scalebox{0.5}{\input{figures/stenosisShapeFlux06.tikz}}
%%  \scalebox{0.5}{\input{figures/stenosisShapeTension07.tikz}}
%%  \scalebox{0.5}{\input{figures/stenosisShapeFlux08.tikz}}
%%  \scalebox{0.5}{\input{figures/stenosisShapeTension09.tikz}}
%%  \scalebox{0.5}{\input{figures/stenosisShapeFlux10.tikz}}
%%  \scalebox{0.5}{\input{figures/stenosisShapeTension11.tikz}}
%%  \scalebox{0.5}{\input{figures/stenosisShapeFlux12.tikz}}
%%  \scalebox{0.5}{\input{figures/stenosisShapeTension02.tikz}}
%%  \scalebox{0.5}{\input{figures/stenosisShapeFlux04.tikz}}
%%  \scalebox{0.5}{\input{figures/stenosisShapeTension06.tikz}}
%%  \scalebox{0.5}{\input{figures/stenosisShapeFlux08.tikz}}
%%  \scalebox{0.5}{\input{figures/stenosisShapeTension10.tikz}}
%%  \scalebox{0.5}{\input{figures/stenosisShapeFlux12.tikz}}
%%  \fi
%  \includegraphics[width=0.45\linewidth]{figures/stenosisShapeTension02.pdf}
%  \includegraphics[width=0.45\linewidth]{figures/stenosisShapeFlux04.pdf}
%  \\
%  \includegraphics[width=0.45\linewidth]{figures/stenosisShapeTension06.pdf}
%  \includegraphics[width=0.45\linewidth]{figures/stenosisShapeFlux08.pdf}
%  \\
%  \includegraphics[width=0.45\linewidth]{figures/stenosisShapeTension10.pdf}
%  \includegraphics[width=0.45\linewidth]{figures/stenosisShapeFlux12.pdf}
%  \caption{\label{fig:stenosisShape} An impermeable (dashed) and
%  semi-permeable (solid) vesicle in a stenosed geometry (black).
%  The vesicles' initial configuration are identical with reduced area
%  $\nu = 0.65$. The color denotes the vesicles' tension (left) and flux
%  (right). The semi-permeable vesicle deflates throughout the
%  constriction (see Figure~\ref{fig:stenosisArea}) and develops higher
%  curvature regions than its impermeable counterpart.}
%\end{figure}
%
%\begin{figure}[htp]
%  \centering
%%  \ifTikz
%%  \scalebox{0.5}{\input{figures/stenosisTension12.tikz}}
%%  \scalebox{0.5}{\input{figures/stenosisFlux12.tikz}}
%%  \scalebox{0.5}{\input{figures/stenosisXvelocity12.tikz}}
%%  \scalebox{0.5}{\input{figures/stenosisYvelocity12.tikz}}
%%  \fi
%  \includegraphics[width=0.45\linewidth]{figures/stenosisTension12.pdf}
%  \includegraphics[width=0.45\linewidth]{figures/stenosisFlux12.pdf} \\
%  \includegraphics[width=0.45\linewidth]{figures/stenosisXvelocity12.pdf}
%  \includegraphics[width=0.45\linewidth]{figures/stenosisYvelocity12.pdf}
%  \caption{\label{fig:stenosisTensionAndFlux} The tension (top left), 
%  flux (top right), $x$-velocity (bottom left), and $y$-velocity (bottom
%  right) of the semi-permeable (red) and impermeable (dashed blue)
%  vesicles in Figure~\ref{fig:stenosisShape} at location $x=50$. The
%  point $\theta=0$ corresponds to the right-most point of the vesicle.}
%\end{figure}
%
%\begin{figure}[htp]
%  \centering
%%  \ifTikz
%%  \input{figures/stenosisArea.tikz}
%%  \fi
%  \includegraphics[width=0.8\linewidth]{figures/stenosisArea}
%  \caption{\label{fig:stenosisArea} The area of a stenosed
%  semi-permeable vesicle, with permeability rate $\beta=10^{-3}$. The
%  $x$-axis is the $x$-coordinate of the center of mass of the vesicle.
%  The vesicle loses fluid at a constant rate throughout the constriction
%  that is between $\pm 60$. The semi-permeable vesicle loses 12.8\% of
%  its fluid.}
%\end{figure}
%
%
%\begin{figure}[htp]
%%  \ifTikz
%%  \input{figures/stenosisBL.tikz}
%%  \fi
%  \includegraphics[width=\linewidth]{figures/stenosisBL.pdf}
%  \caption{\label{fig:stenosisBL} A stenosed semi-impermeable (red) and
%  imperemable (blue) vesicle. The semi-permeable vesicle has 
%  reduced area 0.17, and both vesicles were initialized with reduce area
%  $0.20$. The bottom plot shows the velocity relative to the vesicle
%  inside the lubrication layer.  The vesicle velocity in the square is
%  nearly constant (see Figure~\ref{fig:stenosisRA020Velocity}.}
%\end{figure}
%
%
%\begin{figure}[htp]
%  \centering
%%  \ifTikz
%%  \scalebox{0.5}{\input{figures/stenosisXvelocityRA020.tikz}}
%%  \scalebox{0.5}{\input{figures/stenosisYvelocityRA020.tikz}}
%%  \fi
%  \includegraphics[width=0.45\linewidth]{figures/stenosisXvelocityRA020.pdf}
%  \includegraphics[width=0.45\linewidth]{figures/stenosisYvelocityRA020.pdf}
%  \caption{\label{fig:stenosisRA020Velocity} The $x$ (left) and $y$
%  (right) velocities of the vesicle in Figure~\ref{fig:stenosisBL}. The
%  red lines correspond to the semi-permeable vesicle and the blue dashed
%  lines correspond to the impermeable vesicle.}
%\end{figure}
%
%We consider the effect of stenosis on a semi-permeable vesicle. To allow
%for large amounts of fluid flux through the membrane, we consider
%geometry with a long constricted region. We start by considering a
%vesicle initialized with reduced area $\nu = 0.65$, and we compare the
%vesicle shape when $\beta = 0$ (impermeable) and $\beta = 10^{-3}$
%(permeable). Figure~\ref{fig:stenosisShape} shows the vesicles' shape at
%several locations along the constriction. The colors along the left
%channel represent the tension and those along the right represent the
%flux. In Figure~\ref{fig:stenosisTensionAndFlux}, the tension at
%location $x=20$ and the flux at location $x=30$ are plotted with respect
%to arclength. The value of $\theta=0$ corresponds to the nose of the
%vesicle.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\subsection*{Dimensionless equations}
%The inextensibility condition guarantees that the vesicle's length, $L$,
%is conserved, while its area, $A$, can change due to semi-permeability.
%Therefore, the reduced area $\alpha = 4\pi A/L^2 \in (0,1]$ is not
%conserved. We non-dimensionalize~\eqref{eqn:vesVelocity} by introducing
%the length scale $R_0 = L/2\pi$ and defining the time scale $\tau = \mu
%R_0^3/k_b$, the tension scale $Q = k_b/R_0^2$, and the permeability
%scale $W = R_0/\mu$. The background velocity $\uu_\infty$ is
%non-dimensionalized with $R_0/\tau$.
%
%In each experiment, we initialize an elliptical vesicle with longest
%diameter of 9$\mu$m. We use a length scale of $R_0 = 10^{-6}$m, a
%bending stiffness of $k_b = 10^{-19}$J, and fluid viscosity of $\mu=5
%\times 10^{-2}$kg/ms. Using the relaxation time scale $\tau = \mu
%R_0^3/k_b = 5 \times 10^{-1}$s, the pressure scale is $P = k_b/R_0^3 =
%10^{-1}$Pa. 
% \newpage
% \paragraph{\bf \Huge OLD STUFF}
%
%%%%%%%%%%%%%%%
%\section{{$\Theta-L$} formulation}
%Here, we attempt to apply the formulation described by Sohn et
%al.~\cite{soh-tse-li-voi-low2010} to a semi-permeable vesicle.  In the
%semi-permeable formulation, the vesicle velocity is
%\begin{align}
%  \pderiv{\XX}{t} = \uu(\XX(t)) + \beta (\ff \cdot \nn) \nn.
%\end{align}
%In the {$\Theta-L$} variables, we require a normal velocity $V$, and a
%tangential velocity $T$. Since the inextensibility condition is $T_s +
%\kappa V = 0$, a condition for the vesicle velocity is
%\begin{align}
%  (\uu \cdot \ss)_s = -\kappa(\beta (\ff \cdot \nn) + 
%      \uu \cdot \nn).
%\end{align}
%This equation can be solved following equations (32)--(39)
%in~\cite{soh-tse-li-voi-low2010}, with the only difference being that
%$\tilde{\uu}$ satisfies
%\begin{align}
%  (\tilde{\uu} \cdot \ss)_s + \kappa(\tilde{\uu} \cdot \nn) =
%    -(\vv \cdot \ss)_s - \kappa(\vv \cdot \nn) 
%    - \beta\kappa(\ff \cdot \nn).
%\end{align}
%Compared to equation (40), only the last term is new.
%
%
%%%%%%%%%%%%%%%
%Finally, we write the time stepping method in matrix form as
%\begin{align}
%  A \left(
%    \begin{array}{c}
%      \xx^{N+1} \\ \sigma^{N+1}
%    \end{array}
%  \right) = 
%  \left(
%    \begin{array}{c}
%      \xx^{N} + \Delta t \uu_\infty(\xx^N) \\ \mathbf{1}
%    \end{array}
%  \right),
%\end{align}
%where
%\begin{align}
% A = \left(
%  \begin{array}{cc}
%    I + \Delta t \beta P^N B^N + \Delta t \SS^N B^N & 
%    -\Delta t P^N T^N - \Delta t \SS^N T^N \\
%    D^N & 0
%  \end{array}
%  \right).
%\end{align}
%
%\section{Area of a Vesicle}
%The area inside a single vesicle $\gamma$ parameterized with $\xx$ is
%\begin{align}
%  A(t) = \int_{\gamma} (\xx \cdot \nn)\, ds = 
%    \int_{0}^{2\pi} \left(\xx \cdot \pderiv{\xx}{\theta}^\perp\right)
%    \, d\theta.
%\end{align}
%Taking the time derivative of $A(t)$,
%\begin{align}
%  \dot{A}(t) &= \int_{0}^{2\pi} \left(
%    \dot{\xx} \cdot \pderiv{\xx}{\theta}^{\perp} + 
%    \xx \cdot \pderiv{\dot{\xx}}{\theta}^{\perp}\right) \\
%    \, d\theta.
%  &= \int_{0}^{2\pi} \left(
%    (\dot{\xx} \cdot \nn)\|\xx'(\theta)\| \right) d\theta +
%     \int_{0}^{2\pi}
%    \xx \cdot \frac{d}{dt}\left(\nn |\frac{d\xx}{d\theta}| \right)
%\end{align}
%
%
%
%\begin{align}
%  A(t) = \frac{1}{2}\int_{\gamma} (\xx \cdot \nn)\, ds = 
%    \frac{1}{2}\int_{0}^{2\pi} \left(\xx \cdot \pderiv{\xx}{\theta}^\perp\right)
%    \, d\theta.
%\end{align}
%Taking the time derivative of $A(t)$,
%\begin{align}
%  \dot{A}(t) &= \frac{1}{2}\int_{0}^{2\pi} \left(\dot{\xx} \cdot
%  \pderiv{\xx}{\theta}^{\perp} + 
%    \xx \cdot \pderiv{\dot{\xx}}{\theta}^{\perp}\right) d\theta, \\
%  &= \frac{1}{2}
%  \int_{0}^{2\pi}(\dot{\xx}\cdot\nn)\left\|\pderiv{\xx}{\theta}\right\|d\theta +
%     \frac{1}{2}\int_{0}^{2\pi}
%    \left(\xx \cdot \pderiv{\dot{\xx}}{\theta}^{\perp}\right)d\theta, \\
%  &= \frac{\beta}{2} \int_{\gamma}\left(\ff\cdot\nn\right)ds +
%  \frac{1}{2}\int_{0}^{2\pi}\left(\xx \cdot \pderiv{}{t}\left(\nn
%  \left\|\pderiv{\xx}{\theta}\right\|\right)\right)d\theta, \\
%  &= \frac{\beta}{2} \int_{\gamma}\left(\ff\cdot\nn\right)ds + \frac{1}{2}
%  \int_{\gamma}\left(\xx\cdot\dot{\nn}\right)ds + \frac{1}{2}\int_{0}^{2\pi}
%  \frac{\xx\cdot\nn}{\left\|\pderiv{\xx}{\theta}\right\|}
%  \left(\pderiv{\xx}{\theta}\cdot
%  \pderiv{\dot{\xx}}{\theta}\right)d\theta \\
%  &= \frac{\beta}{2} \int_{\gamma}\left(\ff\cdot\nn\right)ds + \frac{1}{2}
%  \int_{\gamma}\left(\xx\cdot\dot{\nn}\right)ds + \frac{1}{2}\int_{\gamma}
%  (\xx\cdot\nn) (\xx_s \cdot \dot{\xx}_s)\, ds.
%\end{align}
%To satisfy the inextensibilty condition, $\xx_s \cdot \dot{\xx}_s = 0$.
%Therefore the time derivative of $A(t)$ is
%\begin{align}
%     \dot{A}(t) &= \frac{\beta}{2} \int_{\gamma}\left(\ff\cdot\nn\right)ds
%     + \frac{1}{2} \int_{\gamma}\left(\xx\cdot\dot{\nn}\right)ds.
%\end{align}
%which we re-write as
%\begin{align}
%    \dot{A}(t) &= \frac{\beta}{2}
%    \int_{\gamma}\left(\ff\cdot\nn\right)ds-\frac{1}{2}\int_{\gamma}\left(\xx^{\perp}\cdot\dot{\xx}_s\right) ds.
%\end{align}
%Applying integration by parts,
%\begin{align}
%    \dot{A}(t) &= \frac{\beta}{2}
%    \int_{\gamma}\left(\ff\cdot\nn\right)ds+\frac{1}{2}\int_{\gamma}\left(\xx^{\perp}_s\cdot\dot{\xx}\right) ds,
%\end{align}
%which we re-write as
%\begin{align}
%    \dot{A}(t) &= \frac{\beta}{2}
%    \int_{\gamma}\left(\ff\cdot\nn\right)ds+\frac{1}{2}\int_{\gamma}\left(\dot{\xx}\cdot\nn\right),\\
%    \dot{A}(t) &= \beta \int_{\gamma}\left(\ff\cdot\nn\right)ds.
%\end{align}
%
%
%At steady state,
%\begin{align}
%    \dot{A}(t) &= \beta \int_{\gamma}\left(\ff\cdot\nn\right)ds =0.
%\end{align}
%We now replace $\ff$ with its definition and apply integration by parts
%\begin{align}
%   0=\beta\int_{\gamma} 
%    (-k_b\xx_{ssss}+(\sigma\xx_s)_s)\cdot\nn\, ds =
%   \beta\int_{\gamma}\left(k_b\xx_{sss}-\sigma\xx_s\right)\cdot
%     \nn_s\, ds
%\end{align}
%From the Frenet-Serret formulas, we have 
%\begin{align}
%     -\beta\int_{\gamma}
%     \left(k_b\xx_{sss}-\sigma\xx_s\right)\cdot\kappa\xx_s ds &=0.
%\end{align}
%
%Applying integration by parts and Frenet-Serret again
%\begin{align}
%     \beta\int_{\gamma}
%     k_b\kappa\left(\xx_{ss}\cdot\xx_{ss}\right)ds + \beta
%     \int_{\gamma} k_b \kappa_s(\xx_{ss} \cdot \xx_s)ds +\beta\int_{\gamma}\kappa\sigma ds  &= 0,\\
%    \beta\int_{\gamma} k_b\kappa(\kappa\nn \cdot
%     \kappa\nn)+ \beta \int_{\gamma} k_b \kappa_s(\xx_{ss} \cdot
%     \xx_s)ds +\beta\int_{\gamma}\kappa\sigma ds &= 0,
%\end{align}
%which we re-write as
%\begin{align}
%   \beta\int_{\gamma}k_b\kappa^3 ds + \beta \int_{\gamma} k_b
%   \kappa_s(\xx_{ss} \cdot \xx_s)ds +\beta\int_{\gamma} \kappa\sigma ds = 0.
%\end{align}
%Applying Frenet-serret again,
%\begin{align}
%    \beta\int_{\gamma}k_b\kappa^3 ds + \beta \int_{\gamma} k_b
%    \kappa_s(\kappa\nn \cdot \TT)ds +\beta\int_{\gamma} \kappa\sigma ds = 0, \\
%    \beta\int_{\gamma}k_b\kappa^3 ds +\beta\int_{\gamma} \kappa\sigma ds  = 0.
%\end{align}
%
%
%\section{Including a concentration field}
%Based on the work of Yao and Mori~\cite{yao-mor2017}, we introduce a
%chemical concentration field $c$.  The concentration field is governed
%by the advection-diffusion equation
%\begin{align}
%  \pderiv{c}{t} + \nabla \cdot (\uu c) = D\Delta c.
%\end{align}
%If the concentration is diffusion dominated (zero Peclet number), then
%the concentration satisfies
%\begin{align}
%  \Delta c = 0, \quad \xx \in \RR^2.
%\end{align}
%The boundary conditions of the concentration field on the vesicle
%$\gamma$ are
%\begin{align}
%  (\uu c - D\nabla c) \cdot \nn = c \pderiv{\xx}{t} \cdot \nn + 
%    j_c + j_p,
%\end{align}
%where $j_c$ and $j_p$ are the transmembrane chemical flux due to passive
%transport and active transport (pumps), respectively.  In the zero
%Peclet number limit, the absence of an active transport mechanism, and
%the choice for the passive flux, the boundary condition of the
%concentration field is
%\begin{align}
%  -D\pderiv{c}{\nn} = c \pderiv{\xx}{t} + 
%    k_c \jump{c}, \quad \xx \in \gamma.
%\end{align}
%To couple the vesicle dynamics to the concentration, there is a
%difference between the fluid velocity and vesicle velocity
%\begin{align}
%  \uu - \pderiv{\xx}{t} = j_w \nn,
%\end{align}
%where $j_w$ is the water flux through the vesicle.  This is the
%osmophoretic flow.  In~\cite{yao-mor2017}, this flux is
%\begin{align}
%  j_w = -k_w(RT\jump{c} + \ff_\text{mem} \cdot \nn).
%\end{align}
%In our current implementation, we drop the $RT\jump{c}$ term, and
%let $\beta = k_w$.
%
%\section{Preliminary results}
%
%
%\begin{figure}
%	\centering
%	\includegraphics[width=.9\textwidth]{figures/1.jpg}
%	\includegraphics[width=.9\textwidth]{figures/2.jpg}
%	\includegraphics[width=.9\textwidth]{figures/3.jpg}
%	\caption{Reduced area as a function of time for $\beta = [0, 10^{-2},10^{-1}, 1]$ with a shear flow rate $\chi = 0$.}
%%	\label{Shear0}
%\end{figure}
%
%\begin{figure}
%	\centering
%	\includegraphics[width=.9\textwidth]{figures/4.jpg}
%	\includegraphics[width=.9\textwidth]{figures/5.jpg}
%	\includegraphics[width=.9\textwidth]{figures/6.jpg}
%	\caption{Reduced area as a function of time for $\beta = [0, 10^{-2},10^{-1}, 1]$ with a shear flow rate $\chi = 0.5$.}
%%	\label{Shear0}
%\end{figure}
%
%\begin{figure}
%	\centering
%	\includegraphics[width=.9\textwidth]{figures/7.jpg}
%	\includegraphics[width=.9\textwidth]{figures/8.jpg}
%	\includegraphics[width=.9\textwidth]{figures/9.jpg}
%	\caption{Reduced area as a function of time for $\beta = [0, 10^{-2},10^{-1}, 1]$ with a shear flow rate $\chi = 1$.}
%%	\label{Shear0}
%\end{figure}
%
%\begin{figure}
%	\centering
%	\includegraphics[width=.9\textwidth]{figures/10.jpg}
%	\includegraphics[width=.9\textwidth]{figures/11.jpg}
%	\includegraphics[width=.9\textwidth]{figures/12.jpg}
%	\caption{Reduced area as a function of time for $\beta = [0, 10^{-2},10^{-1}, 1]$ with a shear flow rate $\chi = 2$.}
%%	\label{Shear0}
%\end{figure}
% 
% 
% \begin{figure}
% 	\centering
% 	\includegraphics[width=.9\textwidth]{figures/BE1.jpg}
% 	\includegraphics[width=.9\textwidth]{figures/BE2.jpg}
% 	\includegraphics[width=.9\textwidth]{figures/BE3.jpg}
% 	\caption{Bending energy as a function of time for $\beta = [0, 10^{-2},10^{-1}, 1]$ with a shear flow rate $\chi = 0$.}
%% 	\label{Shear0}
% \end{figure}
% 
% \begin{figure}
% 	\centering
% 	\includegraphics[width=.9\textwidth]{figures/BE4.jpg}
% 	\includegraphics[width=.9\textwidth]{figures/BE5.jpg}
% 	\includegraphics[width=.9\textwidth]{figures/BE6.jpg}
% 	\caption{Bending energy as a function of time for $\beta = [0, 10^{-2},10^{-1}, 1]$ with a shear flow rate $\chi = 0.5$.}
%% 	\label{Shear0}
% \end{figure}
% 
% \begin{figure}
% 	\centering
% 	\includegraphics[width=.9\textwidth]{figures/BE7.jpg}
% 	\includegraphics[width=.9\textwidth]{figures/BE8.jpg}
% 	\includegraphics[width=.9\textwidth]{figures/BE9.jpg}
% 	\caption{Bending energy as a function of time for $\beta = [0, 10^{-2},10^{-1}, 1]$ with a shear flow rate $\chi = 1$.}
%% 	\label{Shear0}
% \end{figure}
% 
% \begin{figure}
% 	\centering
% 	\includegraphics[width=.9\textwidth]{figures/BE10.jpg}
% 	\includegraphics[width=.9\textwidth]{figures/BE11.jpg}
% 	\includegraphics[width=.9\textwidth]{figures/BE12.jpg}
% 	\caption{Bending energy as a function of time for $\beta = [0, 10^{-2},10^{-1}, 1]$ with a shear flow rate $\chi = 2$.}
% 	\label{Shear0}
% \end{figure}
%
%When the semi-permeable vesicle reaches a steady-state shape, we have
%\begin{align}
%  \int_{\gamma} \left(-\frac{\kappa^3}{2} + \kappa \Lambda \right) \, ds
%  = 0.
%\end{align}
%Based on numerical observations, the steady-state shape is a circular
%vesicle with constant tension. Since the tension and curvature are
%constant, we can solve for the curvature in terms of the tension $\kappa
%= \sqrt{2\Lambda}$. Therefore, the steady state shape is a circle of
%radius $r = 1/\sqrt{2\Lambda}$, and therefore the length is
%\begin{align}
%  L = \frac{\sqrt{2}\pi}{\sqrt{\Lambda}},
%\end{align}
%which is constant by the local inextensibility constraint. Finally, we
%can compute the final tension in terms of the length using the formula
%\begin{align}
%  \Lambda = \frac{2\pi^2}{L^2}.
%\end{align}
%Therefore, in contrast to a impermeable vesicle, the final tension is
%positive rather than negative.
%
%We start by simulating a star-shaped vesicle in a quiescent flow with
%semi-permeability rate $\beta=1$ (Figure~\ref{fig:starShape}). To
%perform this simulation, we use an adaptive second-order time stepping
%method~\cite{qua-bir2016}. The times and reduced areas are in the
%titles, and note that the shapes are not eqiuspaced in time.
%\begin{figure}[htp]
%  \begin{minipage}{0.55\textwidth}
%%    \ifTikz
%    \input{figures/starRelaxation1.tikz}
%    \input{figures/starRelaxation2.tikz}
%    \input{figures/starRelaxation3.tikz}
%    \input{figures/starRelaxation4.tikz}
%    \input{figures/starRelaxation5.tikz}
%    \input{figures/starRelaxation6.tikz}
%%    \fi
%  \end{minipage}
%  \hfill
%  \begin{minipage}{0.4\textwidth}
%%  \ifTikz
%  \input{figures/starAreaDeriv.tikz}
%%  \fi
%  \end{minipage}
%%  \caption{\label{fig:starShape} A initially star-shaped semi-permeable
%%  vesicle in a shear quiescent flow with $\beta=1$. Note that the time
%%  steps are not equispaced. The area (red) and its derivative (blue) of
%%  an initially star-shaped vesicle as a function of time. The dots
%%  correspond to the time steps shown in six time steps
%%  show on the left.}
%\end{figure}
%
%\begin{table}[htp]
%  \centering
%  \begin{tabular}{|C{1cm}|*{6}{C{2cm}}|}
%    \hline
%    $\beta$ & $10^{0}$ & $10^{-1}$ & $10^{-2}$ & 
%             $10^{-3}$ & $10^{-4}$ & $10^{-5}$ \\
%    $t_\mathrm{SS}$ & $6.45 \times 10^{0}$ & $2.00 \times 10^{1}$ & 
%                      $9.61 \times 10^{1}$ & $8.36 \times 10^{2}$ & 
%                      $8.15 \times 10^{3}$ & $8.03 \times 10^{4}$ \\
%    ratio & --- & 3.10 & 4.81 & 8.71 & 9.74 & 9.85 \\
%    \hline
%  \end{tabular}
%  \caption{\label{tbl:ellipseRelaxTime} The time, $t_{\mathrm{SS}}$, for
%  a semi-permeable vesicle to reach steady state.}
%\end{table}
%
%\begin{table}[htp]
%  \centering
%  \begin{tabular}{|C{1cm}|*{6}{C{2cm}}|}
%    \hline
%    $\beta$ & $10^{0}$ & $10^{-1}$ & $10^{-2}$ & 
%             $10^{-3}$ & $10^{-4}$ & $10^{-5}$ \\
%    $t_\mathrm{SP}$ & $2.17 \times 10^{-2}$ & $1.25 \times 10^{-1}$ & 
%                      $1.59 \times 10^{0}$ & $2.19 \times 10^{1}$ & 
%                      $2.14 \times 10^{2}$ & $2.20 \times 10^{3}$ \\
%    ratio & --- & 5.76 & 12.72 & 13.77 & 9.77 & 10.14 \\
%    \hline
%  \end{tabular}
%  \caption{\label{tbl:ellipseBLTime} The time, $t_\mathrm{SP}$, for a
%  semi-permeable vesicle to begin inflating.}
%\end{table}
%
%%\begin{figure}[htp]
%%\begin{minipage}{0.40\textwidth}
%%\ifTikz
%%\input{figures/starTension.tikz}
%%\fi
%%\end{minipage}
%%\hfill
%%\begin{minipage}{0.55\textwidth}
%%  \centering
%%\includegraphics[width=0.32\textwidth]{figures/StarTensionTime1.pdf}
%%\includegraphics[width=0.32\textwidth]{figures/StarTensionTime2.pdf}
%%\includegraphics[width=0.32\textwidth]{figures/StarTensionTime3.pdf} \\
%%\includegraphics[width=0.32\textwidth]{figures/StarTensionTime4.pdf}
%%\includegraphics[width=0.32\textwidth]{figures/StarTensionTime5.pdf}
%%\includegraphics[width=0.32\textwidth]{figures/StarTensionTime6.pdf}
%%\end{minipage}
%%  \caption{\label{fig:starTension} The tension of the time steps shown
%%  in Figure~\ref{fig:starShape}. The tension at the time horizon $T=1$
%%  is also included. At this point, the vesicle is nearly circular and
%%  the tension is constant and positive. The constant is given by
%%  equation~\eqref{eqn:SSshape}, and this value corresponds to the
%%  dashed black line.}
%%\end{figure}
%
%%\begin{figure}[htp]
%%\begin{minipage}{0.40\textwidth}
%%%\ifTikz
%%\input{figures/starFlux.tikz}
%%%\fi
%%\end{minipage}
%%\hfill
%%\begin{minipage}{0.55\textwidth}
%%\includegraphics[width=0.32\textwidth]{figures/StarFluxTime1.pdf}
%%\includegraphics[width=0.32\textwidth]{figures/StarFluxTime2.pdf}
%%\includegraphics[width=0.32\textwidth]{figures/StarFluxTime3.pdf} \\
%%\includegraphics[width=0.32\textwidth]{figures/StarFluxTime4.pdf}
%%\includegraphics[width=0.32\textwidth]{figures/StarFluxTime5.pdf}
%%\includegraphics[width=0.32\textwidth]{figures/StarFluxTime6.pdf}
%%\end{minipage}
%%  \caption{\label{fig:starFlux} The flux of the time steps shown in
%%  Figure~\ref{fig:starShape}. The tension at the time horizon $T=1$ is
%%  also included. At this point, the vesicle is nearly circular and there
%%  is no flux.}
%%\end{figure}
%\begin{figure}[htp]
%    \ifTikz
%    \input{figures/starRelaxation1.tikz}
%    \quad
%    \input{figures/starRelaxation2.tikz}
%    \quad
%    \input{figures/starRelaxation3.tikz}
%    \quad
%    \input{figures/starRelaxation4.tikz}
%    \quad
%    \input{figures/starRelaxation5.tikz}
%    \fi
%  \caption{\label{fig:starShape} An initially star-shaped semi-permeable
%  vesicle in a quiescent flow with $\beta=1$. Note that the time steps
%  are not equispaced.}
%\end{figure}
%\begin{figure}[htp]
%  \centering
%%  \ifTikz
%  \input{figures/ellipseVelocity1.tikz}
%  \input{figures/ellipseVelocity2.tikz}
%  \input{figures/ellipseVelocity3.tikz}
%%  \fi
%  \caption{\label{fig:vesVelocity} The velocity contributions due to
%  semi-permeability (red) and force balance (blue). The green dashed
%  line is the velocity of an impermeable vesicle with the same initial
%  shape as the semi-permeable vesicle.}
%\end{figure}
%
%\begin{figure}[htp]
%%  \ifTikz
%  \input{figures/ellipseVelocityNorms.tikz}
%%  \fi
%  \caption{\label{fig:vesVelocityNorm} The max norm of the velocity due
%  to the force balance (blue) and semi-permeability. Also included is
%  the velocity of an impermeable vesicle (green). The three black lines
%  correspond to the times reported in Figure~\ref{fig:vesVelocity}.}
%\end{figure}

% \newpage
%\input{figures/shearPhaseDiagramRA2.tikz}
%\input{figures/parabolicPhaseDiagram.tikz}
%\begin{figure}[htp]
%%  \ifTikz
%%  \input{figures/parabolicPhaseDiagram.tikz}
%%  \fi
%  \includegraphics[width=0.9\linewidth]{figures/parabolicPhaseDiagram.pdf}
%  \caption{\label{fig:parabolicPhaseDiagram} The steady state shape of a
%  semi-permeable vesicle with several different permeability rates
%  $\beta$ and flow rates $U$. The numbers reported above each vesicle is
%  its reduced area. At high flow rates, the vesicle self-intersects
%  before reaching a steady state configuration.}
%\end{figure}
%
%
%\begin{table}[htp]
%  \centering
%  \caption{\label{tbl:parabolicRelaxTimes1} The time for a
%  semi-permeable vesicle in a Poiseuille flow to reach its steady state
%  ($t_\mathrm{steady}$) and to begin inflating ($t_\mathrm{inflate}$).
%  The times are reported for the maximum velocities $0.1 \mu m/s$ and
%  $10 \mu m/s$.}
%  \begin{tabular}{C{0.8cm}C{2cm}C{0.8cm}C{2cm}C{0.8cm}}
%    \multicolumn{5}{c}{$U = 0.1 \mu m/s$} \\
%    $\beta$ & $t_\mathrm{steady}$ & ratio & $t_\mathrm{inflate}$ & ratio \\
%    \midrule
%    $10^{0}$  & $3.33 \times 10^{0}$ & ---  
%              & $1.67 \times 10^{-2}$ & --- \\ 
%    $10^{-1}$ & $1.37 \times 10^{1}$ & 4.12 & 
%                $8.73 \times 10^{-1}$ & 5.23 \\
%    $10^{-2}$ & $8.60 \times 10^{1}$ & 6.26 & 
%                $1.11 \times 10^{0}$ & 12.74 \\
%    $10^{-3}$ & $7.75 \times 10^{2}$ & 9.01 & 
%                $1.63 \times 10^{1}$ & 14.70 \\
%    $10^{-4}$ & $7.69 \times 10^{3}$ & 9.93 & 
%                $1.74 \times 10^{2}$ & 10.63 \\
%    $10^{-5}$ & $8.01 \times 10^{4}$ & 10.41 & 
%                $1.66 \times 10^{3}$ & 9.58 \\
%    \bottomrule
%  \end{tabular}
%
%  \begin{tabular}{C{0.8cm}C{2cm}C{0.8cm}C{2cm}C{0.8cm}}
%    \multicolumn{5}{c}{$U = 10 \mu m/s$} \\
%    $\beta$ & $t_\mathrm{steady}$ & ratio & $t_\mathrm{inflate}$ & ratio \\
%    \midrule
%    $10^{0}$  & $3.55 \times 10^{0}$ & ---  
%              & $1.67 \times 10^{-2}$ & --- \\ 
%    $10^{-1}$ & $1.87 \times 10^{1}$ & 5.27 & 
%                $9.01 \times 10^{-2}$ & 5.40 \\
%    $10^{-2}$ & $1.17 \times 10^{2}$ & 6.26 & 
%                $1.52 \times 10^{0}$ & 16.87 \\
%    $10^{-3}$ & $1.06 \times 10^{3}$ & 9.06 & 
%                $2.07 \times 10^{1}$ & 13.62 \\
%    $10^{-4}$ & $1.02 \times 10^{4}$ & 9.62 & 
%                $2.25 \times 10^{2}$ & 10.87 \\
%    $10^{-5}$ & $1.04 \times 10^{5}$ & 10.20 & 
%                $2.25 \times 10^{3}$ & 10.00 \\
%    \bottomrule
%  \end{tabular}
%\end{table}
%The width $W$ is chosen so that $W/R =
%10$, where $R$ is the radius of the vesicle. We investigate the effect
%of the maximum velocity $U$ and the permeability rate $\beta$ on the
%steady state shape of a vesicle initially placed on the center line
%$y=0$. Figure~\ref{fig:parabolicPhaseDiagram} shows the steady state
%configuration for a variety of maximum velocities and permeability
%rates. Similar to the shear example, the maximum velocity strongly
%affects the final shape while the permeability rate's effect is much
%weaker. Instead, the permeability rate determines the time required for
%the vesicle to reach a steady state. In
%Table~\ref{tbl:parabolicRelaxTimes1}, we consider two maximum velocities
%and report the time, $t_\mathrm{steady}$, required for the vesicle's
%area to be within 1.3\% of its maximum area, and the time,
%$t_\mathrm{inflate}$, required for the vesicle's area to exceed 1.3\% of
%its initial area. We see that both times scale with $\beta^{-1}$.

%\begin{figure}[htp]
%  \centering
%  \begin{tabular}{ccc}
%    \ifTikz
%    \scalebox{0.31}{\input{figures/stenosisTension1.tikz}} &
%    \scalebox{0.31}{\input{figures/stenosisTension3.tikz}} &
%    \scalebox{0.31}{\input{figures/stenosisTension5.tikz}} \\
%    \scalebox{0.31}{\input{figures/stenosisTension7.tikz}} &
%    \scalebox{0.31}{\input{figures/stenosisTension9.tikz}} &
%    \scalebox{0.31}{\input{figures/stenosisTension11.tikz}} 
%    \fi
%  \end{tabular}
%  \caption{\label{fig:stenosisTension} The tension of the impermeable
%  (red) and semi-permeable (dashed blue) vesicles in
%  Figure~\ref{fig:stenosisShape}. The point $\theta=0$ corresponds to
%  the right-most point of the vesicle.}
%\end{figure}
%
%\begin{figure}[htp]
%  \centering
%  \begin{tabular}{ccc}
%    \iftikz
%    \scalebox{0.31}{\input{figures/stenosisflux1.tikz}} &
%    \scalebox{0.31}{\input{figures/stenosisflux3.tikz}} &
%    \scalebox{0.31}{\input{figures/stenosisflux5.tikz}} \\
%    \scalebox{0.31}{\input{figures/stenosisflux7.tikz}} &
%    \scalebox{0.31}{\input{figures/stenosisflux9.tikz}} &
%    \scalebox{0.31}{\input{figures/stenosisflux11.tikz}} 
%    \fi
%  \end{tabular}
%  \caption{\label{fig:stenosisFlux} The flux (equivalent to pressure
%  jump) of the impermeable (red) and semi-permeable (dashed blue)
%  vesicles in Figure~\ref{fig:stenosisShape}. The point $\theta=0$
%  corresponds to the right-most point of the vesicle.}
%\end{figure}