### 2020-12-31 vesicle_code/docs/permeableVes/main.tex
starting addressing comments from Howard

\documentclass[9pt,twocolumn,twoside,lineno]{pnas-new}
% Use the lineno option to display guide line numbers if required.

\templatetype{pnasresearcharticle} % Choose template 
% {pnasresearcharticle} = Template for a two-column research article
% {pnasmathematics} %= Template for a one-column mathematics article
% {pnasinvited} %= Template for a PNAS invited submission

%\title{Semipermeable Vesicles: Mathematical Formulation, Numerics, and
%Physics}
\title{Hydrodynamics of a  Semipermeable Vesicle Under Flow and
Confinement}

% Use letters for affiliations, numbers to show equal authorship (if applicable) and to indicate the corresponding author
\author[a]{Bryan Quaife}
\author[a]{Ashley Gannon}
\author[b,1]{Y.-N.~Young}

\affil[a]{Department of Scientific Computing, Florida State University, Tallahessee, FL 32306}
\affil[b]{Department of Mathematical Sciences, New Jersey Institute of
Technology, Newark, NJ 07102}

% Please give the surname of the lead author for the running footer
\leadauthor{Quaife}

% Please add a significance statement to explain the relevance of your work
\significancestatement{Authors must submit a 120-word maximum statement
about the significance of their research paper written at a level
understandable to an undergraduate educated scientist outside their
field of speciality. The primary goal of the significance statement is
to explain the relevance of the work in broad context to a broad
readership. The significance statement appears in the paper itself and
is required for all research papers.}

% Please include corresponding author, author contribution and author declaration information
\authorcontributions{All authors contributed equally.}
%\authordeclaration{Please declare any competing interests here.}
%\equalauthors{\textsuperscript{1}A.O.(Author One) contributed equally to this work with A.T. (Author Two) (remove if not applicable).}
%\correspondingauthor{\textsuperscript{1} E-mail: bquaife\@fsu.edu}
\correspondingauthor{\textsuperscript{1} E-mail: yyoung@njit.edu}

% At least three keywords are required at submission. Please provide three to five keywords, separated by the pipe symbol.
\keywords{Keyword 1 $|$ Keyword 2 $|$ Keyword 3 $|$ ...}

\begin{abstract}
Lipid bilayer membranes have a native (albeit small) permeability for
water molecules. Provided that the bilayer structure stays intact (no membrane poration or rupture) under load, the water influx/efflux is
often assumed negligible in the absence of a osmolarity pressure across the membrane. 
%
In this work we focus on physical regimes where the membrane tension is below the critical value for membrane poration, and use boundary integral simulations to
investigate the effects of native water permeability on the hydrodynamics of
a vesicle under various flows and confinement configurations. 
%
Coupling the membrane permeability with membrane elastic stress, tension, and
mechanical stress in the framework of Helfrich free energy, we
illustrate that, in the absence of an osmotic stress gradient, a giant
unilamellar vesicle may be significantly affected by water
influx/efflux over a sufficiently long time or under a strong
confinement. Our simulation results quantify the conditions for water
permeation to be negligible in terms of the time scales, flow strength
of an external flow, and degree of confinement. These results shed
light on how microfluidic confinement can be utilized to estimate
membrane permeability.
%
%osmotic stress gradient 
%We describe a mathematical model for a semipermeable vesicle and derive
%  a boundary integral equation formulation for simulating its
%  hydrodynamics.  
%
%Please provide an abstract of no more than 250 words in a single paragraph. Abstracts should explain to the general reader the major contributions of the article. References in the abstract must be cited in full within the abstract itself and cited in the text.
\end{abstract}

\dates{This manuscript was compiled on \today}
\doi{\url{www.pnas.org/cgi/doi/10.1073/pnas.XXXXXXXXXX}}

\input my_macros2.tex

\newif\ifTikz
%\Tikztrue
\Tikzfalse

%\usepgfplotslibrary{external}
%\tikzexternalize

\begin{document}

\maketitle
\thispagestyle{firststyle}
\ifthenelse{\boolean{shortarticle}}{\ifthenelse{\boolean{singlecolumn}}{\abscontentformatted}{\abscontent}}{}

\dropcap{W}ater exchange is essential for a living cell to adapt to its
environment over a wide range of time
scales~\cite{CadartVenkovaRechoEtAl2019_NaturePhys,
AlbertsMolecularBiology, YangMaVerkman2001_JBC,
SugieIntaglietta2018_AmJPhysiolHCP,
SaadounPapadopoulosWatanabeEtAl2005_JCS, Verkman2008_JMM,
BerthaudEtAl2016_SM, Keren2011_EurBJ, TaloniKardashSalmanEtAl2015_PRL}.  
%
%Water permeation across the cell membrane due to an osmotic shock has been well studied \cite{AlbertsMolecularBiology}. 
Recent findings confirm that both osmotic and mechanical stresses
contribute to significant water permeation for cells to migrate under
strong confinement~\cite{JiangSun2013_BJ, StrokaJiangChenEtAl2014_Cell,
LiMoriSun2015_PRL, yao-mor2017}, leading to a dynamic surface-to-volume
ratio of a migrating cell. While stress-induced release of messengers in
cells has been well studied~\cite{Wan2008_PNAS, ForsythWan2011_PNAS, 
Russell-PuleriPazAdams2016_AJHCP, ZhangShenHoganBarakatMisbah2018_BJ,
GordonShimmelFrye2020_FP},
%osmotic effects on cellular dynamics have been well studied
%\cite{AlbertsMolecularBiology}, 
mechanically induced water permeation is often associated with membrane
poration or rupture under extreme
stresses~\cite{HarmanBertrandJoos2017_CJP,
RazizadehNikfarPaulLiu2020_BJ}. In this work we use modeling and direct
numerical simulations to show that, over the appropriate time scales or
under strong confinement, the intrinsic membrane permeability to water
can give rise to significantly distinct hydrodynamics of a semipermeable
vesicle without poration or rupture in the lipid bilayer membrane. 
%
%
%One of the most direct ways to induce water permeation across cell membranes is to impose an osmotic shock }.
%
%A gradient in the osmotic stress across the cellular membrane is the most direct way to induce water permeation, yet 
%%
%%In response to external stimuli, cells move, deform, and release signaling molecules, such as ATP in red blood cells \cite{ZhangShenHoganBarakatMisbah2018_BJ} and cyclooxygenase-2 in endothelial cells \cite{Russell-PuleriPazAdams2016_AJHCP}. 
%%
%%Water permeation is often neglected in such release processes, yet 
%%
%recent findings suggest that both osmotic and mechanical stresses can induce water permeation that is significant enough for tumor cells to migrate under strong confinement \cite{StrokaJiangChenEtAl2014_Cell,LiMoriSun2015_PRL,yao-mor2017}, leading to a dynamic surface-to-volume ratio over time. 
%
%
%To elucidate the effects of mechanically induced water permeation (which is a great challenge to quantify in experiments), in this work we use accurate numerical simulations to study the hydrodynamics of a semipermeable vesicle in various configurations.

%Cells also send out chemicals in response to 
%
% Erythrocytes are known to  participate in efficient body water homeostasis by regulating its exchange of water  \cite{YangMaVerkman2001_JBC,SugieIntaglietta2018_AmJPhysiolHCP}.  Cells achieve motility by simultaneously regulating their cell volume and cell membrane tension through trafficking of aquaporin proteins \cite{SaadounPapadopoulosWatanabeEtAl2005_JCS,Verkman2008_JMM,BerthaudEtAl2016_SM} and flattening of caveolae \cite{Keren2011_EurBJ}. 
%Cells also control water transport  as they activate shape fluctuations in the membrane for the bleb formation \cite{TaloniKardashSalmanEtAl2015_PRL}.
%
%
%%\dropcap{W}ater balance is an essential process in the adaptation of a living cell to its environment \cite{AlbertsMolecularBiology}: Erythrocytes are known to  participate in efficient body water homeostasis by regulating its exchange of water  \cite{YangMaVerkman2001_JBC,SugieIntaglietta2018_AmJPhysiolHCP}.  Cells achieve motility by simultaneously regulating their cell volume and cell membrane tension through trafficking of aquaporin proteins \cite{SaadounPapadopoulosWatanabeEtAl2005_JCS,Verkman2008_JMM,BerthaudEtAl2016_SM} and flattening of caveolae \cite{Keren2011_EurBJ}. 
%%Cells also control water transport  as they activate shape fluctuations in the membrane for the bleb formation \cite{TaloniKardashSalmanEtAl2015_PRL}.

The lipid bilayer membrane is permeable to water and water-soluble
molecules~\cite{Dick1964_JTB, FettiplaceHaydon1980_PhysRev,
DeamerBramhall1986_ChemPhysLipids, Grafmueller2019_ABLS}. 
%
%In the presence of an osmotic gradient across the membrane (such as a contrast in the concentration of an impermeable solute), water flows against the osmotic gradient  to neutralize the osmolarity. 
%
Without aquaporins or other membrane channels that promote water
permeability, the intrinsic water permeability is known to depend on the
temperature, lipid composition~\cite{OlbrichRawiczNeedhamEtAl2000_BJ}, 
%\todo[inline]{Same reference twice?} 
and the hydration of
the membrane~\cite{MarrinkBerendsen1994_JPhysChem}. In the red blood
cell membrane, aquaporin-1 channels give rise to a much higher water
permeability of approximately $1.8\times 10^{-2}$ cm/s in
erythrocytes~\cite{YangMaVerkman2001_JBC} while the intrinsic water
permeability of a lipid bilayer membrane is in the range of
$10^{-4}$--$10^{-3}$ cm/s~\cite{ThompsonHuang1966_ANYAS,
FettiplaceHaydon1980_PhysRev, Grafmueller2019_ABLS, Dimova2020_GVB,
BhatiaRobinsonDimova2020_SoftMatt}. 
%
%The swelling/shrinking of a cell is inevitably coupled to changes in
%membrane tension and deformation. Under a strong confinement (such as
%going through a microcapillary or submicron slit),  the transmembrane
%water flow may have comparable contributions from both osmotic and
%mechanical stresses
%\cite{StrokaJiangChenEtAl2014_Cell,LiMoriSun2015_PRL,yao-mor2017}.  
Molecular dynamic (MD) simulations show that a stretched red blood cell
membrane porates when the membrane tension reaches the order of $2$--$4$
mN/m (\cite{RazizadehNikfarPaulLiu2020_BJ} and references therein),
giving rise to enhanced membrane permeability to both water and
macromolecules. Consequently, as long as the lipid bilayer membrane
remains intact (no poration or rupture), red blood cells and vesicles
are assumed to be of the same volume throughout experiments of durations
no longer than tens of minutes.
%
%membrane poration is inevitable when membrane tension is of the order
%of $2-4$ mN/m (\cite{RazizadehNikfarPaulLiu2020_BJ} and references
%therein). 
%%The CGMD results further show that a stretched RBC membrane may have pores that might contribute to transmembrane transport of both water and ATP molecules \cite{RazizadehNikfarPaulLiu2020_BJ}. 
%Thus an important outstanding question in stress-induced ATP release of
%highly deformed RBCs is if there is any transmembrane flow when RBCs
%release ATP as they experience large mechanical load due to confinement
%or large shear stress?  To address such question, the first step we
%take is to focus on the permeable water flux induced by mechanical
%stress.
%%
%%RBCs are known to release ATP under large deformation and microfluidic experiments have been used to study the correlation between RBC deformation and ATP release \cite{Wan2008_PNAS,ForsythWan2011_PNAS,ZhangShenHoganBarakatMisbah2018_BJ}.
%%
%%Why is it interesting to study the dynamics of a semipermeable membrane induced by mechanical stress/confinement?
%%
%%The osmolarity-induced water flux has been well studied, while mechanical pressure may also contribute to the permeable water flow \cite{yao-mor2017}.
%%The response of a cell to an osmotic gradient is inevitably both mechanical (tension redistribution and shape change) and chemical (water flux across the membrane).  
Under what condition is this a good assumption? Do vesicles conserve
volume over hours under flow? What is the dynamic consequence when water
influx/efflux affects the membrane hydrodynamics? To answer these
questions, we investigate the balance between hydrodynamics, tension,
and elasticity of a semipermeable membrane under various flowing
conditions and confinement, see Figure~\ref{fig:sketch}.
%Specifically our numerical simulations yield the quantification of the
%time scale and length scale over which water permeation becomes
%important.

\begin{figure}[htp]
  \centering
  \includegraphics[width=0.5\textwidth]{figures/schematic.pdf}
  \caption{\label{fig:sketch} Schematics showing the effects of
  semipermeability on vesicle hydrodynamics in various configurations. 
  The lipid bilayer membrane (modeled as an interface $\gamma$ of zero
  thickness and an outward normal $\nn$) is permeable to water
  molecules, depending on the mechanical normal stress balance on the
  membrane.  The slipper in (a) is the steady-state shape of a
  semipermeable vesicle in a Poiseuille flow.}
\end{figure}

% In this work we use numerical simulations to examine the hydrodynamics
% of a semipermeable vesicle in two dimensions. 
In microfluidic experiments, it is challenging to keep a freely
suspended red blood cell or vesicle in a steady flow for more than a few
hours. It is also difficult to quantify the vesicle area/volume ratio
accurately.  Using numerical simulations we can quantify the effects of
water permeability on the hydrodynamics of a vesicle under mechanical
stresses. Vesicles (self-enclosed lipid bilayer membranes) have been
used as a model system to study reshaping, remodeling, and scission of
cell membranes due to osmotic stress~\cite{OgleckaEtAl2014_eLife,
CamposSaric2020_bioRxiv, Dimova2020_GVB, BhatiaChrist2020_SoftMatt}.
Under an osmotic shock, a vesicle may undergo oscillatory variation in
its enclosed volume~\cite{ChabanonHoLiedberg2017_BJ}. When trapped in a
microfluidic channel, an osmotic shock induces vesicle shape changes,
tank-treading and tumbling
motions~\cite{BhatiaRobinsonDimova2020_SoftMatt,
BhatiaChrist2020_SoftMatt}, an illustration that the permeable water
flow is inevitably coupled with the mechanical pressure jump across the
membrane~\cite{yao-mor2017}, which could become comparable to osmotic
stress when the membrane deformation is large~\cite{LiMoriSun2015_PRL,
yao-mor2017}. In the absence of an osmotic gradient, we first show that
a freely suspended semipermeable vesicle behaves differently from an
impermeable vesicle over a long time. Next we show that, at normal
physiological conditions, the permeable water flow can be amplified by
extreme confinement and results in observable change in both vesicle
hydrodynamics and reduced area. These results will shed light on the
dynamic aging process of living cells that constantly go through flows
and constrictions over a period of days.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section*{Formulation}
\begin{figure*}[hbp]
  \centering
  \ifTikz
  \input{figures/relaxationComposite.tikz}
  \else
  \includegraphics{figures/relaxationComposite.pdf}
  \fi
  \caption{\label{fig:relaxationComposite} (a) Snapshots of four
  semipermeable vesicles with $\beta = 10^{-3}$ submerged in a fluid
  with no background flow flow. Time goes from left to right. The red
  regions correspond to influx and the blue regions correspond to
  efflux. In all cases, the vesicle reaches a steady state circular
  shape. (b) The reduced area of each of the simulations. The
  particular snapshots in (a) occur at the marks along the curve.}
\end{figure*}
We consider a two-dimensional semipermeable vesicle (permeable only to
water) suspended in a viscous fluid (Figure~\ref{fig:sketch}), and we
let $\gamma$ denote the membrane boundary. We let
$\alpha=4\pi A/L^2$ be the vesicle's reduced area where $A$ is its area
and $L$ is its length. To focus on the effect of water flow on vesicle
hydrodynamics in various conditions, we assume the fluid is the same
inside and outside the vesicle membrane. On the membrane the permeable
water flux gives rise to a difference between the fluid velocity
${\uu}({\xx}\in \gamma)$ and the membrane velocity
$\dot{\xx}$~\cite{yao-mor2017}:
\begin{align}
  \uu - \dot{\xx} = - k_w (R T \triangle c + \ff_\mathrm{mem} \cdot \nn) \nn, \qquad
  \xx \in \gamma,
\end{align}
where $\nn$ is the outward normal (Figure~\ref{fig:sketch}), $k_w$ is
the water permeability, $R$ is the ideal gas constant, $T$ is the
temperature ($^{\circ}$K), $\triangle c$ is contrast in a solute
concentration (mol/volume), and $\ff_\mathrm{mem}$ is the membrane
stress that consists of both bending $ \ff_\mathrm{ben}= -k_b
\xx_{ssss}$ and tension $\ff_\mathrm{ten}=(\sigma \xx_s)_s$ ($s$ is the
arclength along the membrane, see Supplementary Material). Using
$k_w\sim 10^{-12}$ m$^2$s/kg (typical of mammalian cell
membranes~\cite{LiMoriSun2015_PRL}) and an osmolarity $\triangle c =
0.1$ mol/L (consistent with the osmotic forces in vesicle aspiration
experiments~\cite{OlbrichRawiczNeedhamEtAl2000_BJ}), the corresponding
water flow is of the order $v_\mathrm{o}=0.1 \mu$m/s. Assuming that the
membrane bending stiffness $k_b = 10^{-19}$J, a membrane tension
$10^{-3}$N/m (one tenth of the lysis tension), and a characteristic
length scale of $10^{-6}$m, the tension contribution ($v_\mathrm{t}$)
and elastic stress contribution ($v_\mathrm{e}$) to the permeable water
flow can be scaled to $v_\mathrm{o}$ as
$v_\mathrm{o}:v_\mathrm{t}:v_\mathrm{e}=10^{0}:10^{-2}:10^{-6}$.
Furthermore, we note that in the absence of an osmotic stress, the water
permeable flow can be significant when the membrane tension is large due
to the external stress such as the hydrodynamic stress due to an
external flow or confinement. In the following we will explore both.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section*{Effects of membrane water permeability}
We consider a single semipermeable vesicle, with dimensionless
permeability coefficient $\beta$, in familiar configurations to
highlight the effects of membrane water permeability. In free space we
place a semipermeable vesicle in a quiescent flow, a planar shear flow,
and a Poiseuille flow. We also consider a semipermeable vesicle in
confinement: a long close-fitting channel and a contracting channel. For
each example, we fix the vesicle length and examine the enclosed area as
a dynamic consequence of water flow, vesicle shape deformation, and
membrane tension distribution.

In experiments, vesicles under shear flows are measured for a minute
(chapter 19 in~\cite{Dimova2020_GVB}), a duration over which vesicle can
be assumed impermeable~\cite{AbkarianViallat2005_BJ}. Our simulations of
a semipermeable vesicle in these flows show that water permeability may
lead to different vesicle hydrodynamics over a time scale of 10 to 100
minutes. 
%{\bf YNY: is this correct? The shear flow results seem to suggest that
%the permeability may lead to change in area in less than a second???}
For vesicles in a narrow microfluidic channel, the typical time scale is
milliseconds~\cite{abk-fai-sto2006} \todo[inline]{Howard talked about
this statement not always being true}. Our simulations show that strong
confinement can lead to amplification of water permeability at such
short time scale.
%
%all vesicles have a length
%$L$, but the area is not conserved because of semipermeability. We vary
%the initial vesicle reduced area, the permeability rate, and the flow
%rate. For all simulations, the bending modulus is $10^{-19}J$.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection*{Quiescent Flow} 
\begin{figure*}[htp]
  \centering
  \ifTikz
  \input{figures/shearComposite.tikz}
  \else
  \includegraphics{figures/shearComposite.pdf}
  \fi
  \caption{\label{fig:shearComposite} (a) The reduced area of a
  semipermeable vesicle with $\beta = 10^{-3}$ initialized with four
  different reduced areas in a shear flow with two different flow rates.
  (b) The steady state shape of a semipermeable vesicle with varying
  flow rates and semipermeability rates. The circled vesicles correspond
  to the simulations in part (a). \todo[inline]{Runs at smallest value
  of $\beta$ are being redone.}}
\end{figure*}
In the absence of a background flow, a semipermeable vesicle in free
space can be quantified in terms of its area dynamics and its transition
to the equilibrium shape. Our analysis in Supplementary Material shows
that a semipermeable vesicle relaxes to a circle at a rate that is
proportional to $\beta$, and the steady state tension is $\Lambda =
2\pi^2/L^2$. Our analysis further shows that the time for a
semipermeable vesicle to deviate from its initial reduced area and to
reach its steady state is proportional to $\beta^{-1}$. For this reason,
our numerical results are reported in terms of the scaled time $\beta
t$.

In Figure~\ref{fig:relaxationComposite}, we examine the relaxation
dynamics of a semipermeable vesicle with four initial shapes and reduced
areas as labeled. The vesicles on the left are color-coded by the water
permeation flux: red denotes an influx into the vesicle, and blue
denotes an efflux out of the vesicle.
%
First we observe that the initial vesicle shape determines the direction
of water flux. For the vesicle with an initial reduced area
($\alpha(0)=0.12$ in the bottom row) the vesicle deflates (blue regions)
first due to the high interior capillary pressure that corresponds to
the high membrane curvatures. For the other three cases efflux is found
at some locations of high curvature, while mostly water permeates into
the vesicle and significant change in the reduced area is found at
$t\sim 10$ minutes. Over time (about ten hours) all vesicles inflate and
as expected, each vesicle eventually reaches a circular shape with
reduced area $\alpha=1$. 
%%These regions are denoted in blue and are only present at early times. 
%Finally, the vesicles's reduced area do not steady state tension is
%$\Lambda = 2\pi^2/L^2$. 
%Therefore, the times for the vesicle to deviate from its initial reduced area and to reach its
%steady state are proportional to $\beta^{-1}$. 
%
%Each of the four cases in Figure~\ref{fig:relaxationComposite} eventually
%reaches a circular shape with reduced area $\alpha=1$. Next, we observe
%that even though the vesicle is inflating to a circle, 
%%meaning that there is a net flux into the vesicle, 
%there are regions where the flux
%is out of the vesicle. These regions are denoted in blue and are only
%present at early times. 
%
%Finally, the vesicles's reduced area do not
%change by much at early times (up to ten minutes), and

Between $t=10$ seconds and $t=10$ minutes, the intermediate shapes of
semipermeable vesicles resemble the shape of an impermeable vesicle.
This can be explained by considering the two terms that contribute to
the vesicle velocity: (i) the permeability velocity, $\beta(\ff \cdot
\nn) \nn$, and (ii) the force balance velocity, $\SS[\ff]$. Since
$\beta\ll 1$, the permeability velocity is much smaller than the force
balance velocity at early times, and the semipermeable vesicle dynamics
resemble those of an impermeable vesicle. As the force balance velocity
decreases, the permeability velocity dominates, and the vesicle begins
to inflate to a circle with no vanishing force and velocity.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection*{Planar Shear Flow}

We consider a semipermeable vesicle in a planar shear flow
$\uu_{\infty}(\xx) = \dot{\gamma} (y,0)$ with different permeability
$\beta$ and shear rates $\dot{\gamma}$. When $\beta=0$ (impermeable
case), a vesicle's reduced area and viscosity contrast determine whether
it tank treads or tumbles~\cite{fin-lam-sei-gom2008,
kra-win-sei-lip1996}. Focusing on semipermeable vesicles with no
viscosity contrast, we find that a semipermeable vesicle tilts to an
inclination angle and undergoes tank treading dynamics, similar to the
case of an impermeable vesicle.

In Figure~\ref{fig:shearComposite}(a), using four initial reduced areas
and two flow rates, we plot the reduced area of a semipermeable vesicle
with $\beta = 10^{-3}$. As observed in a quiescent flow, the final
vesicle shape is independent of the initial reduced area. However, it
does depend on the flow rate. We also observe that the reduced area
remains constant until $t\sim 10$ minutes, when the vesicle starts to
tank-tread and the normal component of the force-balance velocity
becomes significantly smaller than the permeable velocity, leading to a
changing reduced area.

For a variety of semipermeability rates and flow rates, a semipermeable
vesicle reaches a steady equilibrium as in
Figure~\ref{fig:shearComposite}(a). Since the final shape is independent
of the initial reduced area, we initialize the vesicle with reduced area
$\alpha = 0.65$ for all simulations in
Figure~\ref{fig:shearComposite}(b), where we summarize the steady state
vesicle shape and its reduced area. We see that the flow rate
significantly affects the final vesicle shape, but the semipermeability
rate has a minor effect. Instead, as we saw in the quiescent analysis,
$\beta$ sets a time scale since the time required for a vesicle to reach
its steady state configuration is proportional to $\beta^{-1}$.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection*{Poiseuille Flow}
\begin{figure*}[htp]
  \centering
  \ifTikz
  \input{figures/parabolicComposite.tikz}
  \else
  \includegraphics{figures/parabolicComposite.pdf}
  \fi
  \caption{\label{fig:parabolicComposite} (a) The steady state shape of
  a semipermeable vesicle submerged in a Poiseuille flow with varying
  initial reduced areas and flow rates. The red regions correspond to
  influx and the blue regions correspond to efflux. (b) The steady
  state vertical displacement for four different flow velocities. The
  steady state shapes are superimposed. (c) The reduced area of the
  slipper formed with the flow rate of $800 \mu$m/s. (d) The vesicle
  shape shown with the background imposed flow. The channel width is
  12.5 times larger than the vesicle radius. The corresponding reduced
  area of each vesicle is indicated by the marks in plot (c).}
\end{figure*}

We consider a semipermeable vesicle in the Poiseuille flow $\uu =
U(1-(y/W)^2,0)$, where $U$ is the maximum velocity at $y=0$. The
non-linear effects of the Poiseuille flow on the migration of an
impermeable vesicle have been well-studied~\cite{kao-bir-mis2009}: At a
given value of the maximum velocity $U$, there exists a critical reduced
volume above which the equilibrium vesicle shape is symmetric (parachute
or bullet shapes) and the vesicle stays at $y=0$. Below this critical
reduced volume the equilibrium vesicle moves off the center and is
asymmetric (tank-treading slipper shape). We validated our numerical
codes by comparing against results in Kaoui {\em et
al.}~\cite{kao-bir-mis2009}, see Supplementary Material.

%
%
%We examine the effect of the non-linear Poiseuille flow on the migration of
%semipermeable vesicle. In this experiment, the vesicle is initialized
%above the center line $y=0$, and it migrates towards the center
%line~\cite{dan-vla-mis2009}. 
%
%For an impermeable vesicle, depending on
%the vesicles reduced area and the flow rate, the equilibrium vesicle shape
%is either an axisymmetric parachute or bullet, or an asymmetric
%tank-treading slipper~\cite{kao-bir-mis2009}. 
%We validated our numerical codes by comparing against results in Kaoui {\it et al.} \cite{kao-bir-mis2009}.
%One of the most salient features of an impermeable vesicle in the Poiseuille flow is that, 
%%
%%To validate our methods,
%%in Figure~\ref{fig:parabolicComposite}(a), we construct a phase diagram
%%for the steady state shape of an impermeable vesicle
%%vesicle~\cite{kao-bir-mis2009} (see Figure 2). While the vesicle shapes
%%qualitatively agree, Kaoui et al.~report that the vesicles with
%%$\nu=0.7$ and flow rates $U = 1200 \mu m/s$ and $U = 1600 \mu m/s$ are
%%axisymmetric. In our simulations these vesicles are slightly asymmetric,
%%are undergoing a tank treading motion, and we therefore report them as
%%asymmetric slippers.
%
%Having validated our method, we investigate the dynamics of a semipermeable vesicle in a Poiseuille flow. 
%
%Using several initial
%$U(1-(y/W)^2,0)$, where $U$ is the maximum imposed velocity. We examine
%the effect of the non-linear Poiseuille flow on the migration of
%semipermeable vesicle. 
%

At the beginning of the simulations a semipermeable vesicle is
initialized above the center line $y=0$, and it migrates towards the
center line as in the impermeable case~\cite{dan-vla-mis2009}. 
%
%However,
%for an impermeable vesicle, depending on the initial reduced area and
%the maximum flow velocity $U$, the equilibrium vesicle shape is either
%an axisymmetric parachute or bullet, or an asymmetric tank-treading
%slipper~\cite{kao-bir-mis2009}. 
%
%To validate our methods,
%in Figure~\ref{fig:parabolicComposite}(a), we construct a phase diagram
%for the steady state shape of an impermeable vesicle
%vesicle~\cite{kao-bir-mis2009} (see Figure 2). While the vesicle shapes
%qualitatively agree, Kaoui et al.~report that the vesicles with
%$\alpha=0.7$ and flow rates $U = 1200 \mu m/s$ and $U = 1600 \mu m/s$ are
%axisymmetric. In our simulations these vesicles are slightly asymmetric,
%are undergoing a tank treading motion, and we therefore report them as
%asymmetric slippers.
%
Using several initial reduced areas and flow rates, we recreate the
impermeable vesicle phase diagram for vesicles with a permeability rate
of $\beta = 10^{-3}$. The equilibrium vesicle shape for different $U$
and initial reduced area is summarized in
Figure~\ref{fig:parabolicComposite}(a), and the vertical displacement of
the vesicle's center of mass is in
Figure~\ref{fig:parabolicComposite}(b). The effect of semipermeability
of a vesicle with $\alpha(0) = 0.9$ and $U=800 \mu$m/s on its reduced
area and migration pattern are shown in
Figure~\ref{fig:parabolicComposite}(c) and (d).

As observed for the planar shear flow, the steady state reduced area is
independent of the initial reduced area, and larger flow rates result in
smaller steady state reduced areas. We report the reduced area of the
equilibrium vesicle shape in Figure~\ref{fig:parabolicComposite}(a). At
the smallest flow rate of $U= 200 \mu$m/s, the steady state reduced
area is large, and the result is an axisymmetric bullet. At the larger
flow rates, the steady state reduced area is smaller, and the result is
an asymmetric tank-treading slipper. We observe that the membrane
permeability to water drastically alters the vesicle shape and position
relative to the flow at equilibrium: At $\beta=10^{-3}$ the equilibrium
vesicle depends only on the maximum velocity $U$ of the nonlinear shear
flow. The higher $U$ the smaller the equilibrium reduced area and the
further away vesicle is relative to the flow center at $y=0$.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection*{Semipermeable Vesicle in a Narrow Channel (Stenosis)}
 \begin{figure*}[htp]
  \centering
  \ifTikz
  \input{figures/stenosisComposite.tikz}
  \else
  \includegraphics{figures/stenosisComposite.pdf}
  \fi
  \caption{\label{fig:stenosisComposite} (a) A semipermeable vesicle
  passing through a constricted geometry similar to the experimental
  device in~\cite{abk-fai-sto2006}. (b) A semipermeable vesicle passing
  through a constricted geometry that is over five times longer. (c)
  The excess pressure of an impermeable and semipermeable vesicle in the
  short geometry in (a). (d) The reduced area of the semipermeable
   vesicles in the geometries in (a) and (b) as a function of the
   vesicles' center of mass. In (c) and (d), the $x$-axis is scaled by
   $1/L$ where $L$ is half the length of the constriction. (e) The
   reduced area of the semipermeable vesicles in the geometries in (a)
   and (b) as a function of time. The time axis is scaled by $1/t_R$
   where $t_R$ is the residency time. In parts (c) and (d), the dashed
   lines correspond to the locations of the inlet and outlet. In part
   (e), the dashed lines correspond to the times that the vesicle enters
   and exits the constriction. The colored vesicles show the tension
   distribution in N/m of the circled vesicles. Throughout the entire
   simulation, these vesicle configurations have the largest tension.
   \todo[inline]{Figure out the oscillations in excess pressure}}
\end{figure*}

Cells experience stretching and release messengers when they pass
through capillary vessels~\cite{Wan2008_PNAS, ForsythWan2011_PNAS,
Pak2015_PNAS}.  Many cells exhibit extraordinary flexibility as they go
through narrow channels~\cite{AuStoreyMoore2016_PNAS}, and adjustment of
area to volume ratio may be a key factor for a successful passage.  Here
we quantify the effect of water permeability on a vesicle going through
a stenosis (a long microfluidic channel with a width smaller than the
cell radius).  Based on the findings in~\cite{abk-fai-sto2006}, we first
examine the effect of semipermeability on the pressure fluctuation
across a stenosed geometry. In this experiment, a pressure drop is used
to squeeze a red blood cell through a constriction with diameter $5\mu$m
and length $45\mu$m. We set up a similar computational geometry with a
Poiseuille flow imposed at the inlet and outlet
(Figure~\ref{fig:stenosisComposite}(a)). The flow velocity $U_{max} =
1000 \mu$m/s is maintained at the inlet and outlet and is chosen so that the residency time of the vesicle in the
stenosis channel is $t_R \approx 0.5$s which agrees with the
experimental results~\cite{abk-fai-sto2006}. When the red blood cell
enters the channel, an increase in the pressure drop is required to
maintain a constant flow rate, and the difference in the pressure drop
is called the excess pressure.

We consider both an impermeable and semipermeable vesicle with initial
reduced area $\alpha = 0.65$ passing through a long narrow channel and
compute the excess pressure as a function of vesicle position in the
channel, see Figure~\ref{fig:stenosisComposite}(c). The impermeable case
agrees with experimental results~\cite{abk-fai-sto2006} (see Figure 3).
Our simulations show that less excess pressure is needed to drive a
semipermeable vesicle through the closely-fit channel. Once in the
channel, the vesicle takes a bullet shape with maximum tension in the
front and minimum tension in the back (see
Figure~\ref{fig:stenosisComposite}(a)), consistent with earlier
results~\cite{Pak2015_PNAS, HarmanBertrandJoos2017_CJP}.

\begin{figure}[htp]
  \centering
  \ifTikz
  \input{figures/contractingComposite.tikz}
  \else
  \includegraphics{figures/contractingComposite.pdf}
  \fi
  \caption{\label{fig:contractingComposite} A contracting geometry
  similar to the experimental device in~\cite{wu2015critical}. The
  channel width varies from $18\mu$m to $1.8\mu$m, and it repeats every
  $45\mu$m. The fluid flux along the red regions is into the vesicle,
  and the fluid flux along the blue regions is out of the vesicle.
  (a--c): The ability for a semipermeable vesicle to pass through the
  contraction depends on the semipermeability rate. (d--f): The ability
  for a semipermeable vesicle to pass through the contraction depends
  on the flow rate. The colored vesicles show the tension distribution
  in N/m of the circled vesicles. Throughout the entire simulation,
  these vesicle configurations have the largest tension.}
\end{figure}

Figure~\ref{fig:stenosisComposite}(d)) shows that the vesicle deflates
from the beginning due to the high flow velocity, consistent with the
results for a vesicle in a Poiseuille flow in open space. As the vesicle
enters the stenosis channel, it continues to deflate in the front even
though we observe influx (red segment) at various rear locations in the
membrane. Overall, the vesicle loses about 6.1\% of its initial total
water at a constant rate as shown in
Figure~\ref{fig:stenosisComposite}(d). Thus we expect that a larger
amount of deflation is possible if the vesicle is inside a longer
stenosis: we consider a vesicle passing through a channel of triple the
length in Figure~\ref{fig:stenosisComposite}(b) with $U_{max}=8.4
\mu$m/s, and we observe that vesicle stays inside the stenosis for a
residence time $t_R=16.2$s and the reduced area decreases by 13.2\%
(Figure~\ref{fig:stenosisComposite}(e)).

%\begin{figure*}[htp]
%  \ifTikz
%  \input{figures/stenosisComposite.tikz}
%  \else
%  \includegraphics{figures/stenosisComposite.pdf}
%  \fi
%  \caption{\label{fig:stenosisComposite} (a): A semipermeable vesicle
%  passing through a constricted geometry similar to the experimental
%  device in~\cite{abk-fai-sto2006}. (b): A semipermeable vesicle passing
%  through a constricted geometry that is over five times longer. (c):
%  The excess pressure of an impermeable and semipermeable vesicle in the
%  short geometry in (a). (d): The reduced area of the semipermeable
%  vesicles in the geometries in (a) and (b) as a function of the
%  vesicles' center of mass. The $x$-axis is scaled by $1/L$ where $L$ is
%  half the length of the constriction. (e) The reduced area of the
%  semipermeable vesicles in the geometries in (a) and (b) as a function
%  of time. The time axis is scaled by $1/t_R$ where $t_R$ is the
%  residency time. In parts (c) and (d), the dashed lines correspond to
%  the locations of the inlet and outlet. In part (e), the dashed lines
%  correspond to the times that the vesicle enters and exits the
%  constriction. The colored vesicles show the tension distribution in
%  N/m of the circled vesicles. Throughout the entire simulation, these
%  vesicle configurations have the largest tension.}
%\end{figure*}
%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection*{Semipermeable Vesicle in a Contracting Geometry} 
\begin{figure}[htp]
  \centering
  \ifTikz
  \input{figures/contractingComposite2.tikz}
  \else
  \includegraphics{figures/contractingComposite2.pdf}
  \fi
  \caption{\label{fig:contractingComposite2} Vesicle reduced area and
  velocity vs.~time. Permeability rate is $\beta = 10^{-3}$ and maximum
  velocity is $18\mu$m/s. (a) A semipermeable vesicle passing through
  six periods of the contracting geometry. (b) The reduced area with
  the marks representing shapes in (a). (c) The vesicle velocity with
  the marks representing shapes in (a). In (b) and (c), the bottom
  x-axis is the $x$-coordinate of the center of mass, and the upper
  x-axis is time. The colored vesicles show the tension distribution in
  N/m of the circled vesicles. Throughout the entire simulation, these
  vesicle configurations have the largest tension.\todo[inline]{Make the
  time steps equispaced and the values of $x$ non-uniform. Reverse their
  orders between to and bottom.}}
\end{figure}

Squeezing a cell or an elastic vesicle through a slit in microfluidics
has gained popularity for its application in stress-induced release of
macromolecules~\cite{ShareiEtAl2013_PNAS, Pak2015_PNAS,
ZhangShenHoganBarakatMisbah2018_BJ, LuoBai2019_PoF}. Red blood cells go
through submicron slits many times during their lifetime, and such
process selects healthy, flexible red blood cells to
survive~\cite{wu2015critical,LuPeng2019_PoF}. Here we investigate the
role of membrane permeability to water when a semipermeable vesicle is
squeezed through a contracting microfluidic channel similar
to~\cite{wu2015critical}, where the width of the channel gradually
decreases from $18 \mu$m to $1.8 \mu$m. The channel then immediately
opens again to $18 \mu$m (Figure~\ref{fig:contractingComposite}). A
Poiseuille flow is imposed at the inlet and outlet with a constant
maximum flow rate of $U_{\max}$. We initialize a semipermeable vesicle
with $\alpha = 0.65$ that spans $10 \mu$m of the total channel width and
investigate the effects of the membrane permeability to water and the
maximum flow velocity on whether a vesicle can successfully pass through
the slit. We color the vesicle according to the sign of its flux---red
for influx and blue for efflux.

In Figures~\ref{fig:contractingComposite}a
and~\ref{fig:contractingComposite}b, we plot snapshots of vesicles with
different permeability $\beta$. Figure~\ref{fig:contractingComposite}c
shows the corresponding reduced area as a function of the vesicles'
location. We observe that if the permeability is sufficiently large, the
vesicle inflates gradually over seconds before it reaches the neck, and
then it quickly deflates within a fraction of a second. However, if the
semipermeability rate is too low, the vesicle is too large when it
reaches the narrowest part of the channel, and it is unable to pass. In
Figures~\ref{fig:contractingComposite}d
and~\ref{fig:contractingComposite}e, we plot snapshots of vesicles with
different imposed maximum velocities, and
Figure~\ref{fig:contractingComposite}f shows the reduced area as a
function of the vesicles' location. At low velocities, the vesicle
inflates similar to the unbounded parabolic flow example, and it is
unable to pass through the contraction. However, higher flow rates
result in additional deflation and the vesicle passes through the
contraction.

To understand the effects of water permeability on cells that repeatedly
go through micro-capillary vessels or sub-micron slits, we construct a
contracting geometry (see Figures~\ref{fig:contractingComposite2}) to
simulate a vesicle passing through the contracting geometry six times.
In Figure~\ref{fig:contractingComposite2}b, we show that the vesicle
deflates each time it passes through the contraction followed by a
gradual inflation (over seconds). The total area loss over the history
of the vesicle is almost 50\%. In
Figure~\ref{fig:contractingComposite}c, we show the vesicle velocity as
a function of the vesicle's center of mass. As the vesicle passes
through the contraction, it first accelerates until it reaches the neck
where it abruptly decelerates to go through the contraction. The
variation of the reduced area and velocity with respect to time are also
shown in Figure~\ref{fig:contractingComposite2}.

%
%\begin{itemize}
%  \item Vesicle velocity in plot (a) varies from 5 to 24 microns/s.
%  \item Vesicle velocity in plot (b) varies from 5 to 28 microns/s. Time
%    that vesicle first completely passes through constriction is 4.75s
%  \item Vesicle velocity in plot (c) varies from 0.2 to 2 microns/s.
%  \item Vesicle velocity in plot (d) varies from 5 to 25 microns/s. Time
%    that vesicle first completely passes through constriction is 4.76s
%  \item Vesicle velocity in plot (g) varies from 14 to 95 microns/s (see
%    plot (i)). Time that vesicle first completely passes through
%    constriction is 1.43s
%\end{itemize}

\section*{Discussion}
In this work we investigated the effects of membrane permeability to
water on vesicle hydrodynamics. Without any external flow or
confinement, our analysis shows that a semipermeable vesicle always
inflates to maximize the water content inside the vesicle of a fixed
length (area) in two (three) dimensions. Results from analysis and
numerical simulation show that such relaxation of a semipermeable
vesicle is characterized by a rescaled time $\beta t$. We further
illustrate that the relaxation process consists of at least two phases:
The first phase is dominated by the balance between membrane traction
and viscous stress at early times (up to a minute), when the vesicle
behaves like an impermeable membrane. The second phase sets in after a
few minutes, when water influx begins to dominate and inflate the
vesicle. Over a time scale of a few hours, a semipermeable vesicle is
fully inflated, independent of its initial surface-to-volume ratio.

Under strong confinement we show that water permeation can be amplified
on a time scale of a few seconds as a vesicle is squeezed through a
closely fit channel and a contracting channel. As a semipermeable
vesicle goes through a long closely-fit channel, we observe that the
vesicle loses water at a constant rate, and the longer the channel the
more the loss of water. In the case of a contracting channel, we find
that both permeability and the pressure gradient across the slit must be
sufficiently large for a successful passage. Throughout these
simulations the membrane tension is below the threshold for membrane
poration. Thus we expect the continuum membrane theory to well-capture
the vesicle hydrodynamics, and our simulation results shed light on the
configuration and time scales needed to observe the effects of
permeability on vesicle hydrodynamics under confinement.

% \bibliographystyle{plain}
 \bibliography{refs}

\section*{Supplementary Material: Formulation}
We consider a two-dimensional inextensible semipermeable vesicle,
$\gamma$, suspended in unconfined and confined geometries $\Omega$. The
vesicle is parameterized as $\xx(s,t)$ where $s$ is arclength and $t$ is
time. The suspending fluid is assumed to be viscous so that the fluid
velocity $\uu$ and pressure $p$ are governed by the incompressible
Stokes equations
\begin{align}
  \label{eqn:governing}
  -\nabla p + \mu \triangle \uu = 0, \quad
  \nabla \cdot \uu = 0, \qquad \xx \in \Omega \setminus \gamma,
\end{align}
where $\mu$ is the fluid viscosity. In an unconfined geometry, the
condition $\uu(\xx) \rightarrow \uu_\infty(\xx)$ is enforced as $|\xx|
\rightarrow \infty$, and in a confined geometry, the boundary condition
$\uu(\xx) = \UU(\xx)$ is enforced on $\partial \Omega$, where $\UU(\xx)$
is prescribed. Along the vesicle, mass continuity, force balance, and
local inextensibility are enforced by requiring
\begin{align}
  \label{eqn:jump}
  \jump{\uu} = 0, \quad
  \jump{T\cdot\nn} = \ff_\mathrm{mem}, \quad
  \xx_s \cdot \uu_s = 0,
\end{align}
where $\nn$ is the outward normal of $\gamma$, $\jump{\cdot}$ is the
jump across the interface, and $T$ is the hydrodynamic stress tensor.
The membrane force $\ff_\mathrm{mem}$ is the sum of a bending force
with bending modulus $k_b$ and a force due to the tension $\Lambda$.
Following~\cite{vee-gue-zor-bir2009}, we introduce the variable $\sigma
= \Lambda - 1.5\kappa^2$, where $\kappa$ is the vesicle curvature. Then,
the bending and tension forces corresponding to the Helfrich energy are
$\ff_\mathrm{ben} = -k_b \xx_{ssss}$ and $\ff_\mathrm{ten} = (\sigma
\xx_s)_s$.

Semipermeability is introduced by modifying the vesicle velocity from
the standard no-slip condition $\dot{\xx} = \uu(\xx)$. Instead, we
assume that the flux through the vesicle is proportional to the
mechanical force~\cite{yao-mor2017}. That is, 
\begin{align}
  \label{eqn:vesicleFlux}
  \uu - \dot{\xx} = - k_w (\ff_\mathrm{mem} \cdot \nn) \nn, \qquad
  \xx \in \gamma,
\end{align}
where $k_w$ is the water permeability.

The governing equations are non-dimensionalized by introducing a length
scale that we take to be $R_0 = 10^{-6}$m. Using a bending stiffness of
$k_b = 10^{-19}$J and fluid viscosity $5 \times 10^{-2}$kg/ms, we define
the time scale $\tau = \mu R_0^3/k_b = 0.5s$, the velocity scale $U =
R_0/\tau = 2\mu$m/s, the pressure scale $P = k_b/R_0^3 = 10^{-1}$Pa, the
tension scale $Q = k_b/R_0^2 = 10^{-7}$N/m, and the permeability scale
$W = R_0\mu = 2 \times 10^{-5}$m$^2$s/kg. Another standard dimensionless
variable in vesicle suspensions is the reduced area $\alpha = 4\pi A/L^2
\in (0,1]$, where $L$ is the vesicle length and $A$ is its area. The
length is conserved by the local inextensibility condition, but the area
is not because of semipermeability. Therefore, the reduced area is
dynamic.

Since the fluid satisfies the Stokes equations, the velocity and
pressure can be represented as layer potentials. Given the interfacial
boundary conditions~\eqref{eqn:jump}, the fluid velocity is
\begin{align}
  \uu(\xx) = \uu_\infty(\xx) + \SS[\ff_\mathrm{mem}](\xx), \quad
    \xx \in \Omega,
\end{align}
where $\SS$ is the single-layer potential
\begin{align}
  \SS[\ff](\xx) = \frac{1}{4\pi\mu} \int_{\gamma} \left(
    -\mathbf{I} \log\rho + \frac{\rr \otimes \rr}{\rho^2} \right)
    \ff(\yy) ds_{\yy},
\end{align}
with $\rr = \xx - \yy$ and $\rho = |\rr|$. Imposing the flux
condition~\eqref{eqn:vesicleFlux} and the inextensibility condition, the
vesicle velocity satisfies the boundary integral equation
\begin{align}
  \label{eqn:vesVelocity}
  \dot{\xx} &= \uu_\infty(\xx) + \beta (\ff_\mathrm{mem}\cdot\nn)\nn
  + \SS [\ff_\mathrm{mem}](\xx),  \quad
  \xx_s \cdot \dot{\xx}_s = 0,
\end{align}
where $\beta = k_w/W$ is the dimensionless water permeability. In
confined geometries, the far field term $\uu_\infty$ is replaced by a
double-layer potential with a density defined on $\partial\Omega$. The
unknown density function allows us to satisfy the Dirichlet boundary
condition $\uu(\xx) = \UU(\xx)$ for $\xx \in \partial\Omega$.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section*{Supplementary Material: Numerical Methods}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection*{Discretization in Space}
By using a boundary integral equation formulation, the unknowns are the
vesicle's position and tension and the unknown density function for
confined flows. Both $\gamma$ and $\partial\Omega$ are discretized at
$N$ and $N_\text{wall}$ points, respectively. Derivatives are computed
with Fourier differentiation, and the eighth-order
quadrature~\cite{alp1999} is applied to the weakly singular single-layer
potential. For confined geometries, the trapezoid rule is used to
approximate the double-layer potential since its kernel is smooth and
periodic. Nearly-singular integrals are computed with an
interpolation-based quadrature rule~\cite{qua-bir2014}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection*{Discretization in Time}
To eliminate stringent time step restrictions, we modify a time stepping
method that discretizes high-order derivatives
semi-implicitly~\cite{vee-gue-zor-bir2009}. We introduce the notation
\begin{alignat}{3}
  &B[\xx]\ff = -\frac{d^4}{ds^4} \ff,  \qquad
  &&T[\xx]\sigma = (\sigma \xx_s)_s, \\
  &D[\xx]\ff = \xx_s \cdot \ff_s, 
  &&P[\xx]\ff = (\ff \cdot \nn) \nn,
\end{alignat}
for the differential and projection operators. Then, the
governing equations~\eqref{eqn:vesVelocity} are
\begin{align}
  \dot{\xx} &= \uu_\infty(\xx) + \beta P(B\xx + T\sigma)
  + \SS (B\xx + T\sigma), \quad
  D \dot{\xx} = 0.
\end{align}
We note that the integral and differential operators, which all depend
on the vesicle shape, are non-linear with respect to the vesicle
shape. We construct a time stepping scheme by letting $B^N$ be the
bending operator due to the vesicle configuration at time $t^N$, and use
similar notation for the other operators. Then, a first-order
semi-implicit time stepping method of~\eqref{eqn:vesVelocity} is
\begin{alignat}{3}  
  \frac{\xx^{N+1} - \xx^N}{\triangle t} &= \uu_\infty(\xx^N) 
  &&+ \beta P^N\left(B^N\xx^{N+1} + T^N\sigma^{N+1}\right) \nonumber \\
  & &&+ \SS^N\left(B^N\xx^{N+1} + T^N\sigma^{N+1}\right),  \\
  D^N\xx^{N+1} &= 1,
\end{alignat}
and this linear system is solved with matrix-free GMRES.

To achieve second-order accuracy over long time horizons, we modify an
adaptive spectral deferred correction time stepping
method~\cite{qua-bir2016}. Since the area of the vesicle is not
conserved, we only use the length of the vesicle to estimate the error
which determines if a time step is accepted or not and determines the
subsequent time step size.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection*{Numerical Validation}
To validate our methods, we construct a phase diagram of the steady
state shape of an impermeable vesicle in a Poiseuille flow. Depending on
the flow rate and the reduced area, the steady state vesicle becomes
either a axisymmetric bullet, a axisymmetric parachute, or a tank
treading asymmetric slipper~\cite{kao-bir-mis2009}. Comparing our
results (Figure~\ref{fig:PoiseuillePhase}) with those
in~\cite{kao-bir-mis2009} (see Figure 2), we see that the vesicle shapes
qualitatively agree. We note that Kaoui {\em et al.}~report that the
vesicles with $\alpha=0.7$ and flow rates $U = 1200 \mu$ m/s and $U =
1600 \mu$m/s are axisymmetric.  In our simulations these vesicles are
slightly asymmetric, are undergoing a tank treading motion, and we
therefore report them as asymmetric slippers.
\begin{figure}[htp]
  \centering
  \ifTikz
  \input{figures/parabolicImpermeable.tikz}
  \else
  \includegraphics{figures/parabolicImpermeable.pdf}
  \fi
  \caption{\label{fig:PoiseuillePhase} Phase diagram of an impermeable
  vesicle with various reduced areas submerged in a Poiseuille flow with
  various flow rates. The shapes agree qualitatively with Kaoui {\em et
  al.}~\cite{kao-bir-mis2009} (see Figure 2).}
\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section*{Supplementary Material: Vesicle Area}
In the absence of an imposed flow, the steady state vesicle shape and a
differential equation for the area of the vesicle can be derived. We
start by following the analysis of Veerapaneni {\em et
al.}~\cite{vee-raj-bir-pur2009} by computing the steady state shape of a
semipermeable vesicle in a quiescent flow. The steady state shape is a
stationary point of the Lagrangian
\begin{align}
  \mathcal{L} = \frac{1}{2}\int_{\gamma} \kappa^2 \, ds +
    \Lambda \left(\int_{\gamma} ds  - L \right),
\end{align}
where the Lagrange multiplier $\Lambda$ is the tension. Note the absence
of the pressure Lagrange multiplier since the vesicle's area fluctuates.
Taking the variation of $\mathcal{L}$ with respect to $\gamma$ and
setting it to zero, the steady state shape satisfies $\kappa_{ss} +
\frac{1}{2}\kappa^3 - \Lambda \kappa = 0$, with initial conditions
$\kappa(0) = \kappa_0$ and $\kappa_s(0) = 0$.  Integrating once, we have
\begin{align}
  \frac{\kappa_s^2}{2} + \frac{\kappa^4}{8} - 
    \frac{\Lambda}{2}\kappa^2 = \frac{\kappa_0^4}{8} - 
    \frac{\Lambda}{2}\kappa_0^2.
  \label{eqn:curveODE}
\end{align}
Assuming $\kappa_s \neq 0$,~\eqref{eqn:curveODE} is separable, and can
be solved analytically in terms of the {\em EllipticF} function using
Mathematica. Unlike in the impermeable case, the solution
of~\eqref{eqn:curveODE} is not periodic unless $\kappa(s) = \kappa_0$.
Therefore, the steady state shape of a semipermeable vesicle in a
quiescent flow is a circle with radius $L/2\pi$ and constant tension
$\Lambda = \kappa^2/2 = 2\pi^2/L^2.$ Note that in contrast to a
impermeable vesicle, the tension is positive rather than negative.

We next compute the transient area of the vesicle. Since the area of the
vesicle is
\begin{align}
  A(t) = \frac{1}{2}\int_{\gamma} (\xx \cdot \nn)\, ds,
  \label{eqn:area}
\end{align}
its derivative includes terms due to the time derivative of $\xx \cdot
\nn$ and a term due to interface stretching~\cite{lai-tse-hua2008}, but
this latter term term is zero because of the local inextensibility
condition. Therefore,
\begin{align}
  \dot{A}(t) =
  \frac{1}{2} \int_{\gamma} (\dot{\xx} \cdot \nn)\, ds  + 
  \frac{1}{2} \int_{\gamma} (\xx \cdot \dot{\nn})\, ds.
\end{align}
Applying integration by parts, the two integrals agree, and we have
%\begin{align}
%  \int_\gamma (\xx \cdot \dot{\nn})\, ds 
%    = \int_\gamma \left(\xx \cdot \dot{\xx}_s^\perp \right)
%    = -\int_\gamma \left(\xx^\perp \cdot \dot{\xx}_s\right)\, ds 
%    = \int_\gamma \left(\xx_s^\perp \cdot \dot{\xx}\right)\, ds 
%    = \int_\gamma (\dot{\xx} \cdot \nn)\, ds.
%\end{align}
\begin{align}
  \dot{A}(t) = \int_\gamma \left(\beta (\ff \cdot \nn)\nn 
    + \SS[\ff]\right) \cdot \nn\, ds 
  = \beta \int_\gamma (\ff \cdot \nn)\, ds,
\end{align}
where we have used the incompressibility of the single-layer potential.
Repeatedly applying integration by parts, we have
\begin{align}
  \dot{A}(t) &= \beta \int_\gamma \left( -\xx_{ssss} + 
    (\sigma \xx_s)_s \right) \cdot \nn \, ds \\
  &= -\beta \int_\gamma \left(-\xx_{sss} + \sigma \xx_s 
    \right) \cdot \kappa \xx_s \, ds \\
  &= -\beta \int_\gamma \left(\xx_{ss} \cdot 
    (\kappa \xx_s)_s + \kappa \sigma \right) \, ds \\
  &= -\beta \int_\gamma \left(\kappa (\xx_{ss} \cdot \xx_{ss}) + 
    \kappa_s (\xx_{ss} \cdot \xx_s) + \kappa \sigma \right) 
    \, ds \\
  &= -\beta \int_\gamma \left(\kappa^3 + \kappa \sigma \right) 
    \, ds = \beta \int_\gamma \left(
    \frac{\kappa^3}{2} - \kappa \Lambda \right) \, ds.
  \label{eqn:areaROC}
\end{align}
Note that the integrand in~\eqref{eqn:areaROC} is zero when the vesicle
reaches its steady state which confirms that $\dot{A}(t) = 0$.

\end{document}

% OLD TEXT AND PLOTS FROM THE NUMERICAL EXAMPLES

%We plot the area of a vesicle with six different permeability rates in
%Figure~\ref{fig:ellipseArea}. The area is computed directly from the
%vesicle shape and the time axis is scaled by $\beta$. The plot verifies
%that the time for the vesicle to reach a circular shape is proportional
%to $\beta^{-1}$. We also solve~\eqref{eqn:areaROC} for the $\beta=10$
%simulation and superimpose its solution with dots on
%Figure~\ref{fig:ellipseArea}.

%\begin{figure}[htp]
%  \centering
%%  \ifTikz
%%  \input{figures/ellipseArea.tikz}
%%  \fi
%  \includegraphics[width=0.9\linewidth]{figures/ellipseArea}
%  \caption{\label{fig:ellipseArea} The area of an elliptical
%  semipermeable vesicle, with permeability rate $\beta$ in a quiescent
%  flow. The time axis is scaled by the permeability rate $\beta$. Since
%  the curves collapse to a single curve, the permeability rate $\beta$
%  sets a relaxation time scale.}
%\end{figure}

%We further analyze the transient dynamics by considering the vesicle
%velocity~\eqref{eqn:vesVelocity} for small $\beta$. Two terms contribute
%to the vesicle velocity: (i) the permeability velocity, $\beta(\ff \cdot
%\nn)\nn$, and (ii) the force balance velocity, $\SS[\ff]$. Since $\beta
%\ll 1$, the permeability velocity is much smaller than the force balance
%velocity at early times, and the semipermeable vesicle dynamics
%resembles those of an impermeable vesicle. As the force balance velocity
%decreases, the permeability velocity becomes significant, and the
%vesicle begins to inflate to a circle. This results in the membrane
%force and velocity both converging to zero. Letting $\beta = 10^{-3}$,
%the normal component of the force balance velocity (blue) and the
%permeability velocity (red) are plotted at three times in
%Figure~\ref{fig:vesVelocity}. We also plot the velocity (green) of an
%identically initialized impermeable vesicle. At the early time $t=0.5$,
%the permeability velocity is negligible relative to the force balance
%velocity, and the force balance velocity matches with the impermeable
%vesicle velocity. At intermediate time $t=5$, all three velocities are
%comparable in size, and the semipermeable vesicle begins to inflate.
%Finally, at time $t=50$, the impermeable vesicle is near steady state,
%while the semipermeable vesicle is still inflating. The right-most
%plot of Figure~\ref{fig:vesVelocity} is the maximum norm of the
%different velocity contributions as a function of time, and the black
%vertical lines correspond to the times reported in the other three
%subplots.
%
%\begin{figure*}[htp]
%  \centering
%%  \scalebox{0.6}{\input{figures/ellipseVelocity1.tikz}}
%%  \scalebox{0.6}{\input{figures/ellipseVelocity2.tikz}}
%%  \scalebox{0.6}{\input{figures/ellipseVelocity3.tikz}}
%%  \scalebox{0.6}{\input{figures/ellipseVelocity4.tikz}}
%%  \input{figures/ellipseVelocityNorms.tikz}
%  \includegraphics[width=0.24\linewidth]{figures/ellipseVelocity1.pdf}
%  \includegraphics[width=0.24\linewidth]{figures/ellipseVelocity2.pdf}
%  \includegraphics[width=0.24\linewidth]{figures/ellipseVelocity3.pdf}
%  \includegraphics[width=0.24\linewidth]{figures/ellipseVelocity4.pdf}
%  \caption{\label{fig:vesVelocity} The normal velocity contributions due
%  to force balance (blue) and permeability (red). The green line is the
%  velocity of an impermeable vesicle with the same initial shape as the
%  semipermeable vesicle.}
%\end{figure*}

%We next investigate the interplay between the tension, transmembrane
%flux, and curvature of a semipermeable vesicle. We let $\beta = 1$ and
%initialize the vesicle with regions of both positive and negative
%curvature. The vesicle tension, $\Lambda$, and the flux, $\ff \cdot
%\nn$, are plotted in Figure~\ref{fig:starTensionFlux}. At early times,
%regions of large negative curvature have negative tension and regions
%with large positive curvature have positive tension, and the same result
%holds for an impermeable vesicle. In addition, regions with large
%curvature have a large inflow of fluid (positive flux), while the
%flatter regions have an outflow of fluid. At these early times, the net
%effect is vesicle deflation. As the vesicle shape relaxes, the same
%general tension and flux trends persist, but the net flux is positive
%and the vesicle inflates. At later times, the tension becomes positive
%and the vesicle continues to inflate towards a circle. Finally, the
%tension transitions to a positive constant function, the membrane force
%and flux approach zero, and the vesicle shape reaches its steady
%circular shape.

%\begin{figure*}[htp]
%  \centering
%  \includegraphics[width=0.48\linewidth,trim =2cm 5cm 0cm 5cm, clip=true]{figures/StarTensionTime1.pdf}
%  \includegraphics[width=0.48\linewidth,trim =2cm 5cm 0cm 5cm, clip=true]{figures/StarFluxTime1.pdf}
%
%  \includegraphics[width=0.48\linewidth,trim =2cm 5cm 0cm 5cm, clip=true]{figures/StarTensionTime2.pdf}
%  \includegraphics[width=0.48\linewidth,trim =2cm 5cm 0cm 5cm, clip=true]{figures/StarFluxTime2.pdf}
%
%  \includegraphics[width=0.48\linewidth,trim =2cm 5cm 0cm 5cm, clip=true]{figures/StarTensionTime3.pdf}
%  \includegraphics[width=0.48\linewidth,trim =2cm 5cm 0cm 5cm, clip=true]{figures/StarFluxTime3.pdf}
%
%  \includegraphics[width=0.48\linewidth,trim =2cm 5cm 0cm 5cm, clip=true]{figures/StarTensionTime4.pdf}
%  \includegraphics[width=0.48\linewidth,trim =2cm 5cm 0cm 5cm, clip=true]{figures/StarFluxTime4.pdf}
%
%  \includegraphics[width=0.48\linewidth,trim =2cm 5cm 0cm 5cm, clip=true]{figures/StarTensionTime5.pdf}
%  \includegraphics[width=0.48\linewidth,trim =2cm 5cm 0cm 5cm, clip=true]{figures/StarFluxTime5.pdf}
%
%  \includegraphics[width=0.48\linewidth,trim =2cm 5cm 0cm 5cm, clip=true]{figures/StarTensionTime6.pdf}
%  \includegraphics[width=0.48\linewidth,trim =2cm 5cm 0cm 5cm, clip=true]{figures/StarFluxTime6.pdf}
%  \caption{\label{fig:starTensionFlux} The vesicle tension (left) and
%  flux (right) of the semipermeable vesicle in a quiescent flow with
%  $\beta=1$.}
%\end{figure*}

%\begin{figure*}[htp]
%  \centering
%  \includegraphics[width=0.16\linewidth,trim =2cm 5cm 0cm 5cm, clip=true]{figures/StarTensionTime1.pdf}
%  \includegraphics[width=0.16\linewidth,trim =2cm 5cm 0cm 5cm, clip=true]{figures/StarTensionTime2.pdf}
%  \includegraphics[width=0.16\linewidth,trim =2cm 5cm 0cm 5cm, clip=true]{figures/StarTensionTime3.pdf}
%  \includegraphics[width=0.16\linewidth,trim =2cm 5cm 0cm 5cm, clip=true]{figures/StarTensionTime4.pdf}
%  \includegraphics[width=0.16\linewidth,trim =2cm 5cm 0cm 5cm,clip=true]{figures/StarTensionTime5.pdf}
%  \includegraphics[width=0.16\linewidth,trim =2cm 5cm 0cm 5cm,clip=true]{figures/StarTensionTime6.pdf}
%
%  \includegraphics[width=0.16\linewidth,trim =2cm 5cm 0cm 5cm, clip=true]{figures/StarFluxTime1.pdf}
%  \includegraphics[width=0.16\linewidth,trim =2cm 5cm 0cm 5cm, clip=true]{figures/StarFluxTime2.pdf}
%  \includegraphics[width=0.16\linewidth,trim =2cm 5cm 0cm 5cm, clip=true]{figures/StarFluxTime3.pdf}
%  \includegraphics[width=0.16\linewidth,trim =2cm 5cm 0cm 5cm, clip=true]{figures/StarFluxTime4.pdf}
%  \includegraphics[width=0.16\linewidth,trim =2cm 5cm 0cm 5cm, clip=true]{figures/StarFluxTime5.pdf}
%  \includegraphics[width=0.16\linewidth,trim =2cm 5cm 0cm 5cm, clip=true]{figures/StarFluxTime6.pdf}
%  \caption{\label{fig:starTensionFlux} The vesicle tension (top) and
%  flux (bottom) of the semipermeable vesicle in a quiescent
%  flow with $\beta=1$.}
%\end{figure*}

%\begin{figure}[htp]
%  \centering
%  \ifTikz
%  \input{figures/shearSS.tikz}
%  \fi
%%  \includegraphics[width=0.9\linewidth]{figures/shearSS.pdf}
%  \caption{\label{fig:shearSS} The reduced area of a semipermeable
%  vesicle in a shear flow. The final reduced area is nearly independent
%  of the initial reduced area.}
%\end{figure}

%\begin{figure}[htp]
%  \centering
%  %\ifTikz
%  \input{figures/shearPhaseDiagramRA2.tikz}
%  %\fi
%%  \includegraphics[width=\linewidth]{figures/shearPhaseDiagramRA2.pdf}
%  \caption{\label{fig:shearPhaseDiagram} The steady state shape of a
%  semipermeable vesicle with several different permeability rates
%  $\beta$ and shear rates $\dot{\gamma}$. The final reduced areas are
%  reported for the largest and smallest parameter values. The final
%  reduced area decreases with the flow rate, while the permeability rate
%  has a very small effect on the steady state shape.}
%\end{figure}

%\begin{figure}[htp]
%%  \ifTikz
%%  \input{figures/parabolicOffCenterPhaseImpermeable.tikz}
%%  \fi
%  \includegraphics[width=0.9\linewidth]{figures/parabolicOffCenterPhaseImpermeable.pdf}
%  \caption{\label{fig:parabolicOffCenterImpermeablePhaseDiagram} Phase
%  diagram of an impermeable vesicle with several different flow rates
%  $U$ and reduced areas $\nu$. As shown by Kaoui et
%  al.~\cite{kao-bir-mis2009}, the final vesicle shape is either an
%  axisymmetric parachute or bullet (pink region) or an asymmetric
%  slipper (yellow region).}
%\end{figure}

%\begin{figure}[htp]
%%  \ifTikz
%%  \input{figures/parabolicOffCenterPhaseSemipermeable.tikz}
%%  \fi
%  \includegraphics[width=0.9\linewidth]{figures/parabolicOffCenterPhaseSemipermeable.pdf}
%  \caption{\label{fig:parabolicOffCenterSemipermeablePhaseDiagram} Phase
%  diagram of a semipermeable vesicle, with $\beta = 10^{-3}$, with
%  several different flow rates $U$ and reduced areas $\nu$. The
%  steady state reduced area for each flow rate, which is reported along
%  the top of the vesicles, is nearly independent of the initial reduced
%  area. Low flow rates result in axisymmetric vesicles with large
%  reduced area, and high flow rates result in asymmetric vesicles with
%  low reduced area. Results at higher flow rates are not eporeted since
%  the vesicle self-intersects.}
%\end{figure}

%\begin{figure}[htp]
%%  \ifTikz
%%  \scalebox{0.5}{\input{figures/stenosisShapeTension01.tikz}}
%%  \scalebox{0.5}{\input{figures/stenosisShapeFlux02.tikz}}
%%  \scalebox{0.5}{\input{figures/stenosisShapeTension03.tikz}}
%%  \scalebox{0.5}{\input{figures/stenosisShapeFlux04.tikz}}
%%  \scalebox{0.5}{\input{figures/stenosisShapeTension05.tikz}}
%%  \scalebox{0.5}{\input{figures/stenosisShapeFlux06.tikz}}
%%  \scalebox{0.5}{\input{figures/stenosisShapeTension07.tikz}}
%%  \scalebox{0.5}{\input{figures/stenosisShapeFlux08.tikz}}
%%  \scalebox{0.5}{\input{figures/stenosisShapeTension09.tikz}}
%%  \scalebox{0.5}{\input{figures/stenosisShapeFlux10.tikz}}
%%  \scalebox{0.5}{\input{figures/stenosisShapeTension11.tikz}}
%%  \scalebox{0.5}{\input{figures/stenosisShapeFlux12.tikz}}
%%  \scalebox{0.5}{\input{figures/stenosisShapeTension02.tikz}}
%%  \scalebox{0.5}{\input{figures/stenosisShapeFlux04.tikz}}
%%  \scalebox{0.5}{\input{figures/stenosisShapeTension06.tikz}}
%%  \scalebox{0.5}{\input{figures/stenosisShapeFlux08.tikz}}
%%  \scalebox{0.5}{\input{figures/stenosisShapeTension10.tikz}}
%%  \scalebox{0.5}{\input{figures/stenosisShapeFlux12.tikz}}
%%  \fi
%  \includegraphics[width=0.45\linewidth]{figures/stenosisShapeTension02.pdf}
%  \includegraphics[width=0.45\linewidth]{figures/stenosisShapeFlux04.pdf}
%  \\
%  \includegraphics[width=0.45\linewidth]{figures/stenosisShapeTension06.pdf}
%  \includegraphics[width=0.45\linewidth]{figures/stenosisShapeFlux08.pdf}
%  \\
%  \includegraphics[width=0.45\linewidth]{figures/stenosisShapeTension10.pdf}
%  \includegraphics[width=0.45\linewidth]{figures/stenosisShapeFlux12.pdf}
%  \caption{\label{fig:stenosisShape} An impermeable (dashed) and
%  semipermeable (solid) vesicle in a stenosed geometry (black).
%  The vesicles' initial configuration are identical with reduced area
%  $\nu = 0.65$. The color denotes the vesicles' tension (left) and flux
%  (right). The semipermeable vesicle deflates throughout the
%  constriction (see Figure~\ref{fig:stenosisArea}) and develops higher
%  curvature regions than its impermeable counterpart.}
%\end{figure}
%
%\begin{figure}[htp]
%  \centering
%%  \ifTikz
%%  \scalebox{0.5}{\input{figures/stenosisTension12.tikz}}
%%  \scalebox{0.5}{\input{figures/stenosisFlux12.tikz}}
%%  \scalebox{0.5}{\input{figures/stenosisXvelocity12.tikz}}
%%  \scalebox{0.5}{\input{figures/stenosisYvelocity12.tikz}}
%%  \fi
%  \includegraphics[width=0.45\linewidth]{figures/stenosisTension12.pdf}
%  \includegraphics[width=0.45\linewidth]{figures/stenosisFlux12.pdf} \\
%  \includegraphics[width=0.45\linewidth]{figures/stenosisXvelocity12.pdf}
%  \includegraphics[width=0.45\linewidth]{figures/stenosisYvelocity12.pdf}
%  \caption{\label{fig:stenosisTensionAndFlux} The tension (top left), 
%  flux (top right), $x$-velocity (bottom left), and $y$-velocity (bottom
%  right) of the semipermeable (red) and impermeable (dashed blue)
%  vesicles in Figure~\ref{fig:stenosisShape} at location $x=50$. The
%  point $\theta=0$ corresponds to the right-most point of the vesicle.}
%\end{figure}
%
%\begin{figure}[htp]
%  \centering
%%  \ifTikz
%%  \input{figures/stenosisArea.tikz}
%%  \fi
%  \includegraphics[width=0.8\linewidth]{figures/stenosisArea}
%  \caption{\label{fig:stenosisArea} The area of a stenosed
%  semipermeable vesicle, with permeability rate $\beta=10^{-3}$. The
%  $x$-axis is the $x$-coordinate of the center of mass of the vesicle.
%  The vesicle loses fluid at a constant rate throughout the constriction
%  that is between $\pm 60$. The semipermeable vesicle loses 12.8\% of
%  its fluid.}
%\end{figure}
%
%
%\begin{figure}[htp]
%%  \ifTikz
%%  \input{figures/stenosisBL.tikz}
%%  \fi
%  \includegraphics[width=\linewidth]{figures/stenosisBL.pdf}
%  \caption{\label{fig:stenosisBL} A stenosed semiimpermeable (red) and
%  imperemable (blue) vesicle. The semipermeable vesicle has 
%  reduced area 0.17, and both vesicles were initialized with reduce area
%  $0.20$. The bottom plot shows the velocity relative to the vesicle
%  inside the lubrication layer.  The vesicle velocity in the square is
%  nearly constant (see Figure~\ref{fig:stenosisRA020Velocity}.}
%\end{figure}
%
%
%\begin{figure}[htp]
%  \centering
%%  \ifTikz
%%  \scalebox{0.5}{\input{figures/stenosisXvelocityRA020.tikz}}
%%  \scalebox{0.5}{\input{figures/stenosisYvelocityRA020.tikz}}
%%  \fi
%  \includegraphics[width=0.45\linewidth]{figures/stenosisXvelocityRA020.pdf}
%  \includegraphics[width=0.45\linewidth]{figures/stenosisYvelocityRA020.pdf}
%  \caption{\label{fig:stenosisRA020Velocity} The $x$ (left) and $y$
%  (right) velocities of the vesicle in Figure~\ref{fig:stenosisBL}. The
%  red lines correspond to the semipermeable vesicle and the blue dashed
%  lines correspond to the impermeable vesicle.}
%\end{figure}
%
%We consider the effect of stenosis on a semipermeable vesicle. To allow
%for large amounts of fluid flux through the membrane, we consider
%geometry with a long constricted region. We start by considering a
%vesicle initialized with reduced area $\nu = 0.65$, and we compare the
%vesicle shape when $\beta = 0$ (impermeable) and $\beta = 10^{-3}$
%(permeable). Figure~\ref{fig:stenosisShape} shows the vesicles' shape at
%several locations along the constriction. The colors along the left
%channel represent the tension and those along the right represent the
%flux. In Figure~\ref{fig:stenosisTensionAndFlux}, the tension at
%location $x=20$ and the flux at location $x=30$ are plotted with respect
%to arclength. The value of $\theta=0$ corresponds to the nose of the
%vesicle.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\subsection*{Dimensionless equations}
%The inextensibility condition guarantees that the vesicle's length, $L$,
%is conserved, while its area, $A$, can change due to semipermeability.
%Therefore, the reduced area $\alpha = 4\pi A/L^2 \in (0,1]$ is not
%conserved. We non-dimensionalize~\eqref{eqn:vesVelocity} by introducing
%the length scale $R_0 = L/2\pi$ and defining the time scale $\tau = \mu
%R_0^3/k_b$, the tension scale $Q = k_b/R_0^2$, and the permeability
%scale $W = R_0/\mu$. The background velocity $\uu_\infty$ is
%non-dimensionalized with $R_0/\tau$.
%
%In each experiment, we initialize an elliptical vesicle with longest
%diameter of 9$\mu$m. We use a length scale of $R_0 = 10^{-6}$m, a
%bending stiffness of $k_b = 10^{-19}$J, and fluid viscosity of $\mu=5
%\times 10^{-2}$kg/ms. Using the relaxation time scale $\tau = \mu
%R_0^3/k_b = 5 \times 10^{-1}$s, the pressure scale is $P = k_b/R_0^3 =
%10^{-1}$Pa. 
% \newpage
% \paragraph{\bf \Huge OLD STUFF}
%
%%%%%%%%%%%%%%%
%\section{{$\Theta-L$} formulation}
%Here, we attempt to apply the formulation described by Sohn et
%al.~\cite{soh-tse-li-voi-low2010} to a semipermeable vesicle.  In the
%semipermeable formulation, the vesicle velocity is
%\begin{align}
%  \pderiv{\XX}{t} = \uu(\XX(t)) + \beta (\ff \cdot \nn) \nn.
%\end{align}
%In the {$\Theta-L$} variables, we require a normal velocity $V$, and a
%tangential velocity $T$. Since the inextensibility condition is $T_s +
%\kappa V = 0$, a condition for the vesicle velocity is
%\begin{align}
%  (\uu \cdot \ss)_s = -\kappa(\beta (\ff \cdot \nn) + 
%      \uu \cdot \nn).
%\end{align}
%This equation can be solved following equations (32)--(39)
%in~\cite{soh-tse-li-voi-low2010}, with the only difference being that
%$\tilde{\uu}$ satisfies
%\begin{align}
%  (\tilde{\uu} \cdot \ss)_s + \kappa(\tilde{\uu} \cdot \nn) =
%    -(\vv \cdot \ss)_s - \kappa(\vv \cdot \nn) 
%    - \beta\kappa(\ff \cdot \nn).
%\end{align}
%Compared to equation (40), only the last term is new.
%
%
%%%%%%%%%%%%%%%
%Finally, we write the time stepping method in matrix form as
%\begin{align}
%  A \left(
%    \begin{array}{c}
%      \xx^{N+1} \\ \sigma^{N+1}
%    \end{array}
%  \right) = 
%  \left(
%    \begin{array}{c}
%      \xx^{N} + \Delta t \uu_\infty(\xx^N) \\ \mathbf{1}
%    \end{array}
%  \right),
%\end{align}
%where
%\begin{align}
% A = \left(
%  \begin{array}{cc}
%    I + \Delta t \beta P^N B^N + \Delta t \SS^N B^N & 
%    -\Delta t P^N T^N - \Delta t \SS^N T^N \\
%    D^N & 0
%  \end{array}
%  \right).
%\end{align}
%
%\section{Area of a Vesicle}
%The area inside a single vesicle $\gamma$ parameterized with $\xx$ is
%\begin{align}
%  A(t) = \int_{\gamma} (\xx \cdot \nn)\, ds = 
%    \int_{0}^{2\pi} \left(\xx \cdot \pderiv{\xx}{\theta}^\perp\right)
%    \, d\theta.
%\end{align}
%Taking the time derivative of $A(t)$,
%\begin{align}
%  \dot{A}(t) &= \int_{0}^{2\pi} \left(
%    \dot{\xx} \cdot \pderiv{\xx}{\theta}^{\perp} + 
%    \xx \cdot \pderiv{\dot{\xx}}{\theta}^{\perp}\right) \\
%    \, d\theta.
%  &= \int_{0}^{2\pi} \left(
%    (\dot{\xx} \cdot \nn)\|\xx'(\theta)\| \right) d\theta +
%     \int_{0}^{2\pi}
%    \xx \cdot \frac{d}{dt}\left(\nn |\frac{d\xx}{d\theta}| \right)
%\end{align}
%
%
%
%\begin{align}
%  A(t) = \frac{1}{2}\int_{\gamma} (\xx \cdot \nn)\, ds = 
%    \frac{1}{2}\int_{0}^{2\pi} \left(\xx \cdot \pderiv{\xx}{\theta}^\perp\right)
%    \, d\theta.
%\end{align}
%Taking the time derivative of $A(t)$,
%\begin{align}
%  \dot{A}(t) &= \frac{1}{2}\int_{0}^{2\pi} \left(\dot{\xx} \cdot
%  \pderiv{\xx}{\theta}^{\perp} + 
%    \xx \cdot \pderiv{\dot{\xx}}{\theta}^{\perp}\right) d\theta, \\
%  &= \frac{1}{2}
%  \int_{0}^{2\pi}(\dot{\xx}\cdot\nn)\left\|\pderiv{\xx}{\theta}\right\|d\theta +
%     \frac{1}{2}\int_{0}^{2\pi}
%    \left(\xx \cdot \pderiv{\dot{\xx}}{\theta}^{\perp}\right)d\theta, \\
%  &= \frac{\beta}{2} \int_{\gamma}\left(\ff\cdot\nn\right)ds +
%  \frac{1}{2}\int_{0}^{2\pi}\left(\xx \cdot \pderiv{}{t}\left(\nn
%  \left\|\pderiv{\xx}{\theta}\right\|\right)\right)d\theta, \\
%  &= \frac{\beta}{2} \int_{\gamma}\left(\ff\cdot\nn\right)ds + \frac{1}{2}
%  \int_{\gamma}\left(\xx\cdot\dot{\nn}\right)ds + \frac{1}{2}\int_{0}^{2\pi}
%  \frac{\xx\cdot\nn}{\left\|\pderiv{\xx}{\theta}\right\|}
%  \left(\pderiv{\xx}{\theta}\cdot
%  \pderiv{\dot{\xx}}{\theta}\right)d\theta \\
%  &= \frac{\beta}{2} \int_{\gamma}\left(\ff\cdot\nn\right)ds + \frac{1}{2}
%  \int_{\gamma}\left(\xx\cdot\dot{\nn}\right)ds + \frac{1}{2}\int_{\gamma}
%  (\xx\cdot\nn) (\xx_s \cdot \dot{\xx}_s)\, ds.
%\end{align}
%To satisfy the inextensibilty condition, $\xx_s \cdot \dot{\xx}_s = 0$.
%Therefore the time derivative of $A(t)$ is
%\begin{align}
%     \dot{A}(t) &= \frac{\beta}{2} \int_{\gamma}\left(\ff\cdot\nn\right)ds
%     + \frac{1}{2} \int_{\gamma}\left(\xx\cdot\dot{\nn}\right)ds.
%\end{align}
%which we re-write as
%\begin{align}
%    \dot{A}(t) &= \frac{\beta}{2}
%    \int_{\gamma}\left(\ff\cdot\nn\right)ds-\frac{1}{2}\int_{\gamma}\left(\xx^{\perp}\cdot\dot{\xx}_s\right) ds.
%\end{align}
%Applying integration by parts,
%\begin{align}
%    \dot{A}(t) &= \frac{\beta}{2}
%    \int_{\gamma}\left(\ff\cdot\nn\right)ds+\frac{1}{2}\int_{\gamma}\left(\xx^{\perp}_s\cdot\dot{\xx}\right) ds,
%\end{align}
%which we re-write as
%\begin{align}
%    \dot{A}(t) &= \frac{\beta}{2}
%    \int_{\gamma}\left(\ff\cdot\nn\right)ds+\frac{1}{2}\int_{\gamma}\left(\dot{\xx}\cdot\nn\right),\\
%    \dot{A}(t) &= \beta \int_{\gamma}\left(\ff\cdot\nn\right)ds.
%\end{align}
%
%
%At steady state,
%\begin{align}
%    \dot{A}(t) &= \beta \int_{\gamma}\left(\ff\cdot\nn\right)ds =0.
%\end{align}
%We now replace $\ff$ with its definition and apply integration by parts
%\begin{align}
%   0=\beta\int_{\gamma} 
%    (-k_b\xx_{ssss}+(\sigma\xx_s)_s)\cdot\nn\, ds =
%   \beta\int_{\gamma}\left(k_b\xx_{sss}-\sigma\xx_s\right)\cdot
%     \nn_s\, ds
%\end{align}
%From the Frenet-Serret formulas, we have 
%\begin{align}
%     -\beta\int_{\gamma}
%     \left(k_b\xx_{sss}-\sigma\xx_s\right)\cdot\kappa\xx_s ds &=0.
%\end{align}
%
%Applying integration by parts and Frenet-Serret again
%\begin{align}
%     \beta\int_{\gamma}
%     k_b\kappa\left(\xx_{ss}\cdot\xx_{ss}\right)ds + \beta
%     \int_{\gamma} k_b \kappa_s(\xx_{ss} \cdot \xx_s)ds +\beta\int_{\gamma}\kappa\sigma ds  &= 0,\\
%    \beta\int_{\gamma} k_b\kappa(\kappa\nn \cdot
%     \kappa\nn)+ \beta \int_{\gamma} k_b \kappa_s(\xx_{ss} \cdot
%     \xx_s)ds +\beta\int_{\gamma}\kappa\sigma ds &= 0,
%\end{align}
%which we re-write as
%\begin{align}
%   \beta\int_{\gamma}k_b\kappa^3 ds + \beta \int_{\gamma} k_b
%   \kappa_s(\xx_{ss} \cdot \xx_s)ds +\beta\int_{\gamma} \kappa\sigma ds = 0.
%\end{align}
%Applying Frenet-serret again,
%\begin{align}
%    \beta\int_{\gamma}k_b\kappa^3 ds + \beta \int_{\gamma} k_b
%    \kappa_s(\kappa\nn \cdot \TT)ds +\beta\int_{\gamma} \kappa\sigma ds = 0, \\
%    \beta\int_{\gamma}k_b\kappa^3 ds +\beta\int_{\gamma} \kappa\sigma ds  = 0.
%\end{align}
%
%
%\section{Including a concentration field}
%Based on the work of Yao and Mori~\cite{yao-mor2017}, we introduce a
%chemical concentration field $c$.  The concentration field is governed
%by the advection-diffusion equation
%\begin{align}
%  \pderiv{c}{t} + \nabla \cdot (\uu c) = D\Delta c.
%\end{align}
%If the concentration is diffusion dominated (zero Peclet number), then
%the concentration satisfies
%\begin{align}
%  \Delta c = 0, \quad \xx \in \RR^2.
%\end{align}
%The boundary conditions of the concentration field on the vesicle
%$\gamma$ are
%\begin{align}
%  (\uu c - D\nabla c) \cdot \nn = c \pderiv{\xx}{t} \cdot \nn + 
%    j_c + j_p,
%\end{align}
%where $j_c$ and $j_p$ are the transmembrane chemical flux due to passive
%transport and active transport (pumps), respectively.  In the zero
%Peclet number limit, the absence of an active transport mechanism, and
%the choice for the passive flux, the boundary condition of the
%concentration field is
%\begin{align}
%  -D\pderiv{c}{\nn} = c \pderiv{\xx}{t} + 
%    k_c \jump{c}, \quad \xx \in \gamma.
%\end{align}
%To couple the vesicle dynamics to the concentration, there is a
%difference between the fluid velocity and vesicle velocity
%\begin{align}
%  \uu - \pderiv{\xx}{t} = j_w \nn,
%\end{align}
%where $j_w$ is the water flux through the vesicle.  This is the
%osmophoretic flow.  In~\cite{yao-mor2017}, this flux is
%\begin{align}
%  j_w = -k_w(RT\jump{c} + \ff_\text{mem} \cdot \nn).
%\end{align}
%In our current implementation, we drop the $RT\jump{c}$ term, and
%let $\beta = k_w$.
%
%\section{Preliminary results}
%
%
%\begin{figure}
%	\centering
%	\includegraphics[width=.9\textwidth]{figures/1.jpg}
%	\includegraphics[width=.9\textwidth]{figures/2.jpg}
%	\includegraphics[width=.9\textwidth]{figures/3.jpg}
%	\caption{Reduced area as a function of time for $\beta = [0, 10^{-2},10^{-1}, 1]$ with a shear flow rate $\chi = 0$.}
%%	\label{Shear0}
%\end{figure}
%
%\begin{figure}
%	\centering
%	\includegraphics[width=.9\textwidth]{figures/4.jpg}
%	\includegraphics[width=.9\textwidth]{figures/5.jpg}
%	\includegraphics[width=.9\textwidth]{figures/6.jpg}
%	\caption{Reduced area as a function of time for $\beta = [0, 10^{-2},10^{-1}, 1]$ with a shear flow rate $\chi = 0.5$.}
%%	\label{Shear0}
%\end{figure}
%
%\begin{figure}
%	\centering
%	\includegraphics[width=.9\textwidth]{figures/7.jpg}
%	\includegraphics[width=.9\textwidth]{figures/8.jpg}
%	\includegraphics[width=.9\textwidth]{figures/9.jpg}
%	\caption{Reduced area as a function of time for $\beta = [0, 10^{-2},10^{-1}, 1]$ with a shear flow rate $\chi = 1$.}
%%	\label{Shear0}
%\end{figure}
%
%\begin{figure}
%	\centering
%	\includegraphics[width=.9\textwidth]{figures/10.jpg}
%	\includegraphics[width=.9\textwidth]{figures/11.jpg}
%	\includegraphics[width=.9\textwidth]{figures/12.jpg}
%	\caption{Reduced area as a function of time for $\beta = [0, 10^{-2},10^{-1}, 1]$ with a shear flow rate $\chi = 2$.}
%%	\label{Shear0}
%\end{figure}
% 
% 
% \begin{figure}
% 	\centering
% 	\includegraphics[width=.9\textwidth]{figures/BE1.jpg}
% 	\includegraphics[width=.9\textwidth]{figures/BE2.jpg}
% 	\includegraphics[width=.9\textwidth]{figures/BE3.jpg}
% 	\caption{Bending energy as a function of time for $\beta = [0, 10^{-2},10^{-1}, 1]$ with a shear flow rate $\chi = 0$.}
%% 	\label{Shear0}
% \end{figure}
% 
% \begin{figure}
% 	\centering
% 	\includegraphics[width=.9\textwidth]{figures/BE4.jpg}
% 	\includegraphics[width=.9\textwidth]{figures/BE5.jpg}
% 	\includegraphics[width=.9\textwidth]{figures/BE6.jpg}
% 	\caption{Bending energy as a function of time for $\beta = [0, 10^{-2},10^{-1}, 1]$ with a shear flow rate $\chi = 0.5$.}
%% 	\label{Shear0}
% \end{figure}
% 
% \begin{figure}
% 	\centering
% 	\includegraphics[width=.9\textwidth]{figures/BE7.jpg}
% 	\includegraphics[width=.9\textwidth]{figures/BE8.jpg}
% 	\includegraphics[width=.9\textwidth]{figures/BE9.jpg}
% 	\caption{Bending energy as a function of time for $\beta = [0, 10^{-2},10^{-1}, 1]$ with a shear flow rate $\chi = 1$.}
%% 	\label{Shear0}
% \end{figure}
% 
% \begin{figure}
% 	\centering
% 	\includegraphics[width=.9\textwidth]{figures/BE10.jpg}
% 	\includegraphics[width=.9\textwidth]{figures/BE11.jpg}
% 	\includegraphics[width=.9\textwidth]{figures/BE12.jpg}
% 	\caption{Bending energy as a function of time for $\beta = [0, 10^{-2},10^{-1}, 1]$ with a shear flow rate $\chi = 2$.}
% 	\label{Shear0}
% \end{figure}
%
%When the semipermeable vesicle reaches a steady-state shape, we have
%\begin{align}
%  \int_{\gamma} \left(-\frac{\kappa^3}{2} + \kappa \Lambda \right) \, ds
%  = 0.
%\end{align}
%Based on numerical observations, the steady-state shape is a circular
%vesicle with constant tension. Since the tension and curvature are
%constant, we can solve for the curvature in terms of the tension $\kappa
%= \sqrt{2\Lambda}$. Therefore, the steady state shape is a circle of
%radius $r = 1/\sqrt{2\Lambda}$, and therefore the length is
%\begin{align}
%  L = \frac{\sqrt{2}\pi}{\sqrt{\Lambda}},
%\end{align}
%which is constant by the local inextensibility constraint. Finally, we
%can compute the final tension in terms of the length using the formula
%\begin{align}
%  \Lambda = \frac{2\pi^2}{L^2}.
%\end{align}
%Therefore, in contrast to a impermeable vesicle, the final tension is
%positive rather than negative.
%
%We start by simulating a star-shaped vesicle in a quiescent flow with
%semipermeability rate $\beta=1$ (Figure~\ref{fig:starShape}). To
%perform this simulation, we use an adaptive second-order time stepping
%method~\cite{qua-bir2016}. The times and reduced areas are in the
%titles, and note that the shapes are not eqiuspaced in time.
%\begin{figure}[htp]
%  \begin{minipage}{0.55\textwidth}
%%    \ifTikz
%    \input{figures/starRelaxation1.tikz}
%    \input{figures/starRelaxation2.tikz}
%    \input{figures/starRelaxation3.tikz}
%    \input{figures/starRelaxation4.tikz}
%    \input{figures/starRelaxation5.tikz}
%    \input{figures/starRelaxation6.tikz}
%%    \fi
%  \end{minipage}
%  \hfill
%  \begin{minipage}{0.4\textwidth}
%%  \ifTikz
%  \input{figures/starAreaDeriv.tikz}
%%  \fi
%  \end{minipage}
%%  \caption{\label{fig:starShape} A initially star-shaped semipermeable
%%  vesicle in a shear quiescent flow with $\beta=1$. Note that the time
%%  steps are not equispaced. The area (red) and its derivative (blue) of
%%  an initially star-shaped vesicle as a function of time. The dots
%%  correspond to the time steps shown in six time steps
%%  show on the left.}
%\end{figure}
%
%\begin{table}[htp]
%  \centering
%  \begin{tabular}{|C{1cm}|*{6}{C{2cm}}|}
%    \hline
%    $\beta$ & $10^{0}$ & $10^{-1}$ & $10^{-2}$ & 
%             $10^{-3}$ & $10^{-4}$ & $10^{-5}$ \\
%    $t_\mathrm{SS}$ & $6.45 \times 10^{0}$ & $2.00 \times 10^{1}$ & 
%                      $9.61 \times 10^{1}$ & $8.36 \times 10^{2}$ & 
%                      $8.15 \times 10^{3}$ & $8.03 \times 10^{4}$ \\
%    ratio & --- & 3.10 & 4.81 & 8.71 & 9.74 & 9.85 \\
%    \hline
%  \end{tabular}
%  \caption{\label{tbl:ellipseRelaxTime} The time, $t_{\mathrm{SS}}$, for
%  a semipermeable vesicle to reach steady state.}
%\end{table}
%
%\begin{table}[htp]
%  \centering
%  \begin{tabular}{|C{1cm}|*{6}{C{2cm}}|}
%    \hline
%    $\beta$ & $10^{0}$ & $10^{-1}$ & $10^{-2}$ & 
%             $10^{-3}$ & $10^{-4}$ & $10^{-5}$ \\
%    $t_\mathrm{SP}$ & $2.17 \times 10^{-2}$ & $1.25 \times 10^{-1}$ & 
%                      $1.59 \times 10^{0}$ & $2.19 \times 10^{1}$ & 
%                      $2.14 \times 10^{2}$ & $2.20 \times 10^{3}$ \\
%    ratio & --- & 5.76 & 12.72 & 13.77 & 9.77 & 10.14 \\
%    \hline
%  \end{tabular}
%  \caption{\label{tbl:ellipseBLTime} The time, $t_\mathrm{SP}$, for a
%  semipermeable vesicle to begin inflating.}
%\end{table}
%
%%\begin{figure}[htp]
%%\begin{minipage}{0.40\textwidth}
%%\ifTikz
%%\input{figures/starTension.tikz}
%%\fi
%%\end{minipage}
%%\hfill
%%\begin{minipage}{0.55\textwidth}
%%  \centering
%%\includegraphics[width=0.32\textwidth]{figures/StarTensionTime1.pdf}
%%\includegraphics[width=0.32\textwidth]{figures/StarTensionTime2.pdf}
%%\includegraphics[width=0.32\textwidth]{figures/StarTensionTime3.pdf} \\
%%\includegraphics[width=0.32\textwidth]{figures/StarTensionTime4.pdf}
%%\includegraphics[width=0.32\textwidth]{figures/StarTensionTime5.pdf}
%%\includegraphics[width=0.32\textwidth]{figures/StarTensionTime6.pdf}
%%\end{minipage}
%%  \caption{\label{fig:starTension} The tension of the time steps shown
%%  in Figure~\ref{fig:starShape}. The tension at the time horizon $T=1$
%%  is also included. At this point, the vesicle is nearly circular and
%%  the tension is constant and positive. The constant is given by
%%  equation~\eqref{eqn:SSshape}, and this value corresponds to the
%%  dashed black line.}
%%\end{figure}
%
%%\begin{figure}[htp]
%%\begin{minipage}{0.40\textwidth}
%%%\ifTikz
%%\input{figures/starFlux.tikz}
%%%\fi
%%\end{minipage}
%%\hfill
%%\begin{minipage}{0.55\textwidth}
%%\includegraphics[width=0.32\textwidth]{figures/StarFluxTime1.pdf}
%%\includegraphics[width=0.32\textwidth]{figures/StarFluxTime2.pdf}
%%\includegraphics[width=0.32\textwidth]{figures/StarFluxTime3.pdf} \\
%%\includegraphics[width=0.32\textwidth]{figures/StarFluxTime4.pdf}
%%\includegraphics[width=0.32\textwidth]{figures/StarFluxTime5.pdf}
%%\includegraphics[width=0.32\textwidth]{figures/StarFluxTime6.pdf}
%%\end{minipage}
%%  \caption{\label{fig:starFlux} The flux of the time steps shown in
%%  Figure~\ref{fig:starShape}. The tension at the time horizon $T=1$ is
%%  also included. At this point, the vesicle is nearly circular and there
%%  is no flux.}
%%\end{figure}
%\begin{figure}[htp]
%    \ifTikz
%    \input{figures/starRelaxation1.tikz}
%    \quad
%    \input{figures/starRelaxation2.tikz}
%    \quad
%    \input{figures/starRelaxation3.tikz}
%    \quad
%    \input{figures/starRelaxation4.tikz}
%    \quad
%    \input{figures/starRelaxation5.tikz}
%    \fi
%  \caption{\label{fig:starShape} An initially star-shaped semipermeable
%  vesicle in a quiescent flow with $\beta=1$. Note that the time steps
%  are not equispaced.}
%\end{figure}
%\begin{figure}[htp]
%  \centering
%%  \ifTikz
%  \input{figures/ellipseVelocity1.tikz}
%  \input{figures/ellipseVelocity2.tikz}
%  \input{figures/ellipseVelocity3.tikz}
%%  \fi
%  \caption{\label{fig:vesVelocity} The velocity contributions due to
%  semipermeability (red) and force balance (blue). The green dashed
%  line is the velocity of an impermeable vesicle with the same initial
%  shape as the semipermeable vesicle.}
%\end{figure}
%
%\begin{figure}[htp]
%%  \ifTikz
%  \input{figures/ellipseVelocityNorms.tikz}
%%  \fi
%  \caption{\label{fig:vesVelocityNorm} The max norm of the velocity due
%  to the force balance (blue) and semipermeability. Also included is
%  the velocity of an impermeable vesicle (green). The three black lines
%  correspond to the times reported in Figure~\ref{fig:vesVelocity}.}
%\end{figure}

% \newpage
%\input{figures/shearPhaseDiagramRA2.tikz}
%\input{figures/parabolicPhaseDiagram.tikz}
%\begin{figure}[htp]
%%  \ifTikz
%%  \input{figures/parabolicPhaseDiagram.tikz}
%%  \fi
%  \includegraphics[width=0.9\linewidth]{figures/parabolicPhaseDiagram.pdf}
%  \caption{\label{fig:parabolicPhaseDiagram} The steady state shape of a
%  semipermeable vesicle with several different permeability rates
%  $\beta$ and flow rates $U$. The numbers reported above each vesicle is
%  its reduced area. At high flow rates, the vesicle self-intersects
%  before reaching a steady state configuration.}
%\end{figure}
%
%
%\begin{table}[htp]
%  \centering
%  \caption{\label{tbl:parabolicRelaxTimes1} The time for a
%  semipermeable vesicle in a Poiseuille flow to reach its steady state
%  ($t_\mathrm{steady}$) and to begin inflating ($t_\mathrm{inflate}$).
%  The times are reported for the maximum velocities $0.1 \mu m/s$ and
%  $10 \mu m/s$.}
%  \begin{tabular}{C{0.8cm}C{2cm}C{0.8cm}C{2cm}C{0.8cm}}
%    \multicolumn{5}{c}{$U = 0.1 \mu m/s$} \\
%    $\beta$ & $t_\mathrm{steady}$ & ratio & $t_\mathrm{inflate}$ & ratio \\
%    \midrule
%    $10^{0}$  & $3.33 \times 10^{0}$ & ---  
%              & $1.67 \times 10^{-2}$ & --- \\ 
%    $10^{-1}$ & $1.37 \times 10^{1}$ & 4.12 & 
%                $8.73 \times 10^{-1}$ & 5.23 \\
%    $10^{-2}$ & $8.60 \times 10^{1}$ & 6.26 & 
%                $1.11 \times 10^{0}$ & 12.74 \\
%    $10^{-3}$ & $7.75 \times 10^{2}$ & 9.01 & 
%                $1.63 \times 10^{1}$ & 14.70 \\
%    $10^{-4}$ & $7.69 \times 10^{3}$ & 9.93 & 
%                $1.74 \times 10^{2}$ & 10.63 \\
%    $10^{-5}$ & $8.01 \times 10^{4}$ & 10.41 & 
%                $1.66 \times 10^{3}$ & 9.58 \\
%    \bottomrule
%  \end{tabular}
%
%  \begin{tabular}{C{0.8cm}C{2cm}C{0.8cm}C{2cm}C{0.8cm}}
%    \multicolumn{5}{c}{$U = 10 \mu m/s$} \\
%    $\beta$ & $t_\mathrm{steady}$ & ratio & $t_\mathrm{inflate}$ & ratio \\
%    \midrule
%    $10^{0}$  & $3.55 \times 10^{0}$ & ---  
%              & $1.67 \times 10^{-2}$ & --- \\ 
%    $10^{-1}$ & $1.87 \times 10^{1}$ & 5.27 & 
%                $9.01 \times 10^{-2}$ & 5.40 \\
%    $10^{-2}$ & $1.17 \times 10^{2}$ & 6.26 & 
%                $1.52 \times 10^{0}$ & 16.87 \\
%    $10^{-3}$ & $1.06 \times 10^{3}$ & 9.06 & 
%                $2.07 \times 10^{1}$ & 13.62 \\
%    $10^{-4}$ & $1.02 \times 10^{4}$ & 9.62 & 
%                $2.25 \times 10^{2}$ & 10.87 \\
%    $10^{-5}$ & $1.04 \times 10^{5}$ & 10.20 & 
%                $2.25 \times 10^{3}$ & 10.00 \\
%    \bottomrule
%  \end{tabular}
%\end{table}
%The width $W$ is chosen so that $W/R =
%10$, where $R$ is the radius of the vesicle. We investigate the effect
%of the maximum velocity $U$ and the permeability rate $\beta$ on the
%steady state shape of a vesicle initially placed on the center line
%$y=0$. Figure~\ref{fig:parabolicPhaseDiagram} shows the steady state
%configuration for a variety of maximum velocities and permeability
%rates. Similar to the shear example, the maximum velocity strongly
%affects the final shape while the permeability rate's effect is much
%weaker. Instead, the permeability rate determines the time required for
%the vesicle to reach a steady state. In
%Table~\ref{tbl:parabolicRelaxTimes1}, we consider two maximum velocities
%and report the time, $t_\mathrm{steady}$, required for the vesicle's
%area to be within 1.3\% of its maximum area, and the time,
%$t_\mathrm{inflate}$, required for the vesicle's area to exceed 1.3\% of
%its initial area. We see that both times scale with $\beta^{-1}$.

%\begin{figure}[htp]
%  \centering
%  \begin{tabular}{ccc}
%    \ifTikz
%    \scalebox{0.31}{\input{figures/stenosisTension1.tikz}} &
%    \scalebox{0.31}{\input{figures/stenosisTension3.tikz}} &
%    \scalebox{0.31}{\input{figures/stenosisTension5.tikz}} \\
%    \scalebox{0.31}{\input{figures/stenosisTension7.tikz}} &
%    \scalebox{0.31}{\input{figures/stenosisTension9.tikz}} &
%    \scalebox{0.31}{\input{figures/stenosisTension11.tikz}} 
%    \fi
%  \end{tabular}
%  \caption{\label{fig:stenosisTension} The tension of the impermeable
%  (red) and semipermeable (dashed blue) vesicles in
%  Figure~\ref{fig:stenosisShape}. The point $\theta=0$ corresponds to
%  the right-most point of the vesicle.}
%\end{figure}
%
%\begin{figure}[htp]
%  \centering
%  \begin{tabular}{ccc}
%    \iftikz
%    \scalebox{0.31}{\input{figures/stenosisflux1.tikz}} &
%    \scalebox{0.31}{\input{figures/stenosisflux3.tikz}} &
%    \scalebox{0.31}{\input{figures/stenosisflux5.tikz}} \\
%    \scalebox{0.31}{\input{figures/stenosisflux7.tikz}} &
%    \scalebox{0.31}{\input{figures/stenosisflux9.tikz}} &
%    \scalebox{0.31}{\input{figures/stenosisflux11.tikz}} 
%    \fi
%  \end{tabular}
%  \caption{\label{fig:stenosisFlux} The flux (equivalent to pressure
%  jump) of the impermeable (red) and semipermeable (dashed blue)
%  vesicles in Figure~\ref{fig:stenosisShape}. The point $\theta=0$
%  corresponds to the right-most point of the vesicle.}
%\end{figure}
%This is verified for a
%vesicle with initial reduced area $\nu = 0.65$ and with different
%semipermeability rates. In Table~\ref{tbl:ellipseRelaxTimes} we report
%the required time, $t_\mathrm{steady}$, for the vesicle's reduced area
%to be within 1.3\% of its steady state reduced area, and the required
%time, $t_\mathrm{inflate}$, for the vesicle's reduced area to exceed
%1.3\% of its initial reduced area. Both times are proportional to
%$\beta^{-1}$, and for this reason, our numerical results are reported in
%terms of the scaled time $\beta t$.
%\begin{table}[htp]
%  \centering
%  \caption{\label{tbl:ellipseRelaxTimes} The time for a semipermeable
%  vesicle to its steady state ($t_\mathrm{steady}$) and to begin
%  inflating ($t_\mathrm{inflate}$).}
%  \begin{tabular}{C{0.8cm}C{2cm}C{0.8cm}C{2cm}C{0.8cm}}
%    $\beta$ & $t_\mathrm{steady}$  & ratio 
%            & $t_\mathrm{inflate}$ & ratio \\
%    \midrule
%    $10^{0}$  & $6.45 \times 10^{0}$ & ---  
%              & $2.17 \times 10^{-2}$ & --- \\ 
%    $10^{-1}$ & $2.00 \times 10^{1}$ & 3.10 & 
%                $1.25 \times 10^{-1}$ & 5.76 \\
%    $10^{-2}$ & $9.61 \times 10^{1}$ & 4.81 & 
%                $1.59 \times 10^{0}$ & 12.72 \\
%    $10^{-3}$ & $8.36 \times 10^{2}$ & 8.71 & 
%                $2.19 \times 10^{1}$ & 13.77 \\
%    $10^{-4}$ & $8.15 \times 10^{3}$ & 9.74 & 
%                $2.14 \times 10^{2}$ & 9.77 \\
%    $10^{-5}$ & $8.03 \times 10^{4}$ & 9.85 & 
%                $2.20 \times 10^{3}$ & 10.14 \\
%    \bottomrule
%  \end{tabular}
%\end{table}