### 2021-05-20 vesicle_code/docs/permeableVes/Response2Refs.tex
yny5

\documentclass[12pt]{article}
\usepackage{amsmath,amsfonts,bm,wrapfig}
\usepackage[left=1in,right=1in,top=1in,nohead]{geometry}
\usepackage[dvips]{graphicx}
\usepackage[
pdftitle={nsf si2 bie},
pdfcreator={pdftex},
pdfsubject={paper},
pdfkeywords={some,topics},
hyperindex = {true},  % turn to false for final
colorlinks = {true},  % turn to false for final
linkcolor = {blue},
pagecolor = {blue},
citecolor = {blue}
]{hyperref}
\usepackage{nsf}
\usepackage{paralist}

\usepackage[T1]{fontenc}
\usepackage{ae,aecompl}

\begin{document}

{\large \bf {\bf Response:} to the review report for PRF\_FB10127 Quaife: Hydrodynamics of a semipermeable inextensible membrane under flow and confinement}
\vspace{0.5cm}

We have revised the manuscript ``Hydrodynamics of a semipermeable inextensible membrane under flow and confinement"  
by Bryan Quaife, Ashley Gannon, and Y.-N. Young, 
based on 
the referees' very helpful feedback.  
Thanks to the referees' comments
and suggestions, we believe the revised manuscript is much
improved. 
In the revised manuscript we ensure that
all of their concerns and questions are properly addressed.
Below is a summary of the revisions made in the new manuscript.
The detailed responses to comments of referees are listed in 
responses to Referee 1 and Referee 2.
We hope that the referees and the
editor will now find the paper suitable for publication in 
Physical Review Fluids.

\vspace{0.5cm}
{\large \bf List of changes}
In the following we refer to Referee 1 who has five comments and Referee
2 who has three comments. In the revised manuscript the modifications in
the text are in magenta.

\begin{enumerate}
\item{Second paragraph in page 2, ``....giving rise to membrane rupture and leakage of macromolecules...".}

\item{Second to last paragraph in page 4, ``....suitable for a vesicle in a quiescent flow,"}

\item{New Figure 5(a).}

\item{New Figures 7.}

\item{New \S IV D.}

\item{New Figures 9.}

\item{New Figures 11, and the paragraph above the new Figures 11 in page 17.}

\item{New Figures 13, and the paragraph above the new Figures 13 in page 19.}

\item{Second to last paragraph in page 21.}

\item{Last paragraph in the Discussion session.}

\end{enumerate}

\newpage
\vspace{0.5cm}
{\large \bf {\bf Response:} to Referee 1:}
\vspace{0.5cm}

We thank the referee for their assessment of our work. The following is
our response to each of the specific comments and suggestions.

{\bf The referee wrote:
The paper deals with effect of membrane permeation under equilibrium
and flow conditions. This question is rarely addressed in literature.
The authors highlight the fact that red blood cells have a long life
time, and permeation may become decisive. The work report on several
interesting issues and should be of interest to the community working
on cell dynamics under flow. The paper is clearly written and results
very sound. I support publication after the authors have taken into
account the following points.
}

\noindent
{\bf Response:}: We thank the referee for the confirmation that the topic is of interest to researchers working on cell dynamics under flow.

\begin{enumerate}
\item {\bf 
The referee wrote:
``It is stated in the introduction that stretching (due to
    poration...) give rise to enhanced membrane permeability to
    both water and macromolecules such as ATP. Regarding ATP this is
    possible. In normal conditions ATP is released thanks to protein
    change of conformation that open channels and I am not sure the
    evoked permeability is a similar phenomenon. A few words of
    clarification are needed."}

\noindent
{\bf Response:}: In recent experiments and simulations by Yaling Liu (Lehigh University), it is reported that under very large stress the spectrin network under a red blood cell membrane can be destroyed and the red blood cell membrane can rupture. As a result the hemoglobin inside the RBC is found to reduce in this process. We now see that the source of confusion arises from the fact that we associate such leakage to enhanced membrane permeability to macromolecules.

We agree with the referee that the ATP release is mostly understood when
    it is through the coupling between Px1 (activated by membrane
    tension) and CFTR (activated by membrane deformation), and it is not
    yet well-understood how ATP can be released by diffusing through the
    membrane. We have thus modified this sentence to avoid confusion.

\noindent
{\bf Location:}  The second paragraph of the Introduction, five lines
    above Figure 1: ``....reaches the order of 2-4 mN/m, giving rise to
    membrane rupture and leakage of macromolecules (see Razizadeh {\it
    et al. }$^{20}$ and references therein)."

\item {\bf
The referee wrote:
``What cause more permeation? (i) hydrodynamic stress of suspending
fluid, (ii) tension (normal component of force, say Lagrange
multiplier), (ii) or the tangential force component due to tension
field? It would be nice to check this point."}

\noindent
{\bf Response:} We thank the referee for this point. In our
    formulation, the permeation flux is proportional to the hydrostatic
    pressure and the normal component of the membrane force, which is
    the capillary pressure due to the Lagrange multiplier tension. The
    tangential force component due to tension field does not contribute
    to the permeation flux. To elucidate the relative contribution
    between the hydrostatic pressure and the normal component of the
    membrane force, we have conducted computations and added a short
    paragraph to illustrate the relative contributions in all cases.

\noindent
{\bf Location:} A new \S IV D in page 14, and the new Figures 9 in page 15. The last paragraph in page 17, and the new Figure 11 in page 17. The second paragraph in page 19, and the new Figure 13 in page 19. Second to last paragraph in page 21.

\item {\bf 
The referee wrote:
``Figure 5 is not well explained. Depending on Capillary number it
seems that for all reduced area converge to the same (solid line)
value or continue to decrease (dashed line). I suggest both in text
and in caption to discuss this issue more clearly. Th result (Fig. 5a)
is also not intuitive: when capillary number is large (10) there is a
final reduced area, whereas for smaller value 1, the final reduced
area is smaller. I would have expected that smaller Cae would empty
less the vesicle. Is there any interpretation?"}

\noindent
{\bf Response:} We have mislabeled the $Ca_E$ in Figure 5:  $Ca_E=10^0$
for the solid curves and the $Ca_E=10$ for the dashed curves. To address
the effect of Capillary number on the relaxation of a semipermeable
vesicle, we have continued the simulations for the cases of $Ca_E = 10$
much further to better illustrate the evolution of reduced area towards
equilibrium.

\noindent
{\bf Location:} A new Figure 5(a) in page 11.

\item{\bf
The referee wrote:
``In Fig. 7 B reference is made to steady state? I guess there is
still water flow across membrane? Or the total flux is zero? In other
word is the volume of the shapes shown there the same along the blue
solid line? Then I am confused with inset c which shows variation of
alpha? My question: are shapes in b along blue line long term ones? Is
alpha constant there? Please clarify."}

\noindent
{\bf Response:} The shapes along the blue curve in Figure 7(b) are
equilibrium shapes that enclose a constant area but with an efflux
(blue) and an influx (red) that adds to zero. Inset (c) illustrates the
evolution from initial  the configuration (I) to the equilibrium (shape
III), which corresponds to the case with the largest displacement on the
blue curve. To better present all these results, we have taken out the
insets (c) and (d) and moved them to the new Figure 7(a) and (b). The
new Figure 7(c) and (d) are the old Figure 7(a) and (b).

\noindent
{\bf Location:} Figure 7 in page 13.

\item{\bf
The referee wrote:
``I do not quite understand why permeation time scale is much lower
for confined situations? Is pressure drop fixed, as compared to
previous cases? Does cell see stronger stress now? Why? If the driving
force is the same I do not expect such dramatic change of time scale,
since the mean shear stress experienced by the cell remains of the
same order, as before. A clear discussion of origin (from stress
values point of view, and why confinement would enhance imposed
stress) is necessary. This section is poorly discussed."}

\noindent
{\bf Response:}  We have added new computations to illustrate the relative contributions of membrane tension and bending forces to the permeating fluxes for all cases in \S IV and \S V (see {\bf Response} to the second point), see the new Figures 9, 11, and 13.

For the closely-fit channel the imposed flow at the inlet/outlet is
$U_{max}=1000$ $\mu$m/s to match the experimental conditions. As a
result the membrane tension is dominant over membrane bending in their
contribution to the permeating flux. For the contracting channel, the
imposed flow at the inlet/outlet is smaller and yet the strong
confinement still gives rise to a membrane tension that dominates in
contributing to the permeating flux. We also added new paragraphs to
discuss these new results.

\noindent
{\bf Location:} A new \S IV D in page 14, and the new Figures 9 in page 15. The last paragraph in page 17, and the new Figure 11 in page 17. The second paragraph in page 19, and the new Figure 13 in page 19. Second to last paragraph in page 21.

\end{enumerate}

\newpage
\vspace{0.5cm}
{\large \bf {\bf Response:} to Referee 2:}
\vspace{0.5cm}

We thank the referee for their assessment of our work.  The following is
our response to each of the specific comments and suggestions.

{\bf The referee wrote:
In this theoretical work, a 2D vesicle is simulated (via boundary elements) in unbounded and confined flows, while accounting for the finite permeability of the membrane to water. While this additional physics has, expectedly, little to no impact on the long-time physics (shape, tension) of vesicles in unbounded flows, in confined flows a marked dependence is observed on the permeability rate. In straight channel flow, the rate of water loss is proportional to the residence time in the channel, resulting in a reduction of the 2D reduced area at the channel outlet. In contraction flow, a high permeability rate (as well as a large pressure gradient) is required to force the vesicle through the nozzle and emerge in the open chamber downstream.

Overall, the paper is well written and the main results are clearly presented. A decent amount of time is spent covering vesicles in quiescent fluids and unbounded flows, which are less interesting than Section V, which includes the main interesting results on confined flows. However, these earlier sections help to paint a cohesive picture to put these later results in a clear context and, for that reason, their addition is welcome. My only real critique of the article is that all of the results are based on 2D simulations and, therefore, their applicability to real 3D systems is limited. However, the novelty of adding finite membrane permeability to the 2D boundary element context (which, to the best of my knowledge, has not been done before) merits publication of this article in Physical Review Fluids. I am therefore recommending this article for publication subject to optional revision (see my comments below).}

\noindent
{\bf Response:}: We thank the referee for the confirmation that the topic is interesting and results are useful.

\begin{enumerate}

\item {\bf The referee wrote: ``1. I am slightly confused as to how the length and time scales are derived from the parameters discussed on page 4. As the simulation is 2D, it would seem to me that all physical parameters can only be defined on a “per unit of length” basis. E.g., for a 2D vesicle contour, the surface tension is actually a line tension with units of N, not N/m. Could the authors clarify this?
In particular, reference is often made to how long things will take in the results (seconds,
minutes, days, weeks, etc.), which is why I ask how these time scales are obtained."}

\noindent
{\bf Response:}: We thank the referee for this question. The referee is indeed correct that in a 2D system the system remains constant in the $z$ direction. A two-dimensional vesicle membrane is a tube that is infinitely long in the $z$ direction.
On this infinitely long tube (of which the cross section is independent of $z$), the line tension is the total tangential force along the circumferential surface. Divide this line tension by the length of the tube would give us the membrane tension (force per unit length) in the $x-y$ plane.  
%
%Consequently it is customary to focus on the 2D system on a plane, and the membrane tension is the membrane force per unit length (the length in the $z$ direction). This means that, after the division by the length in the third ($z$) direction, the membrane tension in the plane is force per length.

The dimensionless group of this system is independent of the spatial dimensions of the system. In the absence of any external flow, we have adopted the initial vesicle size $R_0$ for a length scale,  the elastic relaxation in balance with the viscous dissipation at  length scale $R_0$ for a time scale $\tau$, and the elastic force at length scale $R_0$ for membrane tension. The dimensionless permeability is the hydraulic permeability scaled accordingly. Such scaling is natural for the quiescent cases. We used the same scaling for other cases for consistency in presentation to avoid confusion.

\noindent
{\bf Locations:} Second to last paragraph in page 4, fifth line in this paragraph, ``Under this nondimensionalization suitable for a vesicle in a quiescent flow".

\item {\bf The referee wrote: ``2. Could the authors shed some insight as to how these 2D simulations could be reliably connected to real experiments on any quantitative basis? Or is the connection purely
qualitative?"}

\noindent
{\bf Response:} The referee is correct that numerical simulations of
    two-dimensional fluid flow could be ideal and the connection to the
    real experiments is qualitative at best. The studies of
    two-dimensional vesicle hydrodynamics, however, have shown that the
    connection to real experiments or 3D simulations could be more than
    purely qualitative. For example, for a vesicle in a planar shear
    flow, the transition from tank-treading to tumbling is predicted in
    both 2- and 3-dimensions, at similar viscosity contrast. Ghiliotti,
    Biben, and Misbah (JFM, 2010) showed that results for a
    two-dimensional vesicle could be corrected systematically to agree
    with the three-dimensional results. More recently, the
    two-dimensional results in Zhang, Shen, Hogan, Barakat, and Misbah
    are also in more than qualitative agreement with real experiments
    (Biophys.~J, 2018).

    It is true that in some cases the two-dimensional vesicle hydrodynamics may have very little inference to the corresponding three-dimensional vesicle hydrodynamics. We have undertaken the full three-dimensional simulation of a semipermeable vesicle going through strong confinement, where the three-dimensional effects may be important.

\noindent
{\bf Location:} Last paragraph in the Discussion session.

\item {\bf The referee wrote: ``On the last page: ``We observe lose
  of"$\rightarrow$``We observe loss of"."}

\noindent
{\bf Response:} Thank you for catching this error. It has been
corrected.

\noindent
{\bf Location:} Second paragraph in page 21.

 \end{enumerate}

\end{document}